---
title: "Bayes-risk comparisons for eJAB"
author: "eJAB clinical-trials project"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
params:
  refresh: false
  dense_refresh: false
  bic_refresh: false
  seed: 20260724
  null_reps: 2000
  alt_reps: 500
  bic_null_reps: 2000
  bic_alt_reps: 400
  workers: 4
  bic_task_id: null
  bic_shard_dir: null
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 13,
  fig.height = 11,
  dpi = 160
)

library(dplyr)
library(ggplot2)
library(knitr)
library(patchwork)
library(purrr)
library(scales)
library(survival)
library(tidyr)

set.seed(params$seed)

output_dir <- "bayes_risk_outputs/revised_three_comparisons"
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
simulation_file <- file.path(output_dir, "bayes_risk_simulation.rds")
risk_file <- file.path(output_dir, "bayes_risk_grid.csv")
dense_file <- file.path(output_dir, "dense_bayes_risk_grid.rds")
dense_csv <- file.path(output_dir, "dense_bayes_risk_grid.csv")
bic_simulation_file <- file.path(
  output_dir, "large_sample_empirical_bic_simulation.rds"
)
bic_risk_file <- file.path(
  output_dir, "large_sample_empirical_bic_risk_grid.csv"
)

bf01_threshold <- 1
p_threshold <- 0.005
p05_threshold <- 0.05
n_grid <- c(30L, 60L, 120L, 250L, 500L)
theta_grid <- c(0.05, 0.10, 0.20, 0.35, 0.50, 0.80)
bic_n_grid <- unique(round(exp(seq(log(30), log(1e7), length.out = 70))))
bic_theta_grid <- unique(c(
  exp(seq(log(0.001), log(0.02), length.out = 14)),
  seq(0.025, 0.9, length.out = 46),
  seq(0.925, 2, by = 0.025)
))
dense_n_grid <- unique(round(exp(seq(log(30), log(1e7), length.out = 70))))
dense_theta_grid <- unique(c(
  exp(seq(log(0.001), log(0.02), length.out = 14)),
  seq(0.025, 0.9, length.out = 46),
  seq(0.925, 2, by = 0.025)
))
```

# Comparisons

This R Markdown file is the single source for four Bayes-risk comparisons:

1. \(eJAB_{01}\leq1\) versus \(p<0.005\), for all 10 tests.
2. \(eJAB_{01}\leq1\) versus \(p<0.05\), for all 10 tests.
3. \(eJAB_{01}\leq1\) versus \(BF_{01}^{BIC}\leq1\), for the six tests
   with an explicit parametric likelihood. Cox regression and the three rank
   tests are excluded.
4. \(eJAB_{01}\leq1\) versus \(TSBF_{01}\leq1\), for the \(2\times2\),
   \(3\times3\), and \(5\times5\) chi-square independence designs.

The TSBF comparison is restricted to these three chi-square independence
designs. Its test-statistic Bayes-factor function is transcribed from page 6 of
`Bayes_factor_function.pdf`, evaluated at the true standardized effect size,
and uses \(q=(R-1)(C-1)\) for each table dimension.

# Analysis scope and exclusions

```{r scope}
scope <- tribble(
  ~test, ~p_comparison, ~bic_comparison, ~tsbf_comparison, ~tsbf_source,
  "Two-sample t-test", "Included", "Included", "Excluded", NA_character_,
  "Simple linear regression", "Included", "Included", "Excluded", NA_character_,
  "Simple logistic regression", "Included", "Included", "Excluded", NA_character_,
  "Cox proportional-hazards regression", "Included", "Excluded; semiparametric partial likelihood", "Excluded", NA_character_,
  "Wilcoxon signed-rank test", "Included", "Excluded", "Excluded", NA_character_,
  "Mann-Whitney U-test", "Included", "Excluded", "Excluded", NA_character_,
  "One-way ANOVA", "Included", "Included", "Excluded", NA_character_,
  "Repeated-measures ANOVA", "Included", "Included", "Excluded", NA_character_,
  "Chi-square independence test", "Included", "Included", "Included for 2x2, 3x3, and 5x5 designs", "BFF PDF page 6",
  "Kruskal-Wallis test", "Included", "Excluded", "Excluded", NA_character_
)

kable(scope)
write.csv(
  scope,
  file.path(output_dir, "comparison_scope_and_exclusions.csv"),
  row.names = FALSE
)
```

Cox regression and the three rank tests are excluded from the BIC comparison,
which is restricted to tests with an explicit parametric likelihood. TSBF is
evaluated for the 2x2, 3x3, and 5x5 chi-square independence designs.

A simulated replicate is excluded before inspecting any decision-rule result
when the fitted model fails, its native test statistic is non-finite, its
\(p\)-value is non-finite or outside \([0,1]\), or its effective sample size is
not greater than one. No replicate is excluded because its \(p\)-value, test
statistic, effect size, eJAB, BIC BF, or TSBF is unusually large or small.

# eJAB and BIC functions

For eJAB, a test with \(q\) tested degrees of freedom maps its reported
two-sided or right-tail \(p\)-value to

\[
W=F^{-1}_{\chi_q^2}(1-p).
\]

The eJAB approximation is

\[
eJAB_{01}=\sqrt{n_{\mathrm{eff}}}
\exp\left[-\frac12(1-n_{\mathrm{eff}}^{-1/q})W\right]
\]

For the BIC comparison, let
\(\Lambda=2\{\ell(\widehat\theta_1)-\ell(\widehat\theta_0)\}\) be the
likelihood-ratio statistic calculated directly from the fitted alternative and
null models. The BIC Bayes-factor approximation is

\[
BF_{01}^{BIC}=n_{\mathrm{eff}}^{q/2}\exp(-\Lambda/2).
\]

Thus, the BIC decision does not substitute an asymptotic chi-square quantile
for the fitted-model likelihood ratio.

```{r ejab-bic-functions}
clip_p <- function(p) {
  pmin(pmax(p, .Machine$double.xmin), 1 - .Machine$double.eps)
}

ejab01 <- function(p, n_eff, q) {
  W <- qchisq(clip_p(p), df = q, lower.tail = FALSE)
  sqrt(n_eff) * exp(-0.5 * (1 - n_eff^(-1 / q)) * W)
}

bic_bf01_from_lr <- function(lr_statistic, n_eff, q) {
  n_eff^(q / 2) * exp(-pmax(lr_statistic, 0) / 2)
}
```

# Test-statistic Bayes-factor functions

The following functions transcribe the seven BFFs from
`Bayes_factor_function.pdf`. The PDF reports \(BFF_{10}\); the analysis uses
\(TSBF_{01}=1/BFF_{10}\).

```{r tsbf-functions}
bff10_t <- function(t, nu, tau2) {
  r <- 1 + t^2 / nu
  u <- 1 + t^2 / (nu * (1 + tau2))
  k <- tau2 * (nu + 1) / (nu * (1 + tau2))
  exp(
    -1.5 * log1p(tau2) +
      ((nu + 1) / 2) * (log(r) - log(u)) +
      log1p(k * t^2 / u)
  )
}

bff10_z <- function(z, tau2) {
  exp(
    -1.5 * log1p(tau2) +
      log1p(tau2 * z^2 / (1 + tau2)) +
      tau2 * z^2 / (2 * (1 + tau2))
  )
}

bff10_anova <- function(F_stat, groups, n_total, tau2) {
  v <- (n_total - groups) * (1 + tau2)
  exp(
    -((groups + 1) / 2) * log1p(tau2) +
      ((n_total - 1) / 2) *
        (log1p((groups - 1) * F_stat / (n_total - groups)) -
           log1p((groups - 1) * F_stat / v)) +
      log1p((n_total - 1) * tau2 * F_stat /
              (v + (groups - 1) * F_stat))
  )
}

bff10_ranova <- function(F_stat, conditions, subjects, tau2) {
  u <- (subjects - 1) * (1 + tau2)
  exp(
    -((conditions + 1) / 2) * log1p(tau2) +
      (((conditions - 1) * subjects) / 2) *
        (log1p(F_stat / (subjects - 1)) - log1p(F_stat / u)) +
      log1p(subjects * tau2 * F_stat / (u + F_stat))
  )
}

bff10_chisq <- function(chisq_stat, q, tau2) {
  exp(
    -(q / 2 + 1) * log1p(tau2) +
      log1p(tau2 * chisq_stat / (q * (1 + tau2))) +
      tau2 * chisq_stat / (2 * (1 + tau2))
  )
}

bff10_cox <- function(lambda, tau2) {
  bff10_z(sqrt(pmax(lambda, 0)), tau2)
}

yuan_johnson_normal_bf01 <- function(T_star) {
  sqrt(2) * exp(-T_star^2 / 4)
}

yuan_johnson_kw_bf01 <- function(W, groups, tau = 1) {
  q <- groups - 1
  (1 + tau)^(q / 2) * exp(-tau * W / (2 * (1 + tau)))
}

stopifnot(
  all.equal(bff10_t(2, 28, 0), 1),
  all.equal(bff10_z(2, 0), 1),
  all.equal(bff10_anova(2, 4, 40, 0), 1),
  all.equal(bff10_ranova(2, 4, 20, 0), 1),
  all.equal(bff10_chisq(5, 4, 0), 1),
  all.equal(bff10_cox(4, 0), 1),
  all.equal(yuan_johnson_normal_bf01(0), sqrt(2)),
  all.equal(yuan_johnson_kw_bf01(0, 5), 4)
)
```

# Simulation models

All tests use their native row-level statistic. The three rank tests use
heavy-tailed \(t_3\) errors with variance one, matching the earlier simulation.

```{r simulation-models}
test_labels <- c(
  ttest = "Two-sample t-test",
  lm = "Linear regression",
  logistic = "Logistic regression",
  cox = "Cox PH regression",
  wilcoxon = "Wilcoxon signed-rank",
  mannwhitney = "Mann-Whitney",
  anova = "One-way ANOVA",
  ranova = "Repeated-measures ANOVA",
  chisq = "Chi-square independence",
  kruskal = "Kruskal-Wallis"
)

q_by_test <- c(
  ttest = 1L, lm = 1L, logistic = 1L, cox = 1L,
  wilcoxon = 1L, mannwhitney = 1L,
  anova = 3L, ranova = 3L, chisq = 4L, kruskal = 4L
)

bic_test_keys <- c(
  "ttest", "lm", "logistic",
  "anova", "ranova", "chisq"
)
tsbf_test_keys <- "chisq"

contrast_scores <- function(k) {
  x <- seq_len(k) - (k + 1) / 2
  x / sqrt(mean(x^2))
}

chisq_probabilities <- function(theta) {
  log_p <- matrix(0, 3, 3)
  diag(log_p) <- theta
  p <- exp(log_p)
  p / sum(p)
}

chisq_omega2 <- function(theta) {
  p <- chisq_probabilities(theta)
  expected <- outer(rowSums(p), colSums(p))
  sum((p - expected)^2 / expected)
}

chisq_tau2 <- function(n, omega2, q) {
  n * omega2 / q
}

tau2_for <- function(test_key, n, theta, simulation_result = NULL) {
  if (test_key == "ttest") {
    n1 <- floor(n / 2)
    n2 <- n - n1
    return(theta^2 / (2 * (1 / n1 + 1 / n2)))
  }
  if (test_key == "lm") return(n * theta^2 / 2)
  if (test_key == "logistic") return(n * (theta / 2)^2 / 2)
  if (test_key == "cox") {
    events <- simulation_result$events
    return(theta^2 * events / 4)
  }
  if (test_key == "anova") {
    groups <- 4
    scores <- contrast_scores(groups)
    sizes <- rep(floor(n / groups), groups)
    if (n %% groups > 0) {
      sizes[seq_len(n %% groups)] <- sizes[seq_len(n %% groups)] + 1
    }
    weights <- sizes / sum(sizes)
    means <- theta * scores
    omega2 <- sum(weights * (means - sum(weights * means))^2) / (groups - 1)
    return(n * omega2 / 2)
  }
  if (test_key == "ranova") {
    conditions <- 4
    means <- theta * contrast_scores(conditions)
    omega2 <- sum((means - mean(means))^2) / (conditions - 1)
    return(n * omega2 / 2)
  }
  if (test_key == "chisq") {
    return(chisq_tau2(
      n,
      chisq_omega2(theta),
      unname(q_by_test[test_key])
    ))
  }
  NA_real_
}

simulate_one <- function(test_key, n, theta) {
  tryCatch({
    if (test_key == "ttest") {
      n1 <- floor(n / 2)
      n2 <- n - n1
      nu <- n - 2
      noncentrality <- theta / sqrt(1 / n1 + 1 / n2)
      statistic <- rt(1, nu, ncp = noncentrality)
      return(list(
        statistic = statistic,
        p = 2 * pt(-abs(statistic), nu),
        n_eff = n,
        nu = nu,
        bic_lr = n * log1p(statistic^2 / nu)
      ))
    }

    if (test_key == "lm") {
      nu <- n - 2
      centered_x_ss <- rchisq(1, n - 1)
      statistic <- rt(
        1, nu,
        ncp = theta * sqrt(centered_x_ss)
      )
      return(list(
        statistic = statistic,
        p = 2 * pt(-abs(statistic), nu),
        n_eff = n,
        nu = nu,
        bic_lr = n * log1p(statistic^2 / nu)
      ))
    }

    if (test_key == "logistic") {
      # A balanced binary predictor makes the two binomial group counts exact
      # sufficient statistics. This evaluates the same parametric likelihood
      # without allocating an n-row dataset at sample sizes up to 10^7.
      sizes <- c(floor(n / 2), n - floor(n / 2))
      probabilities <- plogis(c(-theta / 2, theta / 2))
      successes <- rbinom(2, sizes, probabilities)
      fitted_probabilities <- successes / sizes
      pooled_probability <- sum(successes) / n
      full_loglik <- sum(dbinom(
        successes, sizes, fitted_probabilities, log = TRUE
      ))
      null_loglik <- sum(dbinom(
        successes, sizes, pooled_probability, log = TRUE
      ))
      likelihood_ratio <- pmax(2 * (full_loglik - null_loglik), 0)
      return(list(
        statistic = sqrt(likelihood_ratio),
        p = pchisq(likelihood_ratio, 1, lower.tail = FALSE),
        n_eff = n,
        bic_lr = likelihood_ratio
      ))
    }

    if (test_key == "cox") {
      x <- sample(rep(c(0, 1), length.out = n))
      event_time <- rexp(n, rate = 0.1 * exp(theta * x))
      censor_time <- rexp(n, rate = 0.025)
      time <- pmin(event_time, censor_time)
      event <- as.integer(event_time <= censor_time)
      fit <- suppressWarnings(coxph(
        Surv(time, event) ~ x,
        ties = "efron",
        control = coxph.control(iter.max = 30)
      ))
      lambda <- pmax(unname(2 * diff(fit$loglik)), 0)
      events <- sum(event)
      return(list(
        statistic = lambda,
        p = pchisq(lambda, 1, lower.tail = FALSE),
        n_eff = events,
        events = events,
        bic_lr = lambda
      ))
    }

    if (test_key == "wilcoxon") {
      d <- theta + rt(n, 3) / sqrt(3)
      ranks <- rank(abs(d), ties.method = "average")
      T_stat <- sum(ranks[d > 0])
      mean0 <- n * (n + 1) / 4
      sd0 <- sqrt(n * (n + 1) * (2 * n + 1) / 24)
      T_star <- (T_stat - mean0) / sd0
      return(list(
        statistic = T_star,
        p = 2 * pnorm(-abs(T_star)),
        n_eff = n,
        bic_lr = NA_real_
      ))
    }

    if (test_key == "mannwhitney") {
      n1 <- floor(n / 2)
      n2 <- n - n1
      x <- rt(n1, 3) / sqrt(3)
      y <- theta + rt(n2, 3) / sqrt(3)
      combined_ranks <- rank(c(x, y), ties.method = "average")
      T_stat <- sum(combined_ranks[seq_len(n1)])
      mean0 <- n1 * (n + 1) / 2
      sd0 <- sqrt(n1 * n2 * (n + 1) / 12)
      T_star <- (T_stat - mean0) / sd0
      return(list(
        statistic = T_star,
        p = 2 * pnorm(-abs(T_star)),
        n_eff = n,
        bic_lr = NA_real_
      ))
    }

    if (test_key == "anova") {
      groups <- 4
      sizes <- rep(floor(n / groups), groups)
      if (n %% groups > 0) {
        sizes[seq_len(n %% groups)] <- sizes[seq_len(n %% groups)] + 1
      }
      means <- theta * contrast_scores(groups)
      weighted_mean <- sum(sizes * means) / n
      noncentrality <- sum(sizes * (means - weighted_mean)^2)
      df1 <- groups - 1
      df2 <- n - groups
      statistic <- rf(1, df1, df2, ncp = noncentrality)
      return(list(
        statistic = statistic,
        p = pf(statistic, df1, df2, lower.tail = FALSE),
        n_eff = n,
        bic_lr = n * log1p(df1 * statistic / df2)
      ))
    }

    if (test_key == "ranova") {
      conditions <- 4
      means <- theta * contrast_scores(conditions)
      df1 <- conditions - 1
      df2 <- (conditions - 1) * (n - 1)
      noncentrality <- n * sum((means - mean(means))^2)
      statistic <- rf(1, df1, df2, ncp = noncentrality)
      return(list(
        statistic = statistic,
        p = pf(statistic, df1, df2, lower.tail = FALSE),
        n_eff = n * (conditions - 1),
        bic_lr = n * conditions * log1p(df1 * statistic / df2)
      ))
    }

    if (test_key == "chisq") {
      counts <- matrix(
        rmultinom(1, n, as.vector(chisq_probabilities(theta))),
        3, 3
      )
      fit <- suppressWarnings(chisq.test(counts, correct = FALSE))
      statistic <- unname(fit$statistic)
      positive <- counts > 0 & fit$expected > 0
      likelihood_ratio <- 2 * sum(
        counts[positive] * log(counts[positive] / fit$expected[positive])
      )
      return(list(
        statistic = statistic,
        p = pchisq(statistic, 4, lower.tail = FALSE),
        n_eff = n,
        bic_lr = pmax(likelihood_ratio, 0)
      ))
    }

    if (test_key == "kruskal") {
      groups <- 5
      sizes <- rep(floor(n / groups), groups)
      if (n %% groups > 0) {
        sizes[seq_len(n %% groups)] <- sizes[seq_len(n %% groups)] + 1
      }
      means <- theta * contrast_scores(groups)
      group <- factor(rep(seq_len(groups), sizes))
      y <- rep(means, sizes) + rt(n, 3) / sqrt(3)
      fit <- kruskal.test(y ~ group)
      statistic <- unname(fit$statistic)
      return(list(
        statistic = statistic,
        p = pchisq(statistic, 4, lower.tail = FALSE),
        n_eff = n,
        bic_lr = NA_real_
      ))
    }

    stop("Unknown test key")
  }, error = function(e) {
    list(
      statistic = NA_real_,
      p = NA_real_,
      n_eff = NA_real_,
      bic_lr = NA_real_
    )
  })
}
```

# Decision rules and Bayes-risk estimation

```{r risk-functions}
valid_result <- function(x, test_key) {
  is.finite(x$statistic) &&
    is.finite(x$p) && x$p >= 0 && x$p <= 1 &&
    is.finite(x$n_eff) && x$n_eff > 1 &&
    (
      !test_key %in% bic_test_keys ||
        (is.finite(x$bic_lr) && x$bic_lr >= 0)
    )
}

tsbf01 <- function(test_key, result, n, theta) {
  if (test_key == "wilcoxon" || test_key == "mannwhitney") {
    return(yuan_johnson_normal_bf01(result$statistic))
  }
  if (test_key == "kruskal") {
    return(yuan_johnson_kw_bf01(result$statistic, groups = 5, tau = 1))
  }

  tau2 <- tau2_for(test_key, n, theta, result)
  bf10 <- switch(
    test_key,
    ttest = bff10_t(result$statistic, result$nu, tau2),
    lm = bff10_t(result$statistic, result$nu, tau2),
    logistic = bff10_z(result$statistic, tau2),
    cox = bff10_cox(result$statistic, tau2),
    anova = bff10_anova(result$statistic, 4, n, tau2),
    ranova = bff10_ranova(result$statistic, 4, n, tau2),
    chisq = bff10_chisq(result$statistic, 4, tau2),
    stop("Unknown test key")
  )
  1 / bf10
}

decisions_for <- function(draws, test_key, n, theta) {
  q <- unname(q_by_test[test_key])
  p <- map_dbl(draws, "p")
  n_eff <- map_dbl(draws, "n_eff")
  ejab <- ejab01(p, n_eff, q)
  bic <- if (test_key %in% bic_test_keys) {
    bic_bf01_from_lr(map_dbl(draws, "bic_lr"), n_eff, q)
  } else {
    rep(NA_real_, length(draws))
  }
  tsbf <- if (test_key %in% tsbf_test_keys) {
    map_dbl(draws, ~ tsbf01(test_key, .x, n, theta))
  } else {
    rep(NA_real_, length(draws))
  }

  tibble(
    reject_p = p < p_threshold,
    reject_ejab = ejab <= bf01_threshold,
    reject_bic = bic <= bf01_threshold,
    reject_tsbf = tsbf <= bf01_threshold
  )
}

simulate_test_n <- function(test_key, n, null_reps, alt_reps, seed,
                            effect_grid = theta_grid) {
  set.seed(seed)

  null_draws <- replicate(
    null_reps,
    simulate_one(test_key, n, 0),
    simplify = FALSE
  )
  null_draws <- null_draws[
    map_lgl(null_draws, valid_result, test_key = test_key)
  ]

  map_dfr(effect_grid, function(theta) {
    alt_draws <- replicate(
      alt_reps,
      simulate_one(test_key, n, theta),
      simplify = FALSE
    )
    alt_draws <- alt_draws[
      map_lgl(alt_draws, valid_result, test_key = test_key)
    ]

    if (!length(null_draws) || !length(alt_draws)) {
      return(tibble(
        test_key = test_key,
        test = unname(test_labels[test_key]),
        n = n,
        theta = theta,
        null_valid_reps = length(null_draws),
        alt_valid_reps = length(alt_draws)
      ))
    }

    null_decisions <- decisions_for(null_draws, test_key, n, theta)
    alt_decisions <- decisions_for(alt_draws, test_key, n, theta)

    alpha_p <- mean(null_decisions$reject_p)
    power_p <- mean(alt_decisions$reject_p)
    alpha_ejab <- mean(null_decisions$reject_ejab)
    power_ejab <- mean(alt_decisions$reject_ejab)
    if (test_key %in% tsbf_test_keys) {
      alpha_tsbf <- mean(null_decisions$reject_tsbf)
      power_tsbf <- mean(alt_decisions$reject_tsbf)
    } else {
      alpha_tsbf <- NA_real_
      power_tsbf <- NA_real_
    }

    if (test_key %in% bic_test_keys) {
      alpha_bic <- mean(null_decisions$reject_bic)
      power_bic <- mean(alt_decisions$reject_bic)
    } else {
      alpha_bic <- NA_real_
      power_bic <- NA_real_
    }

    risk_p <- 0.5 * (alpha_p + 1 - power_p)
    risk_ejab <- 0.5 * (alpha_ejab + 1 - power_ejab)
    risk_bic <- 0.5 * (alpha_bic + 1 - power_bic)
    risk_tsbf <- 0.5 * (alpha_tsbf + 1 - power_tsbf)

    tibble(
      test_key = test_key,
      test = unname(test_labels[test_key]),
      q = unname(q_by_test[test_key]),
      n = n,
      theta = theta,
      null_valid_reps = length(null_draws),
      alt_valid_reps = length(alt_draws),
      alpha_p = alpha_p,
      power_p = power_p,
      alpha_ejab = alpha_ejab,
      power_ejab = power_ejab,
      alpha_bic = alpha_bic,
      power_bic = power_bic,
      alpha_tsbf = alpha_tsbf,
      power_tsbf = power_tsbf,
      risk_p = risk_p,
      risk_ejab = risk_ejab,
      risk_bic = risk_bic,
      risk_tsbf = risk_tsbf,
      delta_p_minus_ejab = risk_p - risk_ejab,
      delta_bic_minus_ejab = risk_bic - risk_ejab,
      delta_tsbf_minus_ejab = risk_tsbf - risk_ejab
    )
  })
}
```

Null draws are generated once for each test and sample size and reused across
the six effect-size cells. For the in-scope chi-square TSBF comparison, the
false-positive rate is recalculated for each effect size because the BFF prior
scale depends on the standardized alternative effect.

```{r run-simulation}
task_grid <- crossing(
  test_key = names(test_labels),
  n = n_grid
) |>
  mutate(seed = params$seed + row_number())

if (isTRUE(params$refresh) || !file.exists(simulation_file)) {
  tasks <- split(task_grid, seq_len(nrow(task_grid)))
  run_task <- function(task) {
    simulate_test_n(
      task$test_key,
      task$n,
      params$null_reps,
      params$alt_reps,
      task$seed
    )
  }

  if (.Platform$OS.type != "windows" && params$workers > 1) {
    result_list <- parallel::mclapply(
      tasks,
      run_task,
      mc.cores = params$workers
    )
  } else {
    result_list <- lapply(tasks, run_task)
  }

  risk <- bind_rows(result_list)
  saveRDS(risk, simulation_file)
} else {
  risk <- readRDS(simulation_file)
}

write.csv(risk, risk_file, row.names = FALSE)

validation_summary <- risk |>
  summarise(
    cells = n(),
    minimum_null_valid_reps = min(null_valid_reps),
    minimum_alt_valid_reps = min(alt_valid_reps),
    nonfinite_required_risks = sum(
      !is.finite(risk_p) | !is.finite(risk_ejab)
    ),
    nonfinite_bic_risks_in_scope = sum(
      test_key %in% bic_test_keys & !is.finite(risk_bic)
    ),
    nonfinite_tsbf_risks_in_scope = sum(
      test_key %in% tsbf_test_keys & !is.finite(risk_tsbf)
    )
  )

kable(validation_summary)
```

## Dense large-sample empirical BIC grid

The BIC comparison uses the same \(70\times104\) extent as the other dense
maps: 70 logarithmically spaced sample sizes from 30 to \(10^7\) and 104 effect
sizes from 0.001 to 2.
Every cell is estimated from model-based Monte Carlo draws and direct
likelihood ratios. The Gaussian panels use exact finite-sample sufficient-
statistic distributions that are algebraically identical to refitting the
nested models. Logistic regression uses exact grouped-binomial sufficient
statistics, and the contingency-table panel uses multinomial counts. No
interpolation or asymptotic chi-square substitution is used to densify the BIC
figure.

```{r dense-empirical-bic-grid}
bic_task_grid <- crossing(
  test_key = bic_test_keys,
  n = bic_n_grid
) |>
  mutate(seed = params$seed + 100000L + row_number())

run_bic_task <- function(task) {
  simulate_test_n(
    task$test_key,
    task$n,
    params$bic_null_reps,
    params$bic_alt_reps,
    task$seed,
    effect_grid = bic_theta_grid
  )
}

# Slurm array mode renders only far enough to evaluate this chunk. Each array
# element owns exactly one deterministic (test_key, n) task and writes its
# result atomically, so interrupted tasks cannot leave plausible partial data.
if (!is.null(params$bic_task_id)) {
  task_id <- as.integer(params$bic_task_id)
  if (length(task_id) != 1L || is.na(task_id) ||
      task_id < 1L || task_id > nrow(bic_task_grid)) {
    stop("bic_task_id must be between 1 and ", nrow(bic_task_grid))
  }
  if (is.null(params$bic_shard_dir) || !nzchar(params$bic_shard_dir)) {
    stop("bic_shard_dir is required in Slurm array mode")
  }

  dir.create(params$bic_shard_dir, recursive = TRUE, showWarnings = FALSE)
  shard <- run_bic_task(bic_task_grid[task_id, , drop = FALSE]) |>
    mutate(task_id = task_id, .before = 1)
  shard_file <- file.path(
    params$bic_shard_dir, sprintf("bic_task_%04d.rds", task_id)
  )
  temporary_file <- paste0(shard_file, ".tmp-", Sys.getpid())
  saveRDS(shard, temporary_file)
  if (!file.rename(temporary_file, shard_file)) {
    unlink(temporary_file)
    stop("Could not atomically publish shard: ", shard_file)
  }
  knitr::knit_exit()
}

if (isTRUE(params$bic_refresh) || !file.exists(bic_simulation_file)) {
  bic_tasks <- split(bic_task_grid, seq_len(nrow(bic_task_grid)))

  if (.Platform$OS.type != "windows" && params$workers > 1) {
    bic_result_list <- parallel::mclapply(
      bic_tasks,
      run_bic_task,
      mc.cores = params$workers
    )
  } else {
    bic_result_list <- lapply(bic_tasks, run_bic_task)
  }

  bic_risk <- bind_rows(bic_result_list)
  saveRDS(bic_risk, bic_simulation_file)
} else {
  bic_risk <- readRDS(bic_simulation_file)
}

write.csv(bic_risk, bic_risk_file, row.names = FALSE)

bic_validation_summary <- bic_risk |>
  summarise(
    cells = n(),
    minimum_null_valid_reps = min(null_valid_reps),
    minimum_alt_valid_reps = min(alt_valid_reps),
    nonfinite_bic_risks = sum(!is.finite(risk_bic))
  )

kable(bic_validation_summary)
```

# Dense large-\(n\) criticality grid

The publication-style maps use a 70-point logarithmic sample-size grid from
\(30\) to \(10^7\) and a 104-point raw effect-size grid from \(0.001\) to
\(2\). At these sample sizes, row-level simulation is not computationally
meaningful. Instead, rejection probabilities are evaluated from each test's
native central/noncentral statistic distribution. The TSBF is always evaluated
from the relevant test-statistic BFF; no generic \(p\)-to-chi-square TSBF is
used.

```{r dense-distributions}
t3_scale <- 1 / sqrt(3)
t3_pdf <- function(x) dt(x / t3_scale, df = 3) / t3_scale
t3_cdf <- function(x) pt(x / t3_scale, df = 3)

mw_ncp_per_n <- function(theta) {
  if (theta == 0) return(0)
  area <- integrate(
    function(e) t3_cdf(theta + e) * t3_pdf(e),
    -Inf, Inf, rel.tol = 1e-8, subdivisions = 300L
  )$value
  3 * (area - 0.5)^2
}

wilcoxon_ncp_per_n <- function(theta) {
  if (theta == 0) return(0)
  abs_cdf_shifted <- function(x) {
    t3_cdf(x - theta) - t3_cdf(-x - theta)
  }
  area <- integrate(
    function(d) abs_cdf_shifted(d) * t3_pdf(d - theta),
    0, Inf, rel.tol = 1e-8, subdivisions = 300L
  )$value
  12 * (area - 0.25)^2
}

kruskal_ncp_per_n <- function(theta) {
  if (theta == 0) return(0)
  locations <- theta * contrast_scores(5)
  pooled_cdf <- function(x) {
    rowMeans(vapply(locations, function(loc) t3_cdf(x - loc), numeric(length(x))))
  }
  rank_means <- vapply(locations, function(loc) {
    integrate(
      function(x) pooled_cdf(x) * t3_pdf(x - loc),
      -Inf, Inf, rel.tol = 1e-8, subdivisions = 300L
    )$value
  }, numeric(1))
  12 * mean((rank_means - 0.5)^2)
}

rank_ncp <- tibble(theta = c(0, dense_theta_grid)) |>
  mutate(
    wilcoxon = map_dbl(theta, wilcoxon_ncp_per_n),
    mannwhitney = map_dbl(theta, mw_ncp_per_n),
    kruskal = map_dbl(theta, kruskal_ncp_per_n)
  )

rank_ncp_value <- function(test_key, theta) {
  rank_ncp[[test_key]][match(theta, rank_ncp$theta)]
}

cox_event_fraction_dense <- function(theta) {
  0.5 * (
    0.1 / (0.1 + 0.025) +
      (0.1 * exp(theta)) / (0.1 * exp(theta) + 0.025)
  )
}

n_eff_dense <- function(test_key, n, theta) {
  if (test_key == "cox") return(n * cox_event_fraction_dense(theta))
  if (test_key == "ranova") return(3 * n)
  n
}

native_ncp <- function(test_key, n, theta) {
  if (theta == 0) return(0)
  switch(
    test_key,
    ttest = theta / sqrt(1 / floor(n / 2) + 1 / (n - floor(n / 2))),
    lm = sqrt(n) * theta,
    logistic = sqrt(n) * theta / 2,
    cox = n_eff_dense("cox", n, theta) * theta^2 / 4,
    wilcoxon = sqrt(n * rank_ncp_value("wilcoxon", theta)),
    mannwhitney = sqrt(n * rank_ncp_value("mannwhitney", theta)),
    anova = n * theta^2,
    ranova = 4 * n * theta^2,
    chisq = n * chisq_omega2(theta),
    kruskal = n * rank_ncp_value("kruskal", theta),
    stop("Unknown test key")
  )
}

native_rejection_at_stat <- function(test_key, n, theta, critical) {
  if (!is.finite(critical)) return(0)
  if (critical <= 0) return(1)
  ncp <- native_ncp(test_key, n, theta)

  probability <- switch(
    test_key,
    ttest = {
      df <- n - 2
      pt(-critical, df = df, ncp = ncp) +
        pt(critical, df = df, ncp = ncp, lower.tail = FALSE)
    },
    lm = {
      df <- n - 2
      pt(-critical, df = df, ncp = ncp) +
        pt(critical, df = df, ncp = ncp, lower.tail = FALSE)
    },
    logistic = {
      pnorm(-critical, mean = ncp) +
        pnorm(critical, mean = ncp, lower.tail = FALSE)
    },
    cox = pchisq(critical, df = 1, ncp = ncp, lower.tail = FALSE),
    wilcoxon = {
      pnorm(-critical, mean = ncp) +
        pnorm(critical, mean = ncp, lower.tail = FALSE)
    },
    mannwhitney = {
      pnorm(-critical, mean = ncp) +
        pnorm(critical, mean = ncp, lower.tail = FALSE)
    },
    anova = pf(
      critical, df1 = 3, df2 = n - 4, ncp = ncp, lower.tail = FALSE
    ),
    ranova = pf(
      critical, df1 = 3, df2 = 3 * (n - 1),
      ncp = ncp, lower.tail = FALSE
    ),
    chisq = pchisq(critical, df = 4, ncp = ncp, lower.tail = FALSE),
    kruskal = pchisq(critical, df = 4, ncp = ncp, lower.tail = FALSE),
    stop("Unknown test key")
  )
  pmin(pmax(probability, 0), 1)
}

native_critical_from_p <- function(test_key, n, p_cut) {
  p_cut <- clip_p(p_cut)
  switch(
    test_key,
    ttest = qt(1 - p_cut / 2, df = n - 2),
    lm = qt(1 - p_cut / 2, df = n - 2),
    logistic = qnorm(1 - p_cut / 2),
    cox = qchisq(1 - p_cut, df = 1),
    wilcoxon = qnorm(1 - p_cut / 2),
    mannwhitney = qnorm(1 - p_cut / 2),
    anova = qf(1 - p_cut, df1 = 3, df2 = n - 4),
    ranova = qf(1 - p_cut, df1 = 3, df2 = 3 * (n - 1)),
    chisq = qchisq(1 - p_cut, df = 4),
    kruskal = qchisq(1 - p_cut, df = 4),
    stop("Unknown test key")
  )
}

p_cut_ejab <- function(n_eff, q) {
  W <- (log(n_eff) - 2 * log(bf01_threshold)) /
    (1 - n_eff^(-1 / q))
  pchisq(W, df = q, lower.tail = FALSE)
}

```

```{r dense-tsbf-thresholds}
log_bff10_t <- function(t, nu, tau2) {
  r <- 1 + t^2 / nu
  u <- 1 + t^2 / (nu * (1 + tau2))
  k <- tau2 * (nu + 1) / (nu * (1 + tau2))
  -1.5 * log1p(tau2) +
    ((nu + 1) / 2) * (log(r) - log(u)) +
    log1p(k * t^2 / u)
}

log_bff10_z <- function(z, tau2) {
  -1.5 * log1p(tau2) +
    log1p(tau2 * z^2 / (1 + tau2)) +
    tau2 * z^2 / (2 * (1 + tau2))
}

log_bff10_anova <- function(F_stat, groups, n_total, tau2) {
  v <- (n_total - groups) * (1 + tau2)
  -((groups + 1) / 2) * log1p(tau2) +
    ((n_total - 1) / 2) *
      (log1p((groups - 1) * F_stat / (n_total - groups)) -
         log1p((groups - 1) * F_stat / v)) +
    log1p((n_total - 1) * tau2 * F_stat /
            (v + (groups - 1) * F_stat))
}

log_bff10_ranova <- function(F_stat, conditions, subjects, tau2) {
  u <- (subjects - 1) * (1 + tau2)
  -((conditions + 1) / 2) * log1p(tau2) +
    (((conditions - 1) * subjects) / 2) *
      (log1p(F_stat / (subjects - 1)) - log1p(F_stat / u)) +
    log1p(subjects * tau2 * F_stat / (u + F_stat))
}

log_bff10_chisq <- function(chisq_stat, q, tau2) {
  -(q / 2 + 1) * log1p(tau2) +
    log1p(tau2 * chisq_stat / (q * (1 + tau2))) +
    tau2 * chisq_stat / (2 * (1 + tau2))
}

critical_root <- function(log_bf_function) {
  target <- log(1 / bf01_threshold)
  if (log_bf_function(0) >= target) return(0)
  upper <- 1
  value <- log_bf_function(upper)
  while ((!is.finite(value) || value < target) && upper < 1e7) {
    upper <- upper * 2
    value <- log_bf_function(upper)
  }
  if (!is.finite(value) || value < target) return(Inf)
  uniroot(
    function(x) log_bf_function(x) - target,
    interval = c(0, upper),
    tol = 1e-8
  )$root
}

estimated_omega_tsbf_critical <- function(q) {
  target <- log(1 / bf01_threshold)
  objective <- function(x) {
    log_bff10_chisq(x, q, tau2 = x / q) - target
  }

  # With BF01 = 1, x = 0 is a measure-zero equality point. For positive x,
  # the plug-in BFF first falls below 1 and then crosses back above 1. The
  # rejection region is determined by this positive upper crossing.
  lower <- sqrt(.Machine$double.eps)
  if (!is.finite(objective(lower)) || objective(lower) >= 0) {
    stop("Could not bracket the positive plug-in TSBF crossing.")
  }

  upper <- max(1, q)
  value <- objective(upper)
  while ((!is.finite(value) || value < 0) && upper < 1e7) {
    upper <- upper * 2
    value <- objective(upper)
  }
  if (!is.finite(value) || value < 0) return(Inf)

  uniroot(objective, interval = c(lower, upper), tol = 1e-8)$root
}

dense_tau2 <- function(test_key, n, theta, data_theta = theta) {
  switch(
    test_key,
    ttest = {
      n1 <- floor(n / 2)
      n2 <- n - n1
      theta^2 / (2 * (1 / n1 + 1 / n2))
    },
    lm = n * theta^2 / 2,
    logistic = n * theta^2 / 8,
    cox = n_eff_dense("cox", n, data_theta) * theta^2 / 4,
    anova = n * theta^2 / 6,
    ranova = 2 * n * theta^2 / 3,
    chisq = chisq_tau2(
      n,
      chisq_omega2(theta),
      unname(q_by_test[test_key])
    ),
    NA_real_
  )
}

dense_tsbf_critical <- function(test_key, n, theta, data_theta = theta) {
  if (test_key %in% c("wilcoxon", "mannwhitney")) {
    return(sqrt(4 * log(sqrt(2) / bf01_threshold)))
  }
  if (test_key == "kruskal") {
    return(4 * log(4 / bf01_threshold))
  }

  tau2 <- dense_tau2(test_key, n, theta, data_theta)
  log_fun <- switch(
    test_key,
    ttest = function(x) log_bff10_t(x, n - 2, tau2),
    lm = function(x) log_bff10_t(x, n - 2, tau2),
    logistic = function(x) log_bff10_z(x, tau2),
    cox = function(x) log_bff10_z(sqrt(x), tau2),
    anova = function(x) log_bff10_anova(x, 4, n, tau2),
    ranova = function(x) log_bff10_ranova(x, 4, n, tau2),
    chisq = function(x) log_bff10_chisq(x, 4, tau2),
    stop("Unknown test key")
  )
  critical_root(log_fun)
}
```

```{r compute-dense-grid}
dense_cell <- function(test_key, n, theta) {
  q <- unname(q_by_test[test_key])
  n_eff_null <- n_eff_dense(test_key, n, 0)
  n_eff_alt <- n_eff_dense(test_key, n, theta)

  p_crit <- native_critical_from_p(test_key, n, p_threshold)
  p05_crit <- native_critical_from_p(test_key, n, p05_threshold)
  ejab_crit_null <- native_critical_from_p(
    test_key, n, p_cut_ejab(n_eff_null, q)
  )
  ejab_crit_alt <- native_critical_from_p(
    test_key, n, p_cut_ejab(n_eff_alt, q)
  )

  alpha_p <- native_rejection_at_stat(test_key, n, 0, p_crit)
  power_p <- native_rejection_at_stat(test_key, n, theta, p_crit)
  alpha_p05 <- native_rejection_at_stat(test_key, n, 0, p05_crit)
  power_p05 <- native_rejection_at_stat(test_key, n, theta, p05_crit)
  alpha_ejab <- native_rejection_at_stat(test_key, n, 0, ejab_crit_null)
  power_ejab <- native_rejection_at_stat(test_key, n, theta, ejab_crit_alt)

  if (test_key %in% tsbf_test_keys) {
    tsbf_crit_null <- dense_tsbf_critical(test_key, n, theta, data_theta = 0)
    tsbf_crit_alt <- dense_tsbf_critical(
      test_key, n, theta, data_theta = theta
    )
    alpha_tsbf <- native_rejection_at_stat(test_key, n, 0, tsbf_crit_null)
    power_tsbf <- native_rejection_at_stat(
      test_key, n, theta, tsbf_crit_alt
    )
  } else {
    alpha_tsbf <- power_tsbf <- NA_real_
  }

  # A fitted-model likelihood ratio is not available in this analytic dense
  # grid, so no BIC risk is calculated here.
  alpha_bic <- power_bic <- NA_real_

  risk_p <- 0.5 * (alpha_p + 1 - power_p)
  risk_p05 <- 0.5 * (alpha_p05 + 1 - power_p05)
  risk_ejab <- 0.5 * (alpha_ejab + 1 - power_ejab)
  risk_bic <- 0.5 * (alpha_bic + 1 - power_bic)
  risk_tsbf <- 0.5 * (alpha_tsbf + 1 - power_tsbf)

  tibble(
    test_key = test_key,
    test = unname(test_labels[test_key]),
    q = q,
    n = n,
    theta = theta,
    risk_p = risk_p,
    risk_p05 = risk_p05,
    risk_ejab = risk_ejab,
    risk_bic = risk_bic,
    risk_tsbf = risk_tsbf,
    delta_p_minus_ejab = risk_p - risk_ejab,
    delta_p05_minus_ejab = risk_p05 - risk_ejab,
    delta_bic_minus_ejab = risk_bic - risk_ejab,
    delta_tsbf_minus_ejab = risk_tsbf - risk_ejab
  )
}

if (isTRUE(params$dense_refresh) || !file.exists(dense_file)) {
  dense_tasks <- crossing(
    test_key = names(test_labels),
    n = dense_n_grid,
    theta = dense_theta_grid
  )
  task_groups <- split(dense_tasks, dense_tasks$test_key)
  run_dense_test <- function(task) {
    pmap_dfr(task, dense_cell)
  }
  if (.Platform$OS.type != "windows" && params$workers > 1) {
    dense_result <- parallel::mclapply(
      task_groups, run_dense_test, mc.cores = params$workers
    )
  } else {
    dense_result <- lapply(task_groups, run_dense_test)
  }
  dense_risk <- bind_rows(dense_result)
  saveRDS(dense_risk, dense_file)
} else {
  dense_risk <- readRDS(dense_file)
}

# Add the p<0.05 comparison to caches created before this comparison existed.
if (!all(c("risk_p05", "delta_p05_minus_ejab") %in% names(dense_risk))) {
  p05_risk <- pmap_dfr(
    dense_risk[c("test_key", "n", "theta")],
    function(test_key, n, theta) {
      p05_crit <- native_critical_from_p(test_key, n, p05_threshold)
      alpha_p05 <- native_rejection_at_stat(test_key, n, 0, p05_crit)
      power_p05 <- native_rejection_at_stat(test_key, n, theta, p05_crit)
      risk_p05 <- 0.5 * (alpha_p05 + 1 - power_p05)
      tibble(
        risk_p05 = risk_p05,
        delta_p05_minus_ejab = risk_p05
      )
    }
  )
  p05_risk$delta_p05_minus_ejab <-
    p05_risk$risk_p05 - dense_risk$risk_ejab
  dense_risk <- bind_cols(dense_risk, p05_risk)
  saveRDS(dense_risk, dense_file)
}

# Older caches contained BIC risks from a central/noncentral-chi-square
# approximation. Keep those values out of both the exported data and plots.
dense_risk <- dense_risk |>
  mutate(
    risk_bic = NA_real_,
    delta_bic_minus_ejab = NA_real_
  )

write.csv(dense_risk, dense_csv, row.names = FALSE)

kable(tibble(
  dense_cells = nrow(dense_risk),
  n_min = min(dense_risk$n),
  n_max = max(dense_risk$n),
  theta_min = min(dense_risk$theta),
  theta_max = max(dense_risk$theta),
  nonfinite_required_risks = sum(
    !is.finite(dense_risk$risk_p) |
      !is.finite(dense_risk$risk_p05) |
      !is.finite(dense_risk$risk_ejab)
  ),
  nonfinite_tsbf_risks_in_scope = sum(
    dense_risk$test_key %in% tsbf_test_keys &
      !is.finite(dense_risk$risk_tsbf)
  )
))
```

# Risk plots

Every risk map starts from the raw Bayes-risk difference

\[
\Delta R(n,\theta)
=R_{\mathrm{comparison}}(n,\theta)-R_{eJAB}(n,\theta).
\]

Within each test panel, the displayed value is

\[
\Delta R_{\mathrm{scaled}}(n,\theta)
=\frac{\Delta R(n,\theta)}{
\max_{n,\theta}|\Delta R(n,\theta)|}.
\]

All panels therefore use the same \([-1,1]\) purple-white-yellow scale without
clipping. Each panel is annotated with the raw maximum absolute risk difference
used as its divisor. Positive/yellow values mean eJAB has lower estimated Bayes
risk; negative/purple values mean the comparison rule has lower estimated
Bayes risk.

Dashed black curves show local alternatives of the form
\(\theta=c/\sqrt{n}\), or \(w=c/\sqrt{n}\) for the table-dimension plots, for
\(c\in\{0.25,0.5,0.75,1,1.5,2,3,4,6,8\}\).

```{r plotting-functions}
facet_labels <- c(
  ttest = "'t-test'~(q == 1)",
  lm = "'Linear regression'~(q == 1)",
  logistic = "'Logistic regression'~(q == 1)",
  cox = "'Cox PH'~(q == 1)",
  wilcoxon = "'Wilcoxon'~(q == 1)",
  mannwhitney = "'Mann-Whitney'~(q == 1)",
  anova = "'ANOVA'~(q == 3)",
  ranova = "'rANOVA'~(q == 3)",
  chisq = "'Chi-square'~(q == 4)",
  kruskal = "'Kruskal-Wallis'~(q == 4)"
)

local_alternative_c_values <- c(
  0.25, 0.5, 0.75, 1, 1.5, 2, 3, 4, 6, 8
)

local_alternative_contours <- function(n_values, effect_values) {
  crossing(
    c_value = local_alternative_c_values,
    log10_n = seq(
      log10(min(n_values)), log10(max(n_values)), length.out = 500
    )
  ) |>
    mutate(theta = c_value / sqrt(10^log10_n)) |>
    filter(
      theta >= min(effect_values),
      theta <= max(effect_values)
    )
}

plot_criticality_map <- function(data, value_col, title,
                                 filename, ncol = 5, height = 7.5,
                                 x_breaks = 2:7,
                                 x_labels = parse(text = paste0("10^", 2:7)),
                                 y_max = 2,
                                 y_breaks = seq(0.5, y_max, by = 0.5),
                                 y_label = expression(
                                   "Raw simulation effect size " * theta
                                 )) {
  scaled_data <- data |>
    filter(theta <= y_max + sqrt(.Machine$double.eps)) |>
    group_by(facet_label) |>
    mutate(
      panel_max_abs_risk_difference = max(
        abs(.data[[value_col]]), na.rm = TRUE
      ),
      panel_max_abs_risk_difference = if_else(
        is.finite(panel_max_abs_risk_difference) &
          panel_max_abs_risk_difference > 0,
        panel_max_abs_risk_difference,
        1
      ),
      scaled_risk_advantage =
        .data[[value_col]] / panel_max_abs_risk_difference
    ) |>
    ungroup()

  panel_annotations <- scaled_data |>
    distinct(facet_label, panel_max_abs_risk_difference) |>
    mutate(
      annotation = sprintf(
        "plain('max')*'|'*Delta*R*'|' == %.4f",
        panel_max_abs_risk_difference
      )
    )

  local_contours <- local_alternative_contours(
    10^data$log10_n,
    data$theta
  )

  p <- ggplot(
    scaled_data,
    aes(
      xmin = xmin, xmax = xmax,
      ymin = ymin, ymax = ymax,
      fill = scaled_risk_advantage
    )
  ) +
    geom_rect() +
    geom_line(
      data = local_contours,
      aes(x = log10_n, y = theta, group = c_value),
      inherit.aes = FALSE,
      color = "black",
      linetype = "dashed",
      linewidth = 0.3,
      alpha = 0.75
    ) +
    geom_label(
      data = panel_annotations,
      aes(x = -Inf, y = Inf, label = annotation),
      inherit.aes = FALSE,
      hjust = -0.05,
      vjust = 1.15,
      size = 3,
      linewidth = 0.2,
      parse = TRUE,
      fill = scales::alpha("white", 0.85)
    ) +
    facet_wrap(
      ~ facet_label,
      ncol = ncol,
      labeller = label_parsed
    ) +
    scale_x_continuous(
      breaks = x_breaks,
      labels = x_labels,
      expand = c(0, 0)
    ) +
    scale_y_continuous(
      breaks = y_breaks,
      limits = c(0, y_max),
      expand = c(0, 0)
    ) +
    scale_fill_gradient2(
      low = "#440154",
      mid = "#F7F7F7",
      high = "#FDE725",
      midpoint = 0,
      limits = c(-1, 1),
      breaks = c(-1, -0.5, 0, 0.5, 1),
      name = expression(frac(Delta*R, plain("max")*abs(Delta*R)))
    ) +
    labs(
      title = title,
      x = expression("Nominal sample size " * n),
      y = y_label
    ) +
    guides(
      fill = guide_colorbar(
        title.position = "right",
        barwidth = grid::unit(0.55, "cm"),
        barheight = grid::unit(8.2, "cm")
      )
    ) +
    theme_bw(base_size = 11) +
    theme(
      panel.grid = element_blank(),
      panel.border = element_rect(color = "black", linewidth = 0.45),
      legend.position = "right",
      legend.title = element_text(angle = 90, hjust = 0.5),
      strip.background = element_blank(),
      strip.text = element_text(size = 11),
      plot.title = element_text(hjust = 0.5, size = 17),
      axis.text.x = element_text(size = 9),
      panel.spacing = grid::unit(0.55, "cm")
    )

  ggsave(
    file.path(output_dir, filename),
    p,
    width = 16,
    height = height,
    dpi = 300
  )
  p
}

cell_boundaries <- function(values, lower = -Inf, upper = Inf) {
  centers <- sort(unique(values))
  middle <- (head(centers, -1) + tail(centers, -1)) / 2
  left <- c(centers[1] - (middle[1] - centers[1]), middle)
  right <- c(middle, centers[length(centers)] +
               (centers[length(centers)] - middle[length(middle)]))
  tibble(
    center = centers,
    lower = pmax(left, lower),
    upper = pmin(right, upper)
  )
}

x_bounds <- cell_boundaries(log10(dense_n_grid))
y_bounds <- cell_boundaries(dense_theta_grid, lower = 0, upper = 2)
empirical_x_bounds <- cell_boundaries(log10(bic_n_grid))
empirical_y_bounds <- cell_boundaries(
  bic_theta_grid, lower = 0, upper = 2
)

plot_data <- dense_risk |>
  mutate(
    log10_n = log10(n),
    test = factor(test, levels = unname(test_labels)),
    facet_label = factor(
      unname(facet_labels[test_key]),
      levels = unname(facet_labels)
    )
  ) |>
  left_join(
    x_bounds |> transmute(log10_n = center, xmin = lower, xmax = upper),
    by = "log10_n"
  ) |>
  left_join(
    y_bounds |> transmute(theta = center, ymin = lower, ymax = upper),
    by = "theta"
  )

# BIC risks are displayed only from the raw-data Monte Carlo experiment.
# In particular, they do not use the dense central/noncentral-chi-square
# rejection-probability approximation used to populate `dense_risk`.
empirical_bic_plot_data <- bic_risk |>
  mutate(
    log10_n = log10(n),
    test = factor(test, levels = unname(test_labels)),
    facet_label = factor(
      unname(facet_labels[test_key]),
      levels = unname(facet_labels[bic_test_keys])
    )
  ) |>
  left_join(
    empirical_x_bounds |>
      transmute(log10_n = center, xmin = lower, xmax = upper),
    by = "log10_n"
  ) |>
  left_join(
    empirical_y_bounds |>
      transmute(theta = center, ymin = lower, ymax = upper),
    by = "theta"
  ) |>
  droplevels()

chisq_dimensions <- tribble(
  ~rows, ~columns, ~table_label, ~table_plot_label, ~q,
  2L, 2L, "2 x 2 (q=1)",  "2 %*% 2~(q == 1)",   1L,
  3L, 3L, "3 x 3 (q=4)",  "3 %*% 3~(q == 4)",   4L,
  5L, 5L, "5 x 5 (q=16)", "5 %*% 5~(q == 16)", 16L
)

chisq_w_max <- 2
chisq_w_grid <- dense_theta_grid[
  dense_theta_grid <= chisq_w_max + sqrt(.Machine$double.eps)
]

estimated_omega_tsbf_criticals <- setNames(
  map_dbl(chisq_dimensions$q, estimated_omega_tsbf_critical),
  chisq_dimensions$q
)

chisq_dimension_cell <- function(
    rows, columns, table_label, table_plot_label, q, n, w) {
  lambda <- n * w^2

  critical_p05 <- qchisq(1 - p05_threshold, df = q)
  critical_ejab <- (
    log(n) - 2 * log(bf01_threshold)
  ) / (1 - n^(-1 / q))

  tau2 <- chisq_tau2(n, w^2, q)
  critical_tsbf <- critical_root(
    function(x) log_bff10_chisq(x, q, tau2)
  )
  critical_tsbf_estimated_omega <- unname(
    estimated_omega_tsbf_criticals[as.character(q)]
  )

  alpha_p05 <- pchisq(critical_p05, df = q, lower.tail = FALSE)
  alpha_ejab <- pchisq(critical_ejab, df = q, lower.tail = FALSE)
  alpha_tsbf <- if (is.finite(critical_tsbf)) {
    pchisq(critical_tsbf, df = q, lower.tail = FALSE)
  } else {
    0
  }
  alpha_tsbf_estimated_omega <- if (
      is.finite(critical_tsbf_estimated_omega)) {
    pchisq(
      critical_tsbf_estimated_omega,
      df = q,
      lower.tail = FALSE
    )
  } else {
    0
  }

  power_p05 <- pchisq(
    critical_p05, df = q, ncp = lambda, lower.tail = FALSE
  )
  power_ejab <- pchisq(
    critical_ejab, df = q, ncp = lambda, lower.tail = FALSE
  )
  power_tsbf <- if (is.finite(critical_tsbf)) {
    pchisq(
      critical_tsbf, df = q, ncp = lambda, lower.tail = FALSE
    )
  } else {
    0
  }
  power_tsbf_estimated_omega <- if (
      is.finite(critical_tsbf_estimated_omega)) {
    pchisq(
      critical_tsbf_estimated_omega,
      df = q,
      ncp = lambda,
      lower.tail = FALSE
    )
  } else {
    0
  }

  risk_p05 <- 0.5 * (alpha_p05 + 1 - power_p05)
  risk_ejab <- 0.5 * (alpha_ejab + 1 - power_ejab)
  risk_tsbf <- 0.5 * (alpha_tsbf + 1 - power_tsbf)
  risk_tsbf_estimated_omega <- 0.5 * (
    alpha_tsbf_estimated_omega + 1 - power_tsbf_estimated_omega
  )

  tibble(
    rows = rows,
    columns = columns,
    table_label = table_label,
    table_plot_label = table_plot_label,
    q = q,
    n = n,
    w = w,
    critical_tsbf_estimated_omega = critical_tsbf_estimated_omega,
    alpha_p05 = alpha_p05,
    alpha_ejab = alpha_ejab,
    alpha_tsbf = alpha_tsbf,
    alpha_tsbf_estimated_omega = alpha_tsbf_estimated_omega,
    power_p05 = power_p05,
    power_ejab = power_ejab,
    power_tsbf = power_tsbf,
    power_tsbf_estimated_omega = power_tsbf_estimated_omega,
    beta_p05 = 1 - power_p05,
    beta_ejab = 1 - power_ejab,
    beta_tsbf = 1 - power_tsbf,
    beta_tsbf_estimated_omega = 1 - power_tsbf_estimated_omega,
    risk_p05 = risk_p05,
    risk_ejab = risk_ejab,
    risk_tsbf = risk_tsbf,
    risk_tsbf_estimated_omega = risk_tsbf_estimated_omega,
    delta_p05_minus_ejab = risk_p05 - risk_ejab,
    delta_tsbf_minus_ejab = risk_tsbf - risk_ejab,
    delta_tsbf_estimated_omega_minus_ejab =
      risk_tsbf_estimated_omega - risk_ejab
  )
}

chisq_dimension_tasks <- crossing(
  chisq_dimensions,
  n = dense_n_grid,
  w = chisq_w_grid
)

chisq_dimension_risk <- pmap_dfr(
  chisq_dimension_tasks,
  chisq_dimension_cell
)

tsbf_dimension_plot_data <- chisq_dimension_risk |>
  transmute(
    test_key = table_label,
    test = table_label,
    q = q,
    n = n,
    theta = w,
    delta_tsbf_minus_ejab = delta_tsbf_minus_ejab,
    log10_n = log10(n),
    facet_label = factor(
      table_plot_label,
      levels = chisq_dimensions$table_plot_label
    )
  ) |>
  left_join(
    x_bounds |> transmute(log10_n = center, xmin = lower, xmax = upper),
    by = "log10_n"
  ) |>
  left_join(
    y_bounds |> transmute(theta = center, ymin = lower, ymax = upper),
    by = "theta"
  )

tsbf_estimated_omega_dimension_plot_data <- chisq_dimension_risk |>
  transmute(
    test_key = table_label,
    test = table_label,
    q = q,
    n = n,
    theta = w,
    delta_tsbf_estimated_omega_minus_ejab =
      delta_tsbf_estimated_omega_minus_ejab,
    log10_n = log10(n),
    facet_label = factor(
      table_plot_label,
      levels = chisq_dimensions$table_plot_label
    )
  ) |>
  left_join(
    x_bounds |> transmute(log10_n = center, xmin = lower, xmax = upper),
    by = "log10_n"
  ) |>
  left_join(
    y_bounds |> transmute(theta = center, ymin = lower, ymax = upper),
    by = "theta"
  )

panel_scaling_summary <- function(data, value_col, comparison_label) {
  data |>
    group_by(test_key, test, q) |>
    summarise(
      panel_max_abs_risk_difference = max(
        abs(.data[[value_col]]), na.rm = TRUE
      ),
      .groups = "drop"
    ) |>
    mutate(comparison = comparison_label, .before = 1)
}

all_risk_panel_scaling_factors <- bind_rows(
  panel_scaling_summary(
    plot_data,
    "delta_p_minus_ejab",
    "p < 0.005 versus eJAB"
  ),
  panel_scaling_summary(
    plot_data,
    "delta_p05_minus_ejab",
    "p < 0.05 versus eJAB"
  ),
  panel_scaling_summary(
    empirical_bic_plot_data,
    "delta_bic_minus_ejab",
    "BIC BF01 <= 1 versus eJAB"
  ),
  panel_scaling_summary(
    tsbf_dimension_plot_data,
    "delta_tsbf_minus_ejab",
    "TSBF01 <= 1 versus eJAB: 2x2, 3x3, and 5x5"
  ),
  panel_scaling_summary(
    tsbf_estimated_omega_dimension_plot_data,
    "delta_tsbf_estimated_omega_minus_ejab",
    "TSBF01 with estimated omega <= 1 versus eJAB: 2x2, 3x3, and 5x5"
  )
)

write.csv(
  all_risk_panel_scaling_factors,
  file.path(output_dir, "all_risk_plot_panel_scaling_factors.csv"),
  row.names = FALSE
)
```

## 1. eJAB versus \(p<0.005\): all 10 tests

```{r plot-p, fig.height=7.5}
p_vs_ejab <- plot_criticality_map(
  plot_data,
  "delta_p_minus_ejab",
  expression(
    "Large-sample Bayes-risk criticality maps: " *
      p < 0.005 ~ "versus" ~ eJAB["01"] <= 1
  ),
  "01_ejab_vs_p005_all_10_tests.png",
  ncol = 5,
  height = 7.5
)
p_vs_ejab
```

## 2. eJAB versus \(p<0.05\): all 10 tests

```{r plot-p05, fig.height=7.5}
p05_vs_ejab <- plot_criticality_map(
  plot_data,
  "delta_p05_minus_ejab",
  expression(
    "Large-sample Bayes-risk criticality maps: " *
      p < 0.05 ~ "versus" ~ eJAB["01"] <= 1
  ),
  "02_ejab_vs_p05_all_10_tests.png",
  ncol = 5,
  height = 7.5
)
p05_vs_ejab
```

## 3. eJAB versus BIC BF: six parametric-likelihood tests

```{r plot-bic, fig.height=7.5}
bic_vs_ejab <- plot_criticality_map(
  empirical_bic_plot_data,
  "delta_bic_minus_ejab",
  expression(
    "Large-sample empirical Bayes-risk criticality maps: " *
      BF["01"]^plain(BIC) <= 1 ~ "versus" ~ eJAB["01"] <= 1
  ),
  "02_ejab_vs_bic_bf01_six_parametric_tests.png",
  ncol = 3,
  height = 7.5,
  x_breaks = 2:7,
  x_labels = parse(text = paste0("10^", 2:7)),
  y_max = 2,
  y_breaks = c(0.5, 1, 1.5, 2)
)

bic_vs_ejab
```

## 4. eJAB versus TSBF: three chi-square table dimensions

```{r plot-tsbf, fig.height=5.5}
tsbf_vs_ejab <- plot_criticality_map(
  tsbf_dimension_plot_data,
  "delta_tsbf_minus_ejab",
  expression(
    "Large-sample Bayes-risk criticality maps: " *
      TSBF["01"] <= 1 ~ "versus" ~ eJAB["01"] <= 1
  ),
  "03_ejab_vs_tsbf01_three_table_sizes.png",
  ncol = 3,
  height = 5.5,
  y_label = expression(
    "True standardized effect size (" * omega * ")"
  )
)
tsbf_vs_ejab
```

## 5. Sensitivity analysis: TSBF using estimated effect size

The primary TSBF map above evaluates the BFF at the population effect size
used to generate each risk cell. As a plug-in sensitivity analysis, this
version estimates Cohen's effect from each realized chi-square statistic:

\[
\widehat\omega=\sqrt{\frac{X^2}{n}},\qquad
\widehat\tau^2=\frac{n}{q}\widehat\omega^2=\frac{X^2}{q}.
\]

Consequently, the BFF depends on the observed \(X^2\) both directly and through
\(\widehat\tau^2\). With the threshold \(BF_{01}\leq1\), \(X^2=0\) is an exact
equality point with probability zero under the continuous chi-square
distribution. For positive \(X^2\), the BFF initially falls below the threshold
and then crosses it from below. The risk calculation uses that positive upper
crossing as the rejection boundary.

This plug-in calculation is useful for the requested sensitivity check, but it
is an empirical-Bayes-style use of the test statistic rather than integration
over uncertainty in \(\omega\).

```{r estimated-omega-tsbf-criticals}
estimated_omega_tsbf_critical_table <- chisq_dimensions |>
  transmute(
    table_label,
    q,
    positive_chisq_critical = unname(
      estimated_omega_tsbf_criticals[as.character(q)]
    ),
    alpha = pchisq(
      positive_chisq_critical,
      df = q,
      lower.tail = FALSE
    )
  )

write.csv(
  estimated_omega_tsbf_critical_table,
  file.path(output_dir, "estimated_omega_tsbf_critical_values.csv"),
  row.names = FALSE
)

kable(
  estimated_omega_tsbf_critical_table,
  digits = 6,
  caption = "Plug-in estimated-effect TSBF rejection boundaries"
)
```

```{r plot-tsbf-estimated-omega, fig.height=5.5}
tsbf_estimated_omega_vs_ejab <- plot_criticality_map(
  tsbf_estimated_omega_dimension_plot_data,
  "delta_tsbf_estimated_omega_minus_ejab",
  expression(
    "Large-sample Bayes-risk sensitivity map: " *
      TSBF["01"](hat(omega)) <= 1 ~ "versus" ~ eJAB["01"] <= 1
  ),
  "03b_ejab_vs_tsbf01_estimated_omega_three_table_sizes.png",
  ncol = 3,
  height = 5.5,
  y_label = expression(
    "Population effect size defining risk (" * omega * ")"
  )
)
tsbf_estimated_omega_vs_ejab
```

## Combined figure

```{r combined-figure, fig.height=30}
combined <- p_vs_ejab / p05_vs_ejab / bic_vs_ejab / tsbf_vs_ejab +
  plot_annotation(
    title = "Four Bayes-risk comparisons for eJAB",
    theme = theme(plot.title = element_text(face = "bold", size = 16))
  )
combined

ggsave(
  file.path(output_dir, "00_all_four_bayes_risk_comparisons.png"),
  combined,
  width = 17,
  height = 30,
  dpi = 300
)
```

# Focused chi-square comparison across table dimensions

This analysis directly addresses the comparison of
\(eJAB_{01}\leq1\) with:

1. \(p<0.05\), and
2. \(TSBF_{01}\leq1\), with the BFF evaluated at the true standardized
   effect size.

The columns correspond to \(2\times2\), \(3\times3\), and \(5\times5\)
independence tests, with \(q=(R-1)(C-1)\) degrees of freedom.
The common vertical axis is Cohen's \(w\), so the effect-size scale is
comparable across table dimensions. Under the noncentral chi-square
approximation,

\[
X^2\mid H_0\sim\chi_q^2,\qquad
X^2\mid w\sim\chi_q^2(nw^2).
\]

Following Eqs. (6.5) and (6.6) of `Bayes_factor_function (1).pdf`, the
chi-square BFF is

\[
BFF_{10}(X^2\mid\tau^2)
=(1+\tau^2)^{-q/2-1}
\left\{1+\frac{\tau^2X^2}{q(1+\tau^2)}\right\}
\exp\left\{\frac{\tau^2X^2}{2(1+\tau^2)}\right\},
\qquad
\tau_\omega^2=\frac{n}{q}\omega^2.
\]

The BFF is evaluated at the true standardized effect, so the implementation
sets \(\tau^2=(n/q)w^2\). The risk calculations use equal prior model
probabilities and equal error costs:

\[
R_d(n,w)=\frac12\{\alpha_d(n,w)+1-\operatorname{power}_d(n,w)\}.
\]

The contingency-table generator below implements the checkerboard probability
perturbation used for these comparisons. For an \(R\times C\) table, there are
\(4\lfloor R/2\rfloor\lfloor C/2\rfloor\) perturbed cells. With

\[
\delta=
\frac{0.5w}{
\sqrt{RC\lfloor R/2\rfloor\lfloor C/2\rfloor}},
\]

the resulting population Cohen's effect satisfies

\[
\sum_{i,j}\frac{(p_{ij}-1/(RC))^2}{1/(RC)}=w^2.
\]

Thus, the noncentral chi-square risk engine below is the large-sample
counterpart of this exact data-generating process. This probability construction
requires

\[
w\leq 2\sqrt{
\frac{\lfloor R/2\rfloor\lfloor C/2\rfloor}{RC}
}.
\]

The maxima are \(1\), \(2/3\), and \(0.8\) for the \(2\times2\),
\(3\times3\), and \(5\times5\) designs, respectively. These bounds apply only
to this particular checkerboard
construction. The noncentral chi-square risk calculation depends on
\((q,n,w)\), rather than on that construction, so the plots use the broader
common range \(0\leq w\leq2\). Values above a design's attainable maximum
should therefore be interpreted as theoretical noncentral chi-square
extrapolations rather than results generated by a contingency table from that
design.

```{r chi-square-data-generator}
chisq_probabilities_from_omega <- function(R, C, omega = 0) {
  stopifnot(
    length(R) == 1, length(C) == 1, length(omega) == 1,
    is.finite(R), is.finite(C), is.finite(omega),
    R >= 2, C >= 2, R == as.integer(R), C == as.integer(C),
    omega >= 0
  )

  K <- R * C
  active_blocks <- floor(R / 2) * floor(C / 2)
  delta <- 0.5 * omega / sqrt(K * active_blocks)

  perturbation <- outer(
    seq_len(R),
    seq_len(C),
    function(i, j) (-1)^(i + j)
  )
  if (R %% 2 == 1) perturbation[R, ] <- 0
  if (C %% 2 == 1) perturbation[, C] <- 0

  probabilities <- matrix(1 / K, nrow = R, ncol = C) +
    delta * perturbation

  probability_tolerance <- sqrt(.Machine$double.eps)
  if (
    any(probabilities < -probability_tolerance) ||
      any(probabilities > 1 + probability_tolerance)
  ) {
    stop("omega produces invalid cell probabilities.")
  }
  pmin(pmax(probabilities, 0), 1)
}

simData6 <- function(
    size, R, C, omega = 0,
    simType = c("jointMulti", "indepMulti")) {
  simType <- match.arg(simType)
  if (
    length(size) != 1 || !is.finite(size) ||
      size < 1 || size != as.integer(size)
  ) {
    stop("size must be a positive integer.")
  }

  prob_H1 <- chisq_probabilities_from_omega(R, C, omega)

  if (simType == "jointMulti") {
    draw <- rmultinom(1, size = size, prob = as.vector(prob_H1))
    matrix(draw, nrow = R, ncol = C)
  } else {
    if (size %% R != 0) {
      stop("size must be divisible by R for balanced fixed row margins.")
    }
    row_size <- size / R
    t(vapply(
      seq_len(R),
      function(i) {
        as.vector(rmultinom(1, size = row_size, prob = prob_H1[i, ]))
      },
      numeric(C)
    ))
  }
}

chisq_design_validation <- chisq_dimensions |>
  select(rows, columns, table_label, q) |>
  crossing(omega = c(0.1, 0.3, 0.5, 0.65)) |>
  mutate(
    probabilities = pmap(
      list(rows, columns, omega),
      chisq_probabilities_from_omega
    ),
    recovered_omega = map_dbl(
      probabilities,
      function(p) {
        p0 <- 1 / length(p)
        sqrt(sum((p - p0)^2 / p0))
      }
    ),
    maximum_row_margin_error = map_dbl(
      probabilities,
      ~ max(abs(rowSums(.x) - 1 / nrow(.x)))
    ),
    maximum_column_margin_error = map_dbl(
      probabilities,
      ~ max(abs(colSums(.x) - 1 / ncol(.x)))
    )
  ) |>
  select(-probabilities)

stopifnot(
  max(abs(
    chisq_design_validation$recovered_omega -
      chisq_design_validation$omega
  )) < 1e-12,
  max(chisq_design_validation$maximum_row_margin_error) < 1e-12,
  max(chisq_design_validation$maximum_column_margin_error) < 1e-12,
  sum(simData6(120, 3, 3, 0.3, "jointMulti")) == 120,
  all(rowSums(
    simData6(120, 3, 3, 0.3, "indepMulti")
  ) == 40)
)

kable(
  chisq_design_validation,
  digits = 12,
  caption = "Validation of the contingency-table effect-size construction"
)
```

```{r chi-square-dimension-risk-grid}
chisq_dimension_risk_file <- file.path(
  output_dir, "chisq_dimensions_risk_grid.csv"
)
write.csv(
  chisq_dimension_risk,
  chisq_dimension_risk_file,
  row.names = FALSE
)

chisq_pattern_agreement <- chisq_dimension_risk |>
  group_by(rows, columns, table_label, q) |>
  summarise(
    grid_cells = n(),
    same_lower_risk_rule = mean(
      sign(delta_p05_minus_ejab) == sign(delta_tsbf_minus_ejab)
    ),
    both_favor_ejab = mean(
      delta_p05_minus_ejab > 0 & delta_tsbf_minus_ejab > 0
    ),
    both_favor_reference = mean(
      delta_p05_minus_ejab < 0 & delta_tsbf_minus_ejab < 0
    ),
    p05_only_favors_ejab = mean(
      delta_p05_minus_ejab > 0 & delta_tsbf_minus_ejab < 0
    ),
    tsbf_only_favors_ejab = mean(
      delta_p05_minus_ejab < 0 & delta_tsbf_minus_ejab > 0
    ),
    .groups = "drop"
  )

write.csv(
  chisq_pattern_agreement,
  file.path(output_dir, "chisq_dimensions_pattern_agreement.csv"),
  row.names = FALSE
)

kable(
  chisq_pattern_agreement,
  digits = 3,
  caption = "Agreement between the two eJAB risk-comparison patterns"
)
```

Each panel below starts with the raw Bayes-risk difference
\(\Delta R=R_{\mathrm{reference}}-R_{eJAB}\), then divides by that panel's
largest absolute raw difference:

\[
\Delta R_{\mathrm{scaled}}
=\frac{\Delta R}{\max_{n,w}|\Delta R|}.
\]

All panels therefore span the common scale \([-1,1]\), while the annotation in
each panel reports the raw \(\max|\Delta R|\) used as its divisor.
Positive/yellow values favor eJAB; negative/purple values favor the rule named
in that row. No risk differences are clipped.

```{r chi-square-dimension-risk-plot, fig.height=8}
chisq_x_bounds <- cell_boundaries(log10(dense_n_grid))
chisq_w_bounds <- cell_boundaries(
  chisq_w_grid, lower = 0, upper = chisq_w_max
)

chisq_dimension_plot_data <- chisq_dimension_risk |>
  select(
    rows, columns, table_label, table_plot_label, q, n, w,
    delta_p05_minus_ejab, delta_tsbf_minus_ejab
  ) |>
  pivot_longer(
    cols = starts_with("delta_"),
    names_to = "comparison",
    values_to = "risk_advantage"
  ) |>
  mutate(
    log10_n = log10(n),
    table_label = factor(
      table_label,
      levels = chisq_dimensions$table_label
    ),
    comparison = factor(
      comparison,
      levels = c(
        "delta_p05_minus_ejab",
        "delta_tsbf_minus_ejab"
      ),
      labels = c(
        "p < 0.05 versus eJAB",
        "TSBF(true w) versus eJAB"
      )
    ),
    comparison_plot_label = factor(
      comparison,
      levels = c(
        "p < 0.05 versus eJAB",
        "TSBF(true w) versus eJAB"
      ),
      labels = c(
        "italic(p) < 0.05~'versus'~eJAB['01'] <= 1",
        "TSBF['01'](omega) <= 1~'versus'~eJAB['01'] <= 1"
      )
    ),
    table_plot_label = factor(
      table_plot_label,
      levels = chisq_dimensions$table_plot_label
    )
  ) |>
  left_join(
    chisq_x_bounds |>
      transmute(log10_n = center, xmin = lower, xmax = upper),
    by = "log10_n"
  ) |>
  left_join(
    chisq_w_bounds |>
      transmute(w = center, ymin = lower, ymax = upper),
    by = "w"
  ) |>
  group_by(comparison, table_label) |>
  mutate(
    panel_max_abs_risk_difference = max(
      abs(risk_advantage), na.rm = TRUE
    ),
    scaled_risk_advantage =
      risk_advantage / panel_max_abs_risk_difference
  ) |>
  ungroup()

chisq_panel_scaling_factors <- chisq_dimension_plot_data |>
  distinct(
    comparison,
    table_label,
    panel_max_abs_risk_difference
  ) |>
  arrange(comparison, table_label)

write.csv(
  chisq_panel_scaling_factors,
  file.path(
    output_dir,
    "chisq_dimensions_panel_scaling_factors.csv"
  ),
  row.names = FALSE
)

kable(
  chisq_panel_scaling_factors,
  digits = 6,
  caption = "Panel-specific maximum absolute raw Bayes-risk differences"
)

chisq_panel_annotations <- chisq_panel_scaling_factors |>
  left_join(
    chisq_dimension_plot_data |>
      distinct(
        comparison, table_label,
        comparison_plot_label, table_plot_label
      ),
    by = c("comparison", "table_label")
  ) |>
  mutate(
    annotation = sprintf(
      "plain('max')*'|'*Delta*R*'|' == %.4f",
      panel_max_abs_risk_difference
    )
  )

chisq_local_contours <- local_alternative_contours(
  dense_n_grid,
  chisq_w_grid
)

chisq_dimension_risk_plot <- ggplot(
  chisq_dimension_plot_data,
  aes(
    xmin = xmin, xmax = xmax,
    ymin = ymin, ymax = ymax,
    fill = scaled_risk_advantage
  )
) +
  geom_rect() +
  geom_line(
    data = chisq_local_contours,
    aes(x = log10_n, y = theta, group = c_value),
    inherit.aes = FALSE,
    color = "black",
    linetype = "dashed",
    linewidth = 0.3,
    alpha = 0.75
  ) +
  geom_label(
    data = chisq_panel_annotations,
    aes(x = -Inf, y = Inf, label = annotation),
    inherit.aes = FALSE,
    hjust = -0.05,
    vjust = 1.15,
    size = 3,
    linewidth = 0.2,
    parse = TRUE,
    fill = scales::alpha("white", 0.85)
  ) +
  facet_grid(
    comparison_plot_label ~ table_plot_label,
    labeller = label_parsed
  ) +
  scale_x_continuous(
    breaks = 2:7,
    labels = parse(text = paste0("10^", 2:7)),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    breaks = seq(0.5, chisq_w_max, by = 0.5),
    limits = c(0, chisq_w_max),
    expand = c(0, 0)
  ) +
  scale_fill_gradient2(
    low = "#440154",
    mid = "#F7F7F7",
    high = "#FDE725",
    midpoint = 0,
    limits = c(-1, 1),
    breaks = c(-1, -0.5, 0, 0.5, 1),
    name = expression(frac(Delta*R, plain("max")*abs(Delta*R)))
  ) +
  labs(
    title = expression(chi^2~"Bayes-risk comparisons across table dimensions"),
    x = expression("Nominal sample size " * n),
    y = expression("True standardized effect size (" * omega * ")")
  ) +
  guides(
    fill = guide_colorbar(
      title.position = "right",
      barwidth = grid::unit(0.55, "cm"),
      barheight = grid::unit(8.2, "cm")
    )
  ) +
  theme_bw(base_size = 11) +
  theme(
    panel.grid = element_blank(),
    panel.border = element_rect(color = "black", linewidth = 0.45),
    legend.position = "right",
    legend.title = element_text(angle = 90, hjust = 0.5),
    strip.background = element_blank(),
    strip.text = element_text(size = 11),
    plot.title = element_text(hjust = 0.5, size = 16),
    axis.text.x = element_text(size = 9),
    panel.spacing = grid::unit(0.55, "cm")
  )

chisq_dimension_risk_plot

ggsave(
  file.path(
    output_dir,
    "04_chisq_dimensions_ejab_vs_p05_and_tsbf.png"
  ),
  chisq_dimension_risk_plot,
  width = 15,
  height = 8,
  dpi = 300
)
```

# Reproducibility

```{r reproducibility}
settings <- tibble(
  setting = c(
    "Seed",
    "Null replicates per test/sample-size cell",
    "Alternative replicates per effect-size cell",
    "BIC null replicates per test/sample-size cell",
    "BIC alternative replicates per effect-size cell",
    "Workers",
    "eJAB/BIC/TSBF BF01 threshold",
    "Classical p-value threshold (stringent)",
    "Classical p-value threshold (conventional)",
    "Simulation cache",
    "Dense empirical BIC cache"
  ),
  value = c(
    params$seed,
    params$null_reps,
    params$alt_reps,
    params$bic_null_reps,
    params$bic_alt_reps,
    params$workers,
    bf01_threshold,
    p_threshold,
    p05_threshold,
    simulation_file,
    bic_simulation_file
  )
)
kable(settings)
sessionInfo()
```
