---
title: "Bayes-risk comparisons for eJAB"
author: "eJAB clinical-trials project"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
params:
  refresh: false
  dense_refresh: false
  bic_refresh: false
  seed: 20260724
  null_reps: 2000
  alt_reps: 500
  bic_null_reps: 2000
  bic_alt_reps: 400
  workers: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 13,
  fig.height = 11,
  dpi = 160
)

library(dplyr)
library(ggplot2)
library(knitr)
library(patchwork)
library(purrr)
library(scales)
library(survival)
library(tidyr)

set.seed(params$seed)

output_dir <- "bayes_risk_outputs/revised_three_comparisons"
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
simulation_file <- file.path(output_dir, "bayes_risk_simulation.rds")
risk_file <- file.path(output_dir, "bayes_risk_grid.csv")
dense_file <- file.path(output_dir, "dense_bayes_risk_grid.rds")
dense_csv <- file.path(output_dir, "dense_bayes_risk_grid.csv")
bic_simulation_file <- file.path(
  output_dir, "large_sample_empirical_bic_simulation.rds"
)
bic_risk_file <- file.path(
  output_dir, "large_sample_empirical_bic_risk_grid.csv"
)

bf01_threshold <- 1 / 3
p_threshold <- 0.005
p05_threshold <- 0.05
n_grid <- c(30L, 60L, 120L, 250L, 500L)
theta_grid <- c(0.05, 0.10, 0.20, 0.35, 0.50, 0.80)
bic_n_grid <- unique(round(exp(seq(log(100), log(1e4), length.out = 24))))
bic_theta_grid <- unique(c(
  exp(seq(log(0.002), log(0.02), length.out = 10)),
  seq(0.025, 0.9, length.out = 22)
))
dense_n_grid <- unique(round(exp(seq(log(30), log(1e7), length.out = 70))))
dense_theta_grid <- unique(c(
  exp(seq(log(0.001), log(0.02), length.out = 14)),
  seq(0.025, 0.9, length.out = 46)
))
```

# Comparisons

This R Markdown file is the single source for four Bayes-risk comparisons:

1. \(eJAB_{01}\leq1/3\) versus \(p<0.005\), for all 10 tests.
2. \(eJAB_{01}\leq1/3\) versus \(p<0.05\), for all 10 tests.
3. \(eJAB_{01}\leq1/3\) versus \(BF_{01}^{BIC}\leq1/3\), for the seven
   non-rank tests. The Cox calculation uses the number of observed events as
   \(n_{\mathrm{eff}}\).
4. \(eJAB_{01}\leq1/3\) versus \(TSBF_{01}\leq1/3\), for all 10 tests.

For the seven non-rank tests, the TSBF functions are transcribed from
`Bayes_factor_function.pdf`. Each function is evaluated at the **true**
standardized effect size. This is preferable to an estimated effect in a
simulation study because the generating value is known and does not add
effect-estimation noise to the decision rule.

For Wilcoxon signed-rank and Mann-Whitney, the TSBF follows Theorem 1 and
Table 1 of Yuan and Johnson (2008). Setting \(\tau=1/C\) gives

\[
TSBF_{01}(T^*)=\sqrt{2}\exp(-T^{*2}/4).
\]

Kruskal-Wallis has a limiting chi-square rather than normal distribution, so
its TSBF uses the paper's Section 5 extension:

\[
TSBF_{01}(W)=(1+\tau)^{(k-1)/2}
 \exp\left\{-\frac{\tau W}{2(1+\tau)}\right\},
\qquad \tau=1.
\]

# Analysis scope and exclusions

```{r scope}
scope <- tribble(
  ~test, ~p_comparison, ~bic_comparison, ~tsbf_comparison, ~tsbf_source,
  "Two-sample t-test", "Included", "Included", "Included", "BFF PDF page 1",
  "Simple linear regression", "Included", "Included", "Included", "BFF PDF page 2",
  "Simple logistic regression", "Included", "Included", "Included", "BFF PDF page 3",
  "Cox proportional-hazards regression", "Included", "Included; events are n_eff", "Included", "BFF PDF page 7",
  "Wilcoxon signed-rank test", "Included", "Excluded", "Included", "Yuan-Johnson Theorem 1 and Table 1",
  "Mann-Whitney U-test", "Included", "Excluded", "Included", "Yuan-Johnson Theorem 1 and Table 1",
  "One-way ANOVA", "Included", "Included", "Included", "BFF PDF page 4",
  "Repeated-measures ANOVA", "Included", "Included", "Included", "BFF PDF page 5",
  "Chi-square independence test", "Included", "Included", "Included", "BFF PDF page 6",
  "Kruskal-Wallis test", "Included", "Excluded", "Included", "Yuan-Johnson Section 5"
)

kable(scope)
write.csv(
  scope,
  file.path(output_dir, "comparison_scope_and_exclusions.csv"),
  row.names = FALSE
)
```

The three rank tests are excluded only from the BIC comparison because that
comparison is restricted to the seven prespecified non-rank test classes. They
remain in the \(p<0.005\) and TSBF comparisons.

A simulated replicate is excluded before inspecting any decision-rule result
when the fitted model fails, its native test statistic is non-finite, its
\(p\)-value is non-finite or outside \([0,1]\), or its effective sample size is
not greater than one. No replicate is excluded because its \(p\)-value, test
statistic, effect size, eJAB, BIC BF, or TSBF is unusually large or small.

# eJAB and BIC functions

For eJAB, a test with \(q\) tested degrees of freedom maps its reported
two-sided or right-tail \(p\)-value to

\[
W=F^{-1}_{\chi_q^2}(1-p).
\]

The eJAB approximation is

\[
eJAB_{01}=\sqrt{n_{\mathrm{eff}}}
\exp\left[-\frac12(1-n_{\mathrm{eff}}^{-1/q})W\right]
\]

For the BIC comparison, let
\(\Lambda=2\{\ell(\widehat\theta_1)-\ell(\widehat\theta_0)\}\) be the
likelihood-ratio statistic calculated directly from the fitted alternative and
null models. The BIC Bayes-factor approximation is

\[
BF_{01}^{BIC}=n_{\mathrm{eff}}^{q/2}\exp(-\Lambda/2).
\]

Thus, the BIC decision does not substitute an asymptotic chi-square quantile
for the fitted-model likelihood ratio.

```{r ejab-bic-functions}
clip_p <- function(p) {
  pmin(pmax(p, .Machine$double.xmin), 1 - .Machine$double.eps)
}

ejab01 <- function(p, n_eff, q) {
  W <- qchisq(clip_p(p), df = q, lower.tail = FALSE)
  sqrt(n_eff) * exp(-0.5 * (1 - n_eff^(-1 / q)) * W)
}

bic_bf01_from_lr <- function(lr_statistic, n_eff, q) {
  n_eff^(q / 2) * exp(-pmax(lr_statistic, 0) / 2)
}
```

# Test-statistic Bayes-factor functions

The following functions transcribe the seven BFFs from
`Bayes_factor_function.pdf`. The PDF reports \(BFF_{10}\); the analysis uses
\(TSBF_{01}=1/BFF_{10}\).

```{r tsbf-functions}
bff10_t <- function(t, nu, tau2) {
  r <- 1 + t^2 / nu
  u <- 1 + t^2 / (nu * (1 + tau2))
  k <- tau2 * (nu + 1) / (nu * (1 + tau2))
  exp(
    -1.5 * log1p(tau2) +
      ((nu + 1) / 2) * (log(r) - log(u)) +
      log1p(k * t^2 / u)
  )
}

bff10_z <- function(z, tau2) {
  exp(
    -1.5 * log1p(tau2) +
      log1p(tau2 * z^2 / (1 + tau2)) +
      tau2 * z^2 / (2 * (1 + tau2))
  )
}

bff10_anova <- function(F_stat, groups, n_total, tau2) {
  v <- (n_total - groups) * (1 + tau2)
  exp(
    -((groups + 1) / 2) * log1p(tau2) +
      ((n_total - 1) / 2) *
        (log1p((groups - 1) * F_stat / (n_total - groups)) -
           log1p((groups - 1) * F_stat / v)) +
      log1p((n_total - 1) * tau2 * F_stat /
              (v + (groups - 1) * F_stat))
  )
}

bff10_ranova <- function(F_stat, conditions, subjects, tau2) {
  u <- (subjects - 1) * (1 + tau2)
  exp(
    -((conditions + 1) / 2) * log1p(tau2) +
      (((conditions - 1) * subjects) / 2) *
        (log1p(F_stat / (subjects - 1)) - log1p(F_stat / u)) +
      log1p(subjects * tau2 * F_stat / (u + F_stat))
  )
}

bff10_chisq <- function(chisq_stat, q, tau2) {
  exp(
    -(q / 2 + 1) * log1p(tau2) +
      log1p(tau2 * chisq_stat / (q * (1 + tau2))) +
      tau2 * chisq_stat / (2 * (1 + tau2))
  )
}

bff10_cox <- function(lambda, tau2) {
  bff10_z(sqrt(pmax(lambda, 0)), tau2)
}

yuan_johnson_normal_bf01 <- function(T_star) {
  sqrt(2) * exp(-T_star^2 / 4)
}

yuan_johnson_kw_bf01 <- function(W, groups, tau = 1) {
  q <- groups - 1
  (1 + tau)^(q / 2) * exp(-tau * W / (2 * (1 + tau)))
}

stopifnot(
  all.equal(bff10_t(2, 28, 0), 1),
  all.equal(bff10_z(2, 0), 1),
  all.equal(bff10_anova(2, 4, 40, 0), 1),
  all.equal(bff10_ranova(2, 4, 20, 0), 1),
  all.equal(bff10_chisq(5, 4, 0), 1),
  all.equal(bff10_cox(4, 0), 1),
  all.equal(yuan_johnson_normal_bf01(0), sqrt(2)),
  all.equal(yuan_johnson_kw_bf01(0, 5), 4)
)
```

# Simulation models

All tests use their native row-level statistic. The three rank tests use
heavy-tailed \(t_3\) errors with variance one, matching the earlier simulation.

```{r simulation-models}
test_labels <- c(
  ttest = "Two-sample t-test",
  lm = "Linear regression",
  logistic = "Logistic regression",
  cox = "Cox PH regression",
  wilcoxon = "Wilcoxon signed-rank",
  mannwhitney = "Mann-Whitney",
  anova = "One-way ANOVA",
  ranova = "Repeated-measures ANOVA",
  chisq = "Chi-square independence",
  kruskal = "Kruskal-Wallis"
)

q_by_test <- c(
  ttest = 1L, lm = 1L, logistic = 1L, cox = 1L,
  wilcoxon = 1L, mannwhitney = 1L,
  anova = 3L, ranova = 3L, chisq = 4L, kruskal = 4L
)

bic_test_keys <- c(
  "ttest", "lm", "logistic", "cox",
  "anova", "ranova", "chisq"
)

contrast_scores <- function(k) {
  x <- seq_len(k) - (k + 1) / 2
  x / sqrt(mean(x^2))
}

chisq_probabilities <- function(theta) {
  log_p <- matrix(0, 3, 3)
  diag(log_p) <- theta
  p <- exp(log_p)
  p / sum(p)
}

chisq_omega2 <- function(theta) {
  p <- chisq_probabilities(theta)
  expected <- outer(rowSums(p), colSums(p))
  sum((p - expected)^2 / expected)
}

tau2_for <- function(test_key, n, theta, simulation_result = NULL) {
  if (test_key == "ttest") {
    n1 <- floor(n / 2)
    n2 <- n - n1
    return(theta^2 / (2 * (1 / n1 + 1 / n2)))
  }
  if (test_key == "lm") return(n * theta^2 / 2)
  if (test_key == "logistic") return(n * (theta / 2)^2 / 2)
  if (test_key == "cox") {
    events <- simulation_result$events
    return(theta^2 * events / 4)
  }
  if (test_key == "anova") {
    groups <- 4
    scores <- contrast_scores(groups)
    sizes <- rep(floor(n / groups), groups)
    if (n %% groups > 0) {
      sizes[seq_len(n %% groups)] <- sizes[seq_len(n %% groups)] + 1
    }
    weights <- sizes / sum(sizes)
    means <- theta * scores
    omega2 <- sum(weights * (means - sum(weights * means))^2) / (groups - 1)
    return(n * omega2 / 2)
  }
  if (test_key == "ranova") {
    conditions <- 4
    means <- theta * contrast_scores(conditions)
    omega2 <- sum((means - mean(means))^2) / (conditions - 1)
    return(n * omega2 / 2)
  }
  if (test_key == "chisq") return(n * chisq_omega2(theta))
  NA_real_
}

simulate_one <- function(test_key, n, theta) {
  tryCatch({
    if (test_key == "ttest") {
      n1 <- floor(n / 2)
      n2 <- n - n1
      nu <- n - 2
      noncentrality <- theta / sqrt(1 / n1 + 1 / n2)
      statistic <- rt(1, nu, ncp = noncentrality)
      return(list(
        statistic = statistic,
        p = 2 * pt(-abs(statistic), nu),
        n_eff = n,
        nu = nu,
        bic_lr = n * log1p(statistic^2 / nu)
      ))
    }

    if (test_key == "lm") {
      nu <- n - 2
      centered_x_ss <- rchisq(1, n - 1)
      statistic <- rt(
        1, nu,
        ncp = theta * sqrt(centered_x_ss)
      )
      return(list(
        statistic = statistic,
        p = 2 * pt(-abs(statistic), nu),
        n_eff = n,
        nu = nu,
        bic_lr = n * log1p(statistic^2 / nu)
      ))
    }

    if (test_key == "logistic") {
      x <- rnorm(n)
      y <- rbinom(n, 1, plogis(theta * x))
      full_model <- suppressWarnings(glm(y ~ x, family = binomial()))
      null_model <- suppressWarnings(glm(y ~ 1, family = binomial()))
      statistic <- unname(summary(full_model)$coefficients["x", "z value"])
      return(list(
        statistic = statistic,
        p = 2 * pnorm(-abs(statistic)),
        n_eff = n,
        bic_lr = pmax(
          2 * (as.numeric(logLik(full_model)) -
                 as.numeric(logLik(null_model))),
          0
        )
      ))
    }

    if (test_key == "cox") {
      x <- sample(rep(c(0, 1), length.out = n))
      event_time <- rexp(n, rate = 0.1 * exp(theta * x))
      censor_time <- rexp(n, rate = 0.025)
      time <- pmin(event_time, censor_time)
      event <- as.integer(event_time <= censor_time)
      fit <- suppressWarnings(coxph(
        Surv(time, event) ~ x,
        ties = "efron",
        control = coxph.control(iter.max = 30)
      ))
      lambda <- pmax(unname(2 * diff(fit$loglik)), 0)
      events <- sum(event)
      return(list(
        statistic = lambda,
        p = pchisq(lambda, 1, lower.tail = FALSE),
        n_eff = events,
        events = events,
        bic_lr = lambda
      ))
    }

    if (test_key == "wilcoxon") {
      d <- theta + rt(n, 3) / sqrt(3)
      ranks <- rank(abs(d), ties.method = "average")
      T_stat <- sum(ranks[d > 0])
      mean0 <- n * (n + 1) / 4
      sd0 <- sqrt(n * (n + 1) * (2 * n + 1) / 24)
      T_star <- (T_stat - mean0) / sd0
      return(list(
        statistic = T_star,
        p = 2 * pnorm(-abs(T_star)),
        n_eff = n,
        bic_lr = NA_real_
      ))
    }

    if (test_key == "mannwhitney") {
      n1 <- floor(n / 2)
      n2 <- n - n1
      x <- rt(n1, 3) / sqrt(3)
      y <- theta + rt(n2, 3) / sqrt(3)
      combined_ranks <- rank(c(x, y), ties.method = "average")
      T_stat <- sum(combined_ranks[seq_len(n1)])
      mean0 <- n1 * (n + 1) / 2
      sd0 <- sqrt(n1 * n2 * (n + 1) / 12)
      T_star <- (T_stat - mean0) / sd0
      return(list(
        statistic = T_star,
        p = 2 * pnorm(-abs(T_star)),
        n_eff = n,
        bic_lr = NA_real_
      ))
    }

    if (test_key == "anova") {
      groups <- 4
      sizes <- rep(floor(n / groups), groups)
      if (n %% groups > 0) {
        sizes[seq_len(n %% groups)] <- sizes[seq_len(n %% groups)] + 1
      }
      means <- theta * contrast_scores(groups)
      weighted_mean <- sum(sizes * means) / n
      noncentrality <- sum(sizes * (means - weighted_mean)^2)
      df1 <- groups - 1
      df2 <- n - groups
      statistic <- rf(1, df1, df2, ncp = noncentrality)
      return(list(
        statistic = statistic,
        p = pf(statistic, df1, df2, lower.tail = FALSE),
        n_eff = n,
        bic_lr = n * log1p(df1 * statistic / df2)
      ))
    }

    if (test_key == "ranova") {
      conditions <- 4
      means <- theta * contrast_scores(conditions)
      df1 <- conditions - 1
      df2 <- (conditions - 1) * (n - 1)
      noncentrality <- n * sum((means - mean(means))^2)
      statistic <- rf(1, df1, df2, ncp = noncentrality)
      return(list(
        statistic = statistic,
        p = pf(statistic, df1, df2, lower.tail = FALSE),
        n_eff = n * (conditions - 1),
        bic_lr = n * conditions * log1p(df1 * statistic / df2)
      ))
    }

    if (test_key == "chisq") {
      counts <- matrix(
        rmultinom(1, n, as.vector(chisq_probabilities(theta))),
        3, 3
      )
      fit <- suppressWarnings(chisq.test(counts, correct = FALSE))
      statistic <- unname(fit$statistic)
      positive <- counts > 0 & fit$expected > 0
      likelihood_ratio <- 2 * sum(
        counts[positive] * log(counts[positive] / fit$expected[positive])
      )
      return(list(
        statistic = statistic,
        p = pchisq(statistic, 4, lower.tail = FALSE),
        n_eff = n,
        bic_lr = pmax(likelihood_ratio, 0)
      ))
    }

    if (test_key == "kruskal") {
      groups <- 5
      sizes <- rep(floor(n / groups), groups)
      if (n %% groups > 0) {
        sizes[seq_len(n %% groups)] <- sizes[seq_len(n %% groups)] + 1
      }
      means <- theta * contrast_scores(groups)
      group <- factor(rep(seq_len(groups), sizes))
      y <- rep(means, sizes) + rt(n, 3) / sqrt(3)
      fit <- kruskal.test(y ~ group)
      statistic <- unname(fit$statistic)
      return(list(
        statistic = statistic,
        p = pchisq(statistic, 4, lower.tail = FALSE),
        n_eff = n,
        bic_lr = NA_real_
      ))
    }

    stop("Unknown test key")
  }, error = function(e) {
    list(
      statistic = NA_real_,
      p = NA_real_,
      n_eff = NA_real_,
      bic_lr = NA_real_
    )
  })
}
```

# Decision rules and Bayes-risk estimation

```{r risk-functions}
valid_result <- function(x, test_key) {
  is.finite(x$statistic) &&
    is.finite(x$p) && x$p >= 0 && x$p <= 1 &&
    is.finite(x$n_eff) && x$n_eff > 1 &&
    (
      !test_key %in% bic_test_keys ||
        (is.finite(x$bic_lr) && x$bic_lr >= 0)
    )
}

tsbf01 <- function(test_key, result, n, theta) {
  if (test_key == "wilcoxon" || test_key == "mannwhitney") {
    return(yuan_johnson_normal_bf01(result$statistic))
  }
  if (test_key == "kruskal") {
    return(yuan_johnson_kw_bf01(result$statistic, groups = 5, tau = 1))
  }

  tau2 <- tau2_for(test_key, n, theta, result)
  bf10 <- switch(
    test_key,
    ttest = bff10_t(result$statistic, result$nu, tau2),
    lm = bff10_t(result$statistic, result$nu, tau2),
    logistic = bff10_z(result$statistic, tau2),
    cox = bff10_cox(result$statistic, tau2),
    anova = bff10_anova(result$statistic, 4, n, tau2),
    ranova = bff10_ranova(result$statistic, 4, n, tau2),
    chisq = bff10_chisq(result$statistic, 4, tau2),
    stop("Unknown test key")
  )
  1 / bf10
}

decisions_for <- function(draws, test_key, n, theta) {
  q <- unname(q_by_test[test_key])
  p <- map_dbl(draws, "p")
  n_eff <- map_dbl(draws, "n_eff")
  ejab <- ejab01(p, n_eff, q)
  bic <- if (test_key %in% bic_test_keys) {
    bic_bf01_from_lr(map_dbl(draws, "bic_lr"), n_eff, q)
  } else {
    rep(NA_real_, length(draws))
  }
  tsbf <- map_dbl(draws, ~ tsbf01(test_key, .x, n, theta))

  tibble(
    reject_p = p < p_threshold,
    reject_ejab = ejab <= bf01_threshold,
    reject_bic = bic <= bf01_threshold,
    reject_tsbf = tsbf <= bf01_threshold
  )
}

simulate_test_n <- function(test_key, n, null_reps, alt_reps, seed,
                            effect_grid = theta_grid) {
  set.seed(seed)

  null_draws <- replicate(
    null_reps,
    simulate_one(test_key, n, 0),
    simplify = FALSE
  )
  null_draws <- null_draws[
    map_lgl(null_draws, valid_result, test_key = test_key)
  ]

  map_dfr(effect_grid, function(theta) {
    alt_draws <- replicate(
      alt_reps,
      simulate_one(test_key, n, theta),
      simplify = FALSE
    )
    alt_draws <- alt_draws[
      map_lgl(alt_draws, valid_result, test_key = test_key)
    ]

    if (!length(null_draws) || !length(alt_draws)) {
      return(tibble(
        test_key = test_key,
        test = unname(test_labels[test_key]),
        n = n,
        theta = theta,
        null_valid_reps = length(null_draws),
        alt_valid_reps = length(alt_draws)
      ))
    }

    null_decisions <- decisions_for(null_draws, test_key, n, theta)
    alt_decisions <- decisions_for(alt_draws, test_key, n, theta)

    alpha_p <- mean(null_decisions$reject_p)
    power_p <- mean(alt_decisions$reject_p)
    alpha_ejab <- mean(null_decisions$reject_ejab)
    power_ejab <- mean(alt_decisions$reject_ejab)
    alpha_tsbf <- mean(null_decisions$reject_tsbf)
    power_tsbf <- mean(alt_decisions$reject_tsbf)

    if (test_key %in% bic_test_keys) {
      alpha_bic <- mean(null_decisions$reject_bic)
      power_bic <- mean(alt_decisions$reject_bic)
    } else {
      alpha_bic <- NA_real_
      power_bic <- NA_real_
    }

    risk_p <- 0.5 * (alpha_p + 1 - power_p)
    risk_ejab <- 0.5 * (alpha_ejab + 1 - power_ejab)
    risk_bic <- 0.5 * (alpha_bic + 1 - power_bic)
    risk_tsbf <- 0.5 * (alpha_tsbf + 1 - power_tsbf)

    tibble(
      test_key = test_key,
      test = unname(test_labels[test_key]),
      q = unname(q_by_test[test_key]),
      n = n,
      theta = theta,
      null_valid_reps = length(null_draws),
      alt_valid_reps = length(alt_draws),
      alpha_p = alpha_p,
      power_p = power_p,
      alpha_ejab = alpha_ejab,
      power_ejab = power_ejab,
      alpha_bic = alpha_bic,
      power_bic = power_bic,
      alpha_tsbf = alpha_tsbf,
      power_tsbf = power_tsbf,
      risk_p = risk_p,
      risk_ejab = risk_ejab,
      risk_bic = risk_bic,
      risk_tsbf = risk_tsbf,
      delta_p_minus_ejab = risk_p - risk_ejab,
      delta_bic_minus_ejab = risk_bic - risk_ejab,
      delta_tsbf_minus_ejab = risk_tsbf - risk_ejab
    )
  })
}
```

Null draws are generated once for each test and sample size and reused across
the six effect-size cells. The TSBF false-positive rate is nevertheless
recalculated for each effect size because the parametric BFF prior scale depends
on the standardized alternative effect.

```{r run-simulation}
task_grid <- crossing(
  test_key = names(test_labels),
  n = n_grid
) |>
  mutate(seed = params$seed + row_number())

if (isTRUE(params$refresh) || !file.exists(simulation_file)) {
  tasks <- split(task_grid, seq_len(nrow(task_grid)))
  run_task <- function(task) {
    simulate_test_n(
      task$test_key,
      task$n,
      params$null_reps,
      params$alt_reps,
      task$seed
    )
  }

  if (.Platform$OS.type != "windows" && params$workers > 1) {
    result_list <- parallel::mclapply(
      tasks,
      run_task,
      mc.cores = params$workers
    )
  } else {
    result_list <- lapply(tasks, run_task)
  }

  risk <- bind_rows(result_list)
  saveRDS(risk, simulation_file)
} else {
  risk <- readRDS(simulation_file)
}

write.csv(risk, risk_file, row.names = FALSE)

validation_summary <- risk |>
  summarise(
    cells = n(),
    minimum_null_valid_reps = min(null_valid_reps),
    minimum_alt_valid_reps = min(alt_valid_reps),
    nonfinite_required_risks = sum(
      !is.finite(risk_p) | !is.finite(risk_ejab) | !is.finite(risk_tsbf)
    ),
    nonfinite_bic_risks_in_scope = sum(
      test_key %in% bic_test_keys & !is.finite(risk_bic)
    )
  )

kable(validation_summary)
```

## Dense large-sample empirical BIC grid

The BIC comparison uses a dedicated \(24\times32\) grid: 24 logarithmically
spaced sample sizes from 100 to 10,000 and 32 effect sizes from 0.002 to 0.9.
Every cell is estimated from model-based Monte Carlo draws and direct
likelihood ratios. The Gaussian panels use exact finite-sample sufficient-
statistic distributions that are algebraically identical to refitting the
nested models. The logistic, Cox, and contingency-table panels use simulated
datasets and their fitted likelihood ratios. No interpolation or asymptotic
chi-square substitution is used to densify the BIC figure.

```{r dense-empirical-bic-grid}
bic_task_grid <- crossing(
  test_key = bic_test_keys,
  n = bic_n_grid
) |>
  mutate(seed = params$seed + 100000L + row_number())

if (isTRUE(params$bic_refresh) || !file.exists(bic_simulation_file)) {
  bic_tasks <- split(bic_task_grid, seq_len(nrow(bic_task_grid)))
  run_bic_task <- function(task) {
    simulate_test_n(
      task$test_key,
      task$n,
      params$bic_null_reps,
      params$bic_alt_reps,
      task$seed,
      effect_grid = bic_theta_grid
    )
  }

  if (.Platform$OS.type != "windows" && params$workers > 1) {
    bic_result_list <- parallel::mclapply(
      bic_tasks,
      run_bic_task,
      mc.cores = params$workers
    )
  } else {
    bic_result_list <- lapply(bic_tasks, run_bic_task)
  }

  bic_risk <- bind_rows(bic_result_list)
  saveRDS(bic_risk, bic_simulation_file)
} else {
  bic_risk <- readRDS(bic_simulation_file)
}

write.csv(bic_risk, bic_risk_file, row.names = FALSE)

bic_validation_summary <- bic_risk |>
  summarise(
    cells = n(),
    minimum_null_valid_reps = min(null_valid_reps),
    minimum_alt_valid_reps = min(alt_valid_reps),
    nonfinite_bic_risks = sum(!is.finite(risk_bic))
  )

kable(bic_validation_summary)
```

# Dense large-\(n\) criticality grid

The publication-style maps use a 70-point logarithmic sample-size grid from
\(30\) to \(10^7\) and a 60-point effect-size grid from \(0.001\) to \(0.9\).
At these sample sizes, row-level simulation is not computationally meaningful.
Instead, rejection probabilities are evaluated from each test's native
central/noncentral statistic distribution. The TSBF is always evaluated from
the relevant test-statistic BFF; no generic \(p\)-to-chi-square TSBF is used.

```{r dense-distributions}
t3_scale <- 1 / sqrt(3)
t3_pdf <- function(x) dt(x / t3_scale, df = 3) / t3_scale
t3_cdf <- function(x) pt(x / t3_scale, df = 3)

mw_ncp_per_n <- function(theta) {
  if (theta == 0) return(0)
  area <- integrate(
    function(e) t3_cdf(theta + e) * t3_pdf(e),
    -Inf, Inf, rel.tol = 1e-8, subdivisions = 300L
  )$value
  3 * (area - 0.5)^2
}

wilcoxon_ncp_per_n <- function(theta) {
  if (theta == 0) return(0)
  abs_cdf_shifted <- function(x) {
    t3_cdf(x - theta) - t3_cdf(-x - theta)
  }
  area <- integrate(
    function(d) abs_cdf_shifted(d) * t3_pdf(d - theta),
    0, Inf, rel.tol = 1e-8, subdivisions = 300L
  )$value
  12 * (area - 0.25)^2
}

kruskal_ncp_per_n <- function(theta) {
  if (theta == 0) return(0)
  locations <- theta * contrast_scores(5)
  pooled_cdf <- function(x) {
    rowMeans(vapply(locations, function(loc) t3_cdf(x - loc), numeric(length(x))))
  }
  rank_means <- vapply(locations, function(loc) {
    integrate(
      function(x) pooled_cdf(x) * t3_pdf(x - loc),
      -Inf, Inf, rel.tol = 1e-8, subdivisions = 300L
    )$value
  }, numeric(1))
  12 * mean((rank_means - 0.5)^2)
}

rank_ncp <- tibble(theta = c(0, dense_theta_grid)) |>
  mutate(
    wilcoxon = map_dbl(theta, wilcoxon_ncp_per_n),
    mannwhitney = map_dbl(theta, mw_ncp_per_n),
    kruskal = map_dbl(theta, kruskal_ncp_per_n)
  )

rank_ncp_value <- function(test_key, theta) {
  rank_ncp[[test_key]][match(theta, rank_ncp$theta)]
}

cox_event_fraction_dense <- function(theta) {
  0.5 * (
    0.1 / (0.1 + 0.025) +
      (0.1 * exp(theta)) / (0.1 * exp(theta) + 0.025)
  )
}

n_eff_dense <- function(test_key, n, theta) {
  if (test_key == "cox") return(n * cox_event_fraction_dense(theta))
  if (test_key == "ranova") return(3 * n)
  n
}

native_ncp <- function(test_key, n, theta) {
  if (theta == 0) return(0)
  switch(
    test_key,
    ttest = theta / sqrt(1 / floor(n / 2) + 1 / (n - floor(n / 2))),
    lm = sqrt(n) * theta,
    logistic = sqrt(n) * theta / 2,
    cox = n_eff_dense("cox", n, theta) * theta^2 / 4,
    wilcoxon = sqrt(n * rank_ncp_value("wilcoxon", theta)),
    mannwhitney = sqrt(n * rank_ncp_value("mannwhitney", theta)),
    anova = n * theta^2,
    ranova = 4 * n * theta^2,
    chisq = n * chisq_omega2(theta),
    kruskal = n * rank_ncp_value("kruskal", theta),
    stop("Unknown test key")
  )
}

native_rejection_at_stat <- function(test_key, n, theta, critical) {
  if (!is.finite(critical)) return(0)
  if (critical <= 0) return(1)
  ncp <- native_ncp(test_key, n, theta)

  probability <- switch(
    test_key,
    ttest = {
      df <- n - 2
      pt(-critical, df = df, ncp = ncp) +
        pt(critical, df = df, ncp = ncp, lower.tail = FALSE)
    },
    lm = {
      df <- n - 2
      pt(-critical, df = df, ncp = ncp) +
        pt(critical, df = df, ncp = ncp, lower.tail = FALSE)
    },
    logistic = {
      pnorm(-critical, mean = ncp) +
        pnorm(critical, mean = ncp, lower.tail = FALSE)
    },
    cox = pchisq(critical, df = 1, ncp = ncp, lower.tail = FALSE),
    wilcoxon = {
      pnorm(-critical, mean = ncp) +
        pnorm(critical, mean = ncp, lower.tail = FALSE)
    },
    mannwhitney = {
      pnorm(-critical, mean = ncp) +
        pnorm(critical, mean = ncp, lower.tail = FALSE)
    },
    anova = pf(
      critical, df1 = 3, df2 = n - 4, ncp = ncp, lower.tail = FALSE
    ),
    ranova = pf(
      critical, df1 = 3, df2 = 3 * (n - 1),
      ncp = ncp, lower.tail = FALSE
    ),
    chisq = pchisq(critical, df = 4, ncp = ncp, lower.tail = FALSE),
    kruskal = pchisq(critical, df = 4, ncp = ncp, lower.tail = FALSE),
    stop("Unknown test key")
  )
  pmin(pmax(probability, 0), 1)
}

native_critical_from_p <- function(test_key, n, p_cut) {
  p_cut <- clip_p(p_cut)
  switch(
    test_key,
    ttest = qt(1 - p_cut / 2, df = n - 2),
    lm = qt(1 - p_cut / 2, df = n - 2),
    logistic = qnorm(1 - p_cut / 2),
    cox = qchisq(1 - p_cut, df = 1),
    wilcoxon = qnorm(1 - p_cut / 2),
    mannwhitney = qnorm(1 - p_cut / 2),
    anova = qf(1 - p_cut, df1 = 3, df2 = n - 4),
    ranova = qf(1 - p_cut, df1 = 3, df2 = 3 * (n - 1)),
    chisq = qchisq(1 - p_cut, df = 4),
    kruskal = qchisq(1 - p_cut, df = 4),
    stop("Unknown test key")
  )
}

p_cut_ejab <- function(n_eff, q) {
  W <- (log(n_eff) - 2 * log(bf01_threshold)) /
    (1 - n_eff^(-1 / q))
  pchisq(W, df = q, lower.tail = FALSE)
}

```

```{r dense-tsbf-thresholds}
log_bff10_t <- function(t, nu, tau2) {
  r <- 1 + t^2 / nu
  u <- 1 + t^2 / (nu * (1 + tau2))
  k <- tau2 * (nu + 1) / (nu * (1 + tau2))
  -1.5 * log1p(tau2) +
    ((nu + 1) / 2) * (log(r) - log(u)) +
    log1p(k * t^2 / u)
}

log_bff10_z <- function(z, tau2) {
  -1.5 * log1p(tau2) +
    log1p(tau2 * z^2 / (1 + tau2)) +
    tau2 * z^2 / (2 * (1 + tau2))
}

log_bff10_anova <- function(F_stat, groups, n_total, tau2) {
  v <- (n_total - groups) * (1 + tau2)
  -((groups + 1) / 2) * log1p(tau2) +
    ((n_total - 1) / 2) *
      (log1p((groups - 1) * F_stat / (n_total - groups)) -
         log1p((groups - 1) * F_stat / v)) +
    log1p((n_total - 1) * tau2 * F_stat /
            (v + (groups - 1) * F_stat))
}

log_bff10_ranova <- function(F_stat, conditions, subjects, tau2) {
  u <- (subjects - 1) * (1 + tau2)
  -((conditions + 1) / 2) * log1p(tau2) +
    (((conditions - 1) * subjects) / 2) *
      (log1p(F_stat / (subjects - 1)) - log1p(F_stat / u)) +
    log1p(subjects * tau2 * F_stat / (u + F_stat))
}

log_bff10_chisq <- function(chisq_stat, q, tau2) {
  -(q / 2 + 1) * log1p(tau2) +
    log1p(tau2 * chisq_stat / (q * (1 + tau2))) +
    tau2 * chisq_stat / (2 * (1 + tau2))
}

critical_root <- function(log_bf_function) {
  target <- log(3)
  if (log_bf_function(0) >= target) return(0)
  upper <- 1
  value <- log_bf_function(upper)
  while ((!is.finite(value) || value < target) && upper < 1e7) {
    upper <- upper * 2
    value <- log_bf_function(upper)
  }
  if (!is.finite(value) || value < target) return(Inf)
  uniroot(
    function(x) log_bf_function(x) - target,
    interval = c(0, upper),
    tol = 1e-8
  )$root
}

dense_tau2 <- function(test_key, n, theta, data_theta = theta) {
  switch(
    test_key,
    ttest = {
      n1 <- floor(n / 2)
      n2 <- n - n1
      theta^2 / (2 * (1 / n1 + 1 / n2))
    },
    lm = n * theta^2 / 2,
    logistic = n * theta^2 / 8,
    cox = n_eff_dense("cox", n, data_theta) * theta^2 / 4,
    anova = n * theta^2 / 6,
    ranova = 2 * n * theta^2 / 3,
    chisq = n * chisq_omega2(theta),
    NA_real_
  )
}

dense_tsbf_critical <- function(test_key, n, theta, data_theta = theta) {
  if (test_key %in% c("wilcoxon", "mannwhitney")) {
    return(sqrt(4 * log(3 * sqrt(2))))
  }
  if (test_key == "kruskal") {
    return(4 * log(12))
  }

  tau2 <- dense_tau2(test_key, n, theta, data_theta)
  log_fun <- switch(
    test_key,
    ttest = function(x) log_bff10_t(x, n - 2, tau2),
    lm = function(x) log_bff10_t(x, n - 2, tau2),
    logistic = function(x) log_bff10_z(x, tau2),
    cox = function(x) log_bff10_z(sqrt(x), tau2),
    anova = function(x) log_bff10_anova(x, 4, n, tau2),
    ranova = function(x) log_bff10_ranova(x, 4, n, tau2),
    chisq = function(x) log_bff10_chisq(x, 4, tau2),
    stop("Unknown test key")
  )
  critical_root(log_fun)
}
```

```{r compute-dense-grid}
dense_cell <- function(test_key, n, theta) {
  q <- unname(q_by_test[test_key])
  n_eff_null <- n_eff_dense(test_key, n, 0)
  n_eff_alt <- n_eff_dense(test_key, n, theta)

  p_crit <- native_critical_from_p(test_key, n, p_threshold)
  p05_crit <- native_critical_from_p(test_key, n, p05_threshold)
  ejab_crit_null <- native_critical_from_p(
    test_key, n, p_cut_ejab(n_eff_null, q)
  )
  ejab_crit_alt <- native_critical_from_p(
    test_key, n, p_cut_ejab(n_eff_alt, q)
  )

  alpha_p <- native_rejection_at_stat(test_key, n, 0, p_crit)
  power_p <- native_rejection_at_stat(test_key, n, theta, p_crit)
  alpha_p05 <- native_rejection_at_stat(test_key, n, 0, p05_crit)
  power_p05 <- native_rejection_at_stat(test_key, n, theta, p05_crit)
  alpha_ejab <- native_rejection_at_stat(test_key, n, 0, ejab_crit_null)
  power_ejab <- native_rejection_at_stat(test_key, n, theta, ejab_crit_alt)

  tsbf_crit_null <- dense_tsbf_critical(test_key, n, theta, data_theta = 0)
  tsbf_crit_alt <- dense_tsbf_critical(
    test_key, n, theta, data_theta = theta
  )
  alpha_tsbf <- native_rejection_at_stat(test_key, n, 0, tsbf_crit_null)
  power_tsbf <- native_rejection_at_stat(
    test_key, n, theta, tsbf_crit_alt
  )

  # A fitted-model likelihood ratio is not available in this analytic dense
  # grid, so no BIC risk is calculated here.
  alpha_bic <- power_bic <- NA_real_

  risk_p <- 0.5 * (alpha_p + 1 - power_p)
  risk_p05 <- 0.5 * (alpha_p05 + 1 - power_p05)
  risk_ejab <- 0.5 * (alpha_ejab + 1 - power_ejab)
  risk_bic <- 0.5 * (alpha_bic + 1 - power_bic)
  risk_tsbf <- 0.5 * (alpha_tsbf + 1 - power_tsbf)

  tibble(
    test_key = test_key,
    test = unname(test_labels[test_key]),
    q = q,
    n = n,
    theta = theta,
    risk_p = risk_p,
    risk_p05 = risk_p05,
    risk_ejab = risk_ejab,
    risk_bic = risk_bic,
    risk_tsbf = risk_tsbf,
    delta_p_minus_ejab = risk_p - risk_ejab,
    delta_p05_minus_ejab = risk_p05 - risk_ejab,
    delta_bic_minus_ejab = risk_bic - risk_ejab,
    delta_tsbf_minus_ejab = risk_tsbf - risk_ejab
  )
}

if (isTRUE(params$dense_refresh) || !file.exists(dense_file)) {
  dense_tasks <- crossing(
    test_key = names(test_labels),
    n = dense_n_grid,
    theta = dense_theta_grid
  )
  task_groups <- split(dense_tasks, dense_tasks$test_key)
  run_dense_test <- function(task) {
    pmap_dfr(task, dense_cell)
  }
  if (.Platform$OS.type != "windows" && params$workers > 1) {
    dense_result <- parallel::mclapply(
      task_groups, run_dense_test, mc.cores = params$workers
    )
  } else {
    dense_result <- lapply(task_groups, run_dense_test)
  }
  dense_risk <- bind_rows(dense_result)
  saveRDS(dense_risk, dense_file)
} else {
  dense_risk <- readRDS(dense_file)
}

# Add the p<0.05 comparison to caches created before this comparison existed.
if (!all(c("risk_p05", "delta_p05_minus_ejab") %in% names(dense_risk))) {
  p05_risk <- pmap_dfr(
    dense_risk[c("test_key", "n", "theta")],
    function(test_key, n, theta) {
      p05_crit <- native_critical_from_p(test_key, n, p05_threshold)
      alpha_p05 <- native_rejection_at_stat(test_key, n, 0, p05_crit)
      power_p05 <- native_rejection_at_stat(test_key, n, theta, p05_crit)
      risk_p05 <- 0.5 * (alpha_p05 + 1 - power_p05)
      tibble(
        risk_p05 = risk_p05,
        delta_p05_minus_ejab = risk_p05
      )
    }
  )
  p05_risk$delta_p05_minus_ejab <-
    p05_risk$risk_p05 - dense_risk$risk_ejab
  dense_risk <- bind_cols(dense_risk, p05_risk)
  saveRDS(dense_risk, dense_file)
}

# Older caches contained BIC risks from a central/noncentral-chi-square
# approximation. Keep those values out of both the exported data and plots.
dense_risk <- dense_risk |>
  mutate(
    risk_bic = NA_real_,
    delta_bic_minus_ejab = NA_real_
  )

write.csv(dense_risk, dense_csv, row.names = FALSE)

kable(tibble(
  dense_cells = nrow(dense_risk),
  n_min = min(dense_risk$n),
  n_max = max(dense_risk$n),
  theta_min = min(dense_risk$theta),
  theta_max = max(dense_risk$theta),
  nonfinite_required_risks = sum(
    !is.finite(dense_risk$risk_p) |
      !is.finite(dense_risk$risk_p05) |
      !is.finite(dense_risk$risk_ejab) |
      !is.finite(dense_risk$risk_tsbf)
  )
))
```

# Risk plots

Every risk map starts from the raw Bayes-risk difference

\[
\Delta R(n,\theta)
=R_{\mathrm{comparison}}(n,\theta)-R_{eJAB}(n,\theta).
\]

Within each test panel, the displayed value is

\[
\Delta R_{\mathrm{scaled}}(n,\theta)
=\frac{\Delta R(n,\theta)}{
\max_{n,\theta}|\Delta R(n,\theta)|}.
\]

All panels therefore use the same \([-1,1]\) purple-white-yellow scale without
clipping. Each panel is annotated with the raw maximum absolute risk difference
used as its divisor. Positive/yellow values mean eJAB has lower estimated Bayes
risk; negative/purple values mean the comparison rule has lower estimated
Bayes risk.

```{r plotting-functions}
facet_labels <- c(
  ttest = "'t-test'~(q == 1)",
  lm = "'Linear regression'~(q == 1)",
  logistic = "'Logistic regression'~(q == 1)",
  cox = "'Cox PH'~(q == 1)",
  wilcoxon = "'Wilcoxon'~(q == 1)",
  mannwhitney = "'Mann-Whitney'~(q == 1)",
  anova = "'ANOVA'~(q == 3)",
  ranova = "'rANOVA'~(q == 3)",
  chisq = "'Chi-square'~(q == 4)",
  kruskal = "'Kruskal-Wallis'~(q == 4)"
)

plot_criticality_map <- function(data, value_col, title,
                                 filename, ncol = 5, height = 7.5,
                                 x_breaks = 2:7,
                                 x_labels = parse(text = paste0("10^", 2:7))) {
  scaled_data <- data |>
    group_by(facet_label) |>
    mutate(
      panel_max_abs_risk_difference = max(
        abs(.data[[value_col]]), na.rm = TRUE
      ),
      panel_max_abs_risk_difference = if_else(
        is.finite(panel_max_abs_risk_difference) &
          panel_max_abs_risk_difference > 0,
        panel_max_abs_risk_difference,
        1
      ),
      scaled_risk_advantage =
        .data[[value_col]] / panel_max_abs_risk_difference
    ) |>
    ungroup()

  panel_annotations <- scaled_data |>
    distinct(facet_label, panel_max_abs_risk_difference) |>
    mutate(
      annotation = sprintf(
        "plain('max')*'|'*Delta*R*'|' == %.4f",
        panel_max_abs_risk_difference
      )
    )

  p <- ggplot(
    scaled_data,
    aes(
      xmin = xmin, xmax = xmax,
      ymin = ymin, ymax = ymax,
      fill = scaled_risk_advantage
    )
  ) +
    geom_rect() +
    geom_label(
      data = panel_annotations,
      aes(x = -Inf, y = Inf, label = annotation),
      inherit.aes = FALSE,
      hjust = -0.05,
      vjust = 1.15,
      size = 3,
      linewidth = 0.2,
      parse = TRUE,
      fill = scales::alpha("white", 0.85)
    ) +
    facet_wrap(
      ~ facet_label,
      ncol = ncol,
      labeller = label_parsed
    ) +
    scale_x_continuous(
      breaks = x_breaks,
      labels = x_labels,
      expand = c(0, 0)
    ) +
    scale_y_continuous(
      breaks = c(0.2, 0.4, 0.6, 0.8),
      limits = c(0, 0.91),
      expand = c(0, 0)
    ) +
    scale_fill_gradient2(
      low = "#440154",
      mid = "#F7F7F7",
      high = "#FDE725",
      midpoint = 0,
      limits = c(-1, 1),
      breaks = c(-1, -0.5, 0, 0.5, 1),
      name = expression(frac(Delta*R, plain("max")*abs(Delta*R)))
    ) +
    labs(
      title = title,
      x = expression("Nominal sample size " * n),
      y = expression("Raw simulation effect size " * theta)
    ) +
    guides(
      fill = guide_colorbar(
        title.position = "right",
        barwidth = grid::unit(0.55, "cm"),
        barheight = grid::unit(8.2, "cm")
      )
    ) +
    theme_bw(base_size = 11) +
    theme(
      panel.grid = element_blank(),
      panel.border = element_rect(color = "black", linewidth = 0.45),
      legend.position = "right",
      legend.title = element_text(angle = 90, hjust = 0.5),
      strip.background = element_blank(),
      strip.text = element_text(size = 11),
      plot.title = element_text(hjust = 0.5, size = 17),
      axis.text.x = element_text(size = 9),
      panel.spacing = grid::unit(0.55, "cm")
    )

  ggsave(
    file.path(output_dir, filename),
    p,
    width = 16,
    height = height,
    dpi = 300
  )
  p
}

cell_boundaries <- function(values, lower = -Inf, upper = Inf) {
  centers <- sort(unique(values))
  middle <- (head(centers, -1) + tail(centers, -1)) / 2
  left <- c(centers[1] - (middle[1] - centers[1]), middle)
  right <- c(middle, centers[length(centers)] +
               (centers[length(centers)] - middle[length(middle)]))
  tibble(
    center = centers,
    lower = pmax(left, lower),
    upper = pmin(right, upper)
  )
}

x_bounds <- cell_boundaries(log10(dense_n_grid))
y_bounds <- cell_boundaries(dense_theta_grid, lower = 0, upper = 0.91)
empirical_x_bounds <- cell_boundaries(log10(bic_n_grid))
empirical_y_bounds <- cell_boundaries(
  bic_theta_grid, lower = 0, upper = 0.91
)

plot_data <- dense_risk |>
  mutate(
    log10_n = log10(n),
    test = factor(test, levels = unname(test_labels)),
    facet_label = factor(
      unname(facet_labels[test_key]),
      levels = unname(facet_labels)
    )
  ) |>
  left_join(
    x_bounds |> transmute(log10_n = center, xmin = lower, xmax = upper),
    by = "log10_n"
  ) |>
  left_join(
    y_bounds |> transmute(theta = center, ymin = lower, ymax = upper),
    by = "theta"
  )

# BIC risks are displayed only from the raw-data Monte Carlo experiment.
# In particular, they do not use the dense central/noncentral-chi-square
# rejection-probability approximation used to populate `dense_risk`.
empirical_bic_plot_data <- bic_risk |>
  mutate(
    log10_n = log10(n),
    test = factor(test, levels = unname(test_labels)),
    facet_label = factor(
      unname(facet_labels[test_key]),
      levels = unname(facet_labels[bic_test_keys])
    )
  ) |>
  left_join(
    empirical_x_bounds |>
      transmute(log10_n = center, xmin = lower, xmax = upper),
    by = "log10_n"
  ) |>
  left_join(
    empirical_y_bounds |>
      transmute(theta = center, ymin = lower, ymax = upper),
    by = "theta"
  ) |>
  droplevels()

panel_scaling_summary <- function(data, value_col, comparison_label) {
  data |>
    group_by(test_key, test, q) |>
    summarise(
      panel_max_abs_risk_difference = max(
        abs(.data[[value_col]]), na.rm = TRUE
      ),
      .groups = "drop"
    ) |>
    mutate(comparison = comparison_label, .before = 1)
}

all_risk_panel_scaling_factors <- bind_rows(
  panel_scaling_summary(
    plot_data,
    "delta_p_minus_ejab",
    "p < 0.005 versus eJAB"
  ),
  panel_scaling_summary(
    plot_data,
    "delta_p05_minus_ejab",
    "p < 0.05 versus eJAB"
  ),
  panel_scaling_summary(
    empirical_bic_plot_data,
    "delta_bic_minus_ejab",
    "BIC BF01 <= 1/3 versus eJAB"
  ),
  panel_scaling_summary(
    plot_data,
    "delta_tsbf_minus_ejab",
    "TSBF01 <= 1/3 versus eJAB"
  )
)

write.csv(
  all_risk_panel_scaling_factors,
  file.path(output_dir, "all_risk_plot_panel_scaling_factors.csv"),
  row.names = FALSE
)
```

## 1. eJAB versus \(p<0.005\): all 10 tests

```{r plot-p, fig.height=7.5}
p_vs_ejab <- plot_criticality_map(
  plot_data,
  "delta_p_minus_ejab",
  expression(
    "Large-sample Bayes-risk criticality maps: " *
      p < 0.005 ~ "versus" ~ eJAB["01"] <= 1/3
  ),
  "01_ejab_vs_p005_all_10_tests.png",
  ncol = 5,
  height = 7.5
)
p_vs_ejab
```

## 2. eJAB versus \(p<0.05\): all 10 tests

```{r plot-p05, fig.height=7.5}
p05_vs_ejab <- plot_criticality_map(
  plot_data,
  "delta_p05_minus_ejab",
  expression(
    "Large-sample Bayes-risk criticality maps: " *
      p < 0.05 ~ "versus" ~ eJAB["01"] <= 1/3
  ),
  "02_ejab_vs_p05_all_10_tests.png",
  ncol = 5,
  height = 7.5
)
p05_vs_ejab
```

## 3. eJAB versus BIC BF: seven non-rank tests, raw-data Monte Carlo

```{r plot-bic, fig.height=7.5}
bic_contours <- crossing(
  c_value = c(0.25, 0.5, 0.75, 1, 1.5, 2, 3, 4, 6, 8),
  log10_n = seq(2, 4, length.out = 500)
) |>
  mutate(theta = c_value / sqrt(10^log10_n)) |>
  filter(theta >= min(bic_theta_grid), theta <= max(bic_theta_grid))

bic_vs_ejab <- plot_criticality_map(
  empirical_bic_plot_data,
  "delta_bic_minus_ejab",
  expression(
    "Large-sample empirical Bayes-risk criticality maps: " *
      BF["01"]^plain(BIC) <= 1/3 ~ "versus" ~ eJAB["01"] <= 1/3
  ),
  "02_ejab_vs_bic_bf01_seven_tests.png",
  ncol = 4,
  height = 7.5,
  x_breaks = 2:4,
  x_labels = parse(text = paste0("10^", 2:4))
) +
  geom_line(
    data = bic_contours,
    aes(x = log10_n, y = theta, group = c_value),
    inherit.aes = FALSE,
    color = "black",
    linetype = "dashed",
    linewidth = 0.3,
    alpha = 0.75
  )

ggsave(
  file.path(output_dir, "02_ejab_vs_bic_bf01_seven_tests.png"),
  bic_vs_ejab,
  width = 16,
  height = 7.5,
  dpi = 300
)

bic_vs_ejab
```

## 4. eJAB versus TSBF: all 10 tests

```{r plot-tsbf, fig.height=7.5}
tsbf_vs_ejab <- plot_criticality_map(
  plot_data,
  "delta_tsbf_minus_ejab",
  expression(
    "Large-sample Bayes-risk criticality maps: " *
      TSBF["01"] <= 1/3 ~ "versus" ~ eJAB["01"] <= 1/3
  ),
  "03_ejab_vs_tsbf01_all_10_tests.png",
  ncol = 5,
  height = 7.5
)
tsbf_vs_ejab
```

## Combined figure

```{r combined-figure, fig.height=30}
combined <- p_vs_ejab / p05_vs_ejab / bic_vs_ejab / tsbf_vs_ejab +
  plot_annotation(
    title = "Four Bayes-risk comparisons for eJAB",
    theme = theme(plot.title = element_text(face = "bold", size = 16))
  )
combined

ggsave(
  file.path(output_dir, "00_all_four_bayes_risk_comparisons.png"),
  combined,
  width = 17,
  height = 30,
  dpi = 300
)
```

# Focused chi-square comparison across table dimensions

This analysis directly addresses the comparison of
\(eJAB_{01}\leq1/3\) with:

1. \(p<0.05\), and
2. \(TSBF_{01}\leq1/3\), with the BFF evaluated at the true standardized
   effect size.

The columns correspond to \(2\times2\), \(3\times3\), and \(5\times5\)
independence tests, with \(q=(d-1)^2\) degrees of freedom. The common vertical
axis is Cohen's \(w\), so the effect-size scale is comparable across table
dimensions. Under the noncentral chi-square approximation,

\[
X^2\mid H_0\sim\chi_q^2,\qquad
X^2\mid w\sim\chi_q^2(nw^2).
\]

Following Eqs. (6.5) and (6.6) of `Bayes_factor_function (1).pdf`, the
chi-square BFF is

\[
BFF_{10}(X^2\mid\tau^2)
=(1+\tau^2)^{-q/2-1}
\left\{1+\frac{\tau^2X^2}{q(1+\tau^2)}\right\}
\exp\left\{\frac{\tau^2X^2}{2(1+\tau^2)}\right\},
\qquad
\tau_\omega^2=n\omega^2.
\]

The BFF is evaluated at the true standardized effect, so the implementation
sets \(\tau^2=nw^2\). The risk calculations use equal prior model probabilities
and equal error costs:

\[
R_d(n,w)=\frac12\{\alpha_d(n,w)+1-\operatorname{power}_d(n,w)\}.
\]

The contingency-table generator below implements the checkerboard probability
perturbation used for these comparisons. For an \(R\times C\) table, there are
\(4\lfloor R/2\rfloor\lfloor C/2\rfloor\) perturbed cells. With

\[
\delta=
\frac{0.5w}{
\sqrt{RC\lfloor R/2\rfloor\lfloor C/2\rfloor}},
\]

the resulting population Cohen's effect satisfies

\[
\sum_{i,j}\frac{(p_{ij}-1/(RC))^2}{1/(RC)}=w^2.
\]

Thus, the noncentral chi-square risk engine below is the large-sample
counterpart of this exact data-generating process. This probability construction
requires

\[
w\leq 2\sqrt{
\frac{\lfloor R/2\rfloor\lfloor C/2\rfloor}{RC}
}.
\]

The maxima are \(1\), \(2/3\), and \(0.8\) for the \(2\times2\),
\(3\times3\), and \(5\times5\) designs, respectively. The plots therefore use
the common valid range \(0<w\leq2/3\).

```{r chi-square-data-generator}
chisq_probabilities_from_omega <- function(R, C, omega = 0) {
  stopifnot(
    length(R) == 1, length(C) == 1, length(omega) == 1,
    is.finite(R), is.finite(C), is.finite(omega),
    R >= 2, C >= 2, R == as.integer(R), C == as.integer(C),
    omega >= 0
  )

  K <- R * C
  active_blocks <- floor(R / 2) * floor(C / 2)
  delta <- 0.5 * omega / sqrt(K * active_blocks)

  perturbation <- outer(
    seq_len(R),
    seq_len(C),
    function(i, j) (-1)^(i + j)
  )
  if (R %% 2 == 1) perturbation[R, ] <- 0
  if (C %% 2 == 1) perturbation[, C] <- 0

  probabilities <- matrix(1 / K, nrow = R, ncol = C) +
    delta * perturbation

  probability_tolerance <- sqrt(.Machine$double.eps)
  if (
    any(probabilities < -probability_tolerance) ||
      any(probabilities > 1 + probability_tolerance)
  ) {
    stop("omega produces invalid cell probabilities.")
  }
  pmin(pmax(probabilities, 0), 1)
}

simData6 <- function(
    size, R, C, omega = 0,
    simType = c("jointMulti", "indepMulti")) {
  simType <- match.arg(simType)
  if (
    length(size) != 1 || !is.finite(size) ||
      size < 1 || size != as.integer(size)
  ) {
    stop("size must be a positive integer.")
  }

  prob_H1 <- chisq_probabilities_from_omega(R, C, omega)

  if (simType == "jointMulti") {
    draw <- rmultinom(1, size = size, prob = as.vector(prob_H1))
    matrix(draw, nrow = R, ncol = C)
  } else {
    if (size %% R != 0) {
      stop("size must be divisible by R for balanced fixed row margins.")
    }
    row_size <- size / R
    t(vapply(
      seq_len(R),
      function(i) {
        as.vector(rmultinom(1, size = row_size, prob = prob_H1[i, ]))
      },
      numeric(C)
    ))
  }
}

chisq_design_validation <- crossing(
  dimension = c(2L, 3L, 5L),
  omega = c(0.1, 0.3, 0.5, 0.65)
) |>
  mutate(
    probabilities = map2(
      dimension,
      omega,
      ~ chisq_probabilities_from_omega(.x, .x, .y)
    ),
    recovered_omega = map_dbl(
      probabilities,
      function(p) {
        p0 <- 1 / length(p)
        sqrt(sum((p - p0)^2 / p0))
      }
    ),
    maximum_row_margin_error = map_dbl(
      probabilities,
      ~ max(abs(rowSums(.x) - 1 / nrow(.x)))
    ),
    maximum_column_margin_error = map_dbl(
      probabilities,
      ~ max(abs(colSums(.x) - 1 / ncol(.x)))
    )
  ) |>
  select(-probabilities)

stopifnot(
  max(abs(
    chisq_design_validation$recovered_omega -
      chisq_design_validation$omega
  )) < 1e-12,
  max(chisq_design_validation$maximum_row_margin_error) < 1e-12,
  max(chisq_design_validation$maximum_column_margin_error) < 1e-12,
  sum(simData6(120, 3, 3, 0.3, "jointMulti")) == 120,
  all(rowSums(
    simData6(120, 3, 3, 0.3, "indepMulti")
  ) == 40)
)

kable(
  chisq_design_validation,
  digits = 12,
  caption = "Validation of the contingency-table effect-size construction"
)
```

```{r chi-square-dimension-risk-grid}
chisq_dimensions <- tibble(
  dimension = c(2L, 3L, 5L),
  table_label = c("2 x 2 (q=1)", "3 x 3 (q=4)", "5 x 5 (q=16)"),
  q = c(1L, 4L, 16L)
)

chisq_w_grid <- dense_theta_grid[
  dense_theta_grid <= 2 / 3 + sqrt(.Machine$double.eps)
]

chisq_dimension_cell <- function(dimension, table_label, q, n, w) {
  lambda <- n * w^2

  critical_p05 <- qchisq(1 - p05_threshold, df = q)
  critical_ejab <- (
    log(n) - 2 * log(bf01_threshold)
  ) / (1 - n^(-1 / q))

  tau2 <- n * w^2
  critical_tsbf <- critical_root(
    function(x) log_bff10_chisq(x, q, tau2)
  )

  alpha_p05 <- pchisq(critical_p05, df = q, lower.tail = FALSE)
  alpha_ejab <- pchisq(critical_ejab, df = q, lower.tail = FALSE)
  alpha_tsbf <- if (is.finite(critical_tsbf)) {
    pchisq(critical_tsbf, df = q, lower.tail = FALSE)
  } else {
    0
  }

  power_p05 <- pchisq(
    critical_p05, df = q, ncp = lambda, lower.tail = FALSE
  )
  power_ejab <- pchisq(
    critical_ejab, df = q, ncp = lambda, lower.tail = FALSE
  )
  power_tsbf <- if (is.finite(critical_tsbf)) {
    pchisq(
      critical_tsbf, df = q, ncp = lambda, lower.tail = FALSE
    )
  } else {
    0
  }

  risk_p05 <- 0.5 * (alpha_p05 + 1 - power_p05)
  risk_ejab <- 0.5 * (alpha_ejab + 1 - power_ejab)
  risk_tsbf <- 0.5 * (alpha_tsbf + 1 - power_tsbf)

  tibble(
    dimension = dimension,
    table_label = table_label,
    q = q,
    n = n,
    w = w,
    alpha_p05 = alpha_p05,
    alpha_ejab = alpha_ejab,
    alpha_tsbf = alpha_tsbf,
    power_p05 = power_p05,
    power_ejab = power_ejab,
    power_tsbf = power_tsbf,
    beta_p05 = 1 - power_p05,
    beta_ejab = 1 - power_ejab,
    beta_tsbf = 1 - power_tsbf,
    risk_p05 = risk_p05,
    risk_ejab = risk_ejab,
    risk_tsbf = risk_tsbf,
    delta_p05_minus_ejab = risk_p05 - risk_ejab,
    delta_tsbf_minus_ejab = risk_tsbf - risk_ejab
  )
}

chisq_dimension_tasks <- crossing(
  chisq_dimensions,
  n = dense_n_grid,
  w = chisq_w_grid
)

chisq_dimension_risk <- pmap_dfr(
  chisq_dimension_tasks,
  chisq_dimension_cell
)

chisq_dimension_risk_file <- file.path(
  output_dir, "chisq_dimensions_risk_grid.csv"
)
write.csv(
  chisq_dimension_risk,
  chisq_dimension_risk_file,
  row.names = FALSE
)

chisq_pattern_agreement <- chisq_dimension_risk |>
  group_by(dimension, table_label, q) |>
  summarise(
    grid_cells = n(),
    same_lower_risk_rule = mean(
      sign(delta_p05_minus_ejab) == sign(delta_tsbf_minus_ejab)
    ),
    both_favor_ejab = mean(
      delta_p05_minus_ejab > 0 & delta_tsbf_minus_ejab > 0
    ),
    both_favor_reference = mean(
      delta_p05_minus_ejab < 0 & delta_tsbf_minus_ejab < 0
    ),
    p05_only_favors_ejab = mean(
      delta_p05_minus_ejab > 0 & delta_tsbf_minus_ejab < 0
    ),
    tsbf_only_favors_ejab = mean(
      delta_p05_minus_ejab < 0 & delta_tsbf_minus_ejab > 0
    ),
    .groups = "drop"
  )

write.csv(
  chisq_pattern_agreement,
  file.path(output_dir, "chisq_dimensions_pattern_agreement.csv"),
  row.names = FALSE
)

kable(
  chisq_pattern_agreement,
  digits = 3,
  caption = "Agreement between the two eJAB risk-comparison patterns"
)
```

Each panel below starts with the raw Bayes-risk difference
\(\Delta R=R_{\mathrm{reference}}-R_{eJAB}\), then divides by that panel's
largest absolute raw difference:

\[
\Delta R_{\mathrm{scaled}}
=\frac{\Delta R}{\max_{n,w}|\Delta R|}.
\]

All panels therefore span the common scale \([-1,1]\), while the annotation in
each panel reports the raw \(\max|\Delta R|\) used as its divisor.
Positive/yellow values favor eJAB; negative/purple values favor the rule named
in that row. No risk differences are clipped.

```{r chi-square-dimension-risk-plot, fig.height=8}
chisq_x_bounds <- cell_boundaries(log10(dense_n_grid))
chisq_w_bounds <- cell_boundaries(
  chisq_w_grid, lower = 0, upper = 2 / 3
)

chisq_dimension_plot_data <- chisq_dimension_risk |>
  select(
    dimension, table_label, q, n, w,
    delta_p05_minus_ejab, delta_tsbf_minus_ejab
  ) |>
  pivot_longer(
    cols = starts_with("delta_"),
    names_to = "comparison",
    values_to = "risk_advantage"
  ) |>
  mutate(
    log10_n = log10(n),
    table_label = factor(
      table_label,
      levels = chisq_dimensions$table_label
    ),
    comparison = factor(
      comparison,
      levels = c(
        "delta_p05_minus_ejab",
        "delta_tsbf_minus_ejab"
      ),
      labels = c(
        "p < 0.05 versus eJAB",
        "TSBF(true w) versus eJAB"
      )
    ),
    comparison_plot_label = factor(
      comparison,
      levels = c(
        "p < 0.05 versus eJAB",
        "TSBF(true w) versus eJAB"
      ),
      labels = c(
        "italic(p) < 0.05~'versus'~eJAB['01'] <= 1/3",
        "TSBF['01'](omega) <= 1/3~'versus'~eJAB['01'] <= 1/3"
      )
    ),
    table_plot_label = factor(
      table_label,
      levels = chisq_dimensions$table_label,
      labels = c(
        "2 %*% 2~(q == 1)",
        "3 %*% 3~(q == 4)",
        "5 %*% 5~(q == 16)"
      )
    )
  ) |>
  left_join(
    chisq_x_bounds |>
      transmute(log10_n = center, xmin = lower, xmax = upper),
    by = "log10_n"
  ) |>
  left_join(
    chisq_w_bounds |>
      transmute(w = center, ymin = lower, ymax = upper),
    by = "w"
  ) |>
  group_by(comparison, table_label) |>
  mutate(
    panel_max_abs_risk_difference = max(
      abs(risk_advantage), na.rm = TRUE
    ),
    scaled_risk_advantage =
      risk_advantage / panel_max_abs_risk_difference
  ) |>
  ungroup()

chisq_panel_scaling_factors <- chisq_dimension_plot_data |>
  distinct(
    comparison,
    table_label,
    panel_max_abs_risk_difference
  ) |>
  arrange(comparison, table_label)

write.csv(
  chisq_panel_scaling_factors,
  file.path(
    output_dir,
    "chisq_dimensions_panel_scaling_factors.csv"
  ),
  row.names = FALSE
)

kable(
  chisq_panel_scaling_factors,
  digits = 6,
  caption = "Panel-specific maximum absolute raw Bayes-risk differences"
)

chisq_panel_annotations <- chisq_panel_scaling_factors |>
  left_join(
    chisq_dimension_plot_data |>
      distinct(
        comparison, table_label,
        comparison_plot_label, table_plot_label
      ),
    by = c("comparison", "table_label")
  ) |>
  mutate(
    annotation = sprintf(
      "plain('max')*'|'*Delta*R*'|' == %.4f",
      panel_max_abs_risk_difference
    )
  )

chisq_dimension_risk_plot <- ggplot(
  chisq_dimension_plot_data,
  aes(
    xmin = xmin, xmax = xmax,
    ymin = ymin, ymax = ymax,
    fill = scaled_risk_advantage
  )
) +
  geom_rect() +
  geom_label(
    data = chisq_panel_annotations,
    aes(x = -Inf, y = Inf, label = annotation),
    inherit.aes = FALSE,
    hjust = -0.05,
    vjust = 1.15,
    size = 3,
    linewidth = 0.2,
    parse = TRUE,
    fill = scales::alpha("white", 0.85)
  ) +
  facet_grid(
    comparison_plot_label ~ table_plot_label,
    labeller = label_parsed
  ) +
  scale_x_continuous(
    breaks = 2:7,
    labels = parse(text = paste0("10^", 2:7)),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    breaks = c(0.2, 0.4, 0.6),
    limits = c(0, 2 / 3),
    expand = c(0, 0)
  ) +
  scale_fill_gradient2(
    low = "#440154",
    mid = "#F7F7F7",
    high = "#FDE725",
    midpoint = 0,
    limits = c(-1, 1),
    breaks = c(-1, -0.5, 0, 0.5, 1),
    name = expression(frac(Delta*R, plain("max")*abs(Delta*R)))
  ) +
  labs(
    title = expression(chi^2~"Bayes-risk comparisons across table dimensions"),
    x = expression("Nominal sample size " * n),
    y = expression("True standardized effect size (" * omega * ")")
  ) +
  guides(
    fill = guide_colorbar(
      title.position = "right",
      barwidth = grid::unit(0.55, "cm"),
      barheight = grid::unit(8.2, "cm")
    )
  ) +
  theme_bw(base_size = 11) +
  theme(
    panel.grid = element_blank(),
    panel.border = element_rect(color = "black", linewidth = 0.45),
    legend.position = "right",
    legend.title = element_text(angle = 90, hjust = 0.5),
    strip.background = element_blank(),
    strip.text = element_text(size = 11),
    plot.title = element_text(hjust = 0.5, size = 16),
    axis.text.x = element_text(size = 9),
    panel.spacing = grid::unit(0.55, "cm")
  )

chisq_dimension_risk_plot

ggsave(
  file.path(
    output_dir,
    "04_chisq_dimensions_ejab_vs_p05_and_tsbf.png"
  ),
  chisq_dimension_risk_plot,
  width = 15,
  height = 8,
  dpi = 300
)
```

# Reproducibility

```{r reproducibility}
settings <- tibble(
  setting = c(
    "Seed",
    "Null replicates per test/sample-size cell",
    "Alternative replicates per effect-size cell",
    "BIC null replicates per test/sample-size cell",
    "BIC alternative replicates per effect-size cell",
    "Workers",
    "eJAB/BIC/TSBF BF01 threshold",
    "Classical p-value threshold (stringent)",
    "Classical p-value threshold (conventional)",
    "Simulation cache",
    "Dense empirical BIC cache"
  ),
  value = c(
    params$seed,
    params$null_reps,
    params$alt_reps,
    params$bic_null_reps,
    params$bic_alt_reps,
    params$workers,
    bf01_threshold,
    p_threshold,
    p05_threshold,
    simulation_file,
    bic_simulation_file
  )
)
kable(settings)
sessionInfo()
```
