---
title: "Generalized Jeffreys's Approximate Objective Bayes Factor"
subtitle: "Part Two: Simple Linear Regression Simulation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
runthis <- T # whether to run chunks 11-15 & 17-18

if(!require(BayesFactor)) {install.packages("BayesFactor")}
if(!require(rstan)) {install.packages("rstan")}
if(!require(polspline)) {install.packages("polspline")}
if(!require(ggplot2)) {install.packages("ggplot2"); library(ggplot2)}
if(!require(plotly)) {install.packages("plotly"); library(plotly)}
if(!require(remotes)) {install.packages("remotes")}
if(!require(klippy)) {remotes::install_github("rlesur/klippy")}
```
```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```
============================================================================================================ 

Open Science Framework: <https://osf.io/qeprj/>

&nbsp;

# {.tabset .tabset-fade .tabset-pills}
## 1. Methods {.tabset .tabset-fade}

### "*BayesFactor*" R Package

&nbsp;

\begin{align} \tag{1}
    \mathcal{M}_1:\ y_i&=\beta_0+\beta_1 x_i+\epsilon_i\\ \text{versus}\quad\mathcal{M}_0:\ y_i&=\beta_0+\epsilon_i,\\ \qquad\qquad&\epsilon_{i}\overset{\text{i.i.d.}}{\sim}\mathcal{N}(0,\ \sigma_\epsilon^2)\quad\text{ for } i=1,\dotsb,n
\end{align}

\begin{equation} \tag{2}
    \pi(\beta_0,\sigma_\epsilon^2)\propto 1/\ \sigma_\epsilon^{2}
\end{equation}

\begin{equation} \tag{3}
    \beta_1\sim\mathcal{N}\left(0,\ ng\cdot\sigma_\epsilon^2\cdot(\boldsymbol{x}^\top\boldsymbol{x})^{-1}\right)
\end{equation}

\begin{equation} \tag{4}
    g\sim\text{Inv-Gamma}(1/2,\ \color{lightgray}{\text{scale}=}\varphi/2)
\end{equation}

By default, the prior scale on the standardized slope is $\varphi=\sqrt{2}/4$.

The scales are medium, wide, and ultrawide for $\varphi\in\{\sqrt{2}/4,\ 1/2,\ \sqrt{2}/2\}$ in linear regression, $h\in\{1/2,\ \sqrt{2}/2,\ 1\}$ in analysis of variance, and $\gamma\in\{\sqrt{2}/2,\ 1,\ \sqrt{2}\}$ in [$t$-tests](https://rpubs.com/sherloconan/1203268), respectively.

See `?BayesFactor::regressionBF`.

&nbsp;

[Equivalently](https://en.wikipedia.org/wiki/Inverse-gamma_distribution#Derivation_from_Gamma_distribution), $g^{-1}\sim\text{Gamma}(1/2,\ \color{lightgray}{\text{rate}=}\varphi/2)$.&nbsp;&nbsp; [Or](https://en.wikipedia.org/wiki/Scaled_inverse_chi-squared_distribution), $g\sim\text{Scale-inv-}\chi^2(1,\ \color{lightgray}{h^2=}\varphi)$.

&nbsp;

<a href="#top"> scroll to top </a>

&nbsp;
&nbsp;


### BIC Approximation

&nbsp;

Approximating Bayes Factors from Bayesian Information Criterion (BIC)

&nbsp;
&nbsp;

The BIC for model $i$ with the number of free parameters $k_i$ and the sample size $n$ is $\text{BIC}(\mathcal{M}_i)=-2\ln{\mathcal{L}_i}+k_i\ln{n}\quad$ for $n\gg k_i$,

where $\mathcal{L}_i=p(\boldsymbol{y}\mid\hat{\boldsymbol{\theta}}_i,\mathcal{M}_i)$ is the maximum likelihood.

&nbsp;

Use the second-order Taylor approximation of log-marginal likelihood of each model.

\begin{equation} \tag{8}
\mathit{BF}_{10}\approx\exp\left\{\frac{\text{BIC}(\mathcal{M}_0)-\text{BIC}(\mathcal{M}_1)}{2}\right\}
\end{equation}

Derivation in Wagenmakers ([2007](https://dx.doi.org/10.3758/BF03194105), Appendix B).

&nbsp;

```{r SBC, eval=F}
?BIC
```

&nbsp;

<a href="#top"> scroll to top </a>

&nbsp;
&nbsp;


### Approximate Bayes Factor

&nbsp;

The test statistic in a Wald test is $W=\left(\frac{\hat{\theta}-\theta_0}{\text{se}(\hat{\theta})}\right)^2\sim\chi_1^2$ under $\mathcal{H}_0$, where $\hat{\theta}$ is the maximum likelihood estimate, and $\text{se}(\hat{\theta})$ is the standard error.

Wagenmakers ([2022](https://doi.org/10.31234/osf.io/egydq); its Eq. 6) reviewed the Jeffreys-style approximate Bayes factor (JAB) and proposed a piecewise approximation (WAB) from $p$-value and effective sample size $N$ in Eq. 10.

The complication with JAB is that the approximation is valid **only** when $\text{se}(\hat{\theta})\ll\sigma_g$, where $\sigma_g$ is the scale of the prior distribution $g(\theta)$. Note that with very small sample sizes, this assumption is not true and Eq. 9 is biased against $\mathcal{H}_0$.

\begin{equation} \tag{9}
\textit{JAB}_{01}=A\cdot\sqrt{N}\exp\left\{-\frac{1}{2}W\right\}
\end{equation}

A normal unit-information prior for $\theta$ yields $A=1$. Based on the Jeffreys prior, the general approximation specifies $A=\sqrt{\pi/2}\approx1.253$.

&nbsp;

Wagenmakers's piecewise approximation is

\[
\textit{WAB}_{01}\approx
\begin{cases} \tag{10}
  \text{$\ p^{\frac{1}{4}}\sqrt{N}$,} & \text{$p>0.5$} \\
  \\
  \text{$\ \sqrt{pN}$,} & \text{$0.1<p\leqslant0.5$ (simpler)} \\
  \\
  \text{$\ \frac{4}{3}p^{\frac{2}{3}}\sqrt{N}$,} & \text{$0.1<p\leqslant0.5$ (more precise)} \\
  \\
  \text{$\ 3p\sqrt{N}$,} & \text{$p\leqslant0.1$}
\end{cases} 
\]

&nbsp;

```{r Wald, eval=F}
qchisq(p, 1, lower.tail=F) # inverse CDF of chisq(1): 1-p → W = t² → JAB
qchisq(1-p, 1) # same
```

&nbsp;
&nbsp;
&nbsp;

**Note**: The two-sided $p$-value of the $t$-statistic differs from the $p$-value of the Wald statistic, particularly when the sample size is insufficient.

```{r pVal}
# suppose t(𝜈 = 10 + 10 - 2) = 0.77, compute the p-values
c(pt(0.77, 18, lower.tail=F)*2, pchisq(0.77^2, 1, lower.tail=F))

# suppose a two-sided p = 0.45, compute the test statistics t² and W
c(qt(0.45/2, 18, lower.tail=F)^2, qchisq(0.45, 1, lower.tail=F)) # try increasing 𝜈
```

&nbsp;

<a href="#top"> scroll to top </a>

&nbsp;
&nbsp;


### Savage–Dickey Density Ratio

&nbsp;

**The Savage–Dickey density ratio** is a special form of the Bayes factor for **nested** models by dividing the value of <br><span style="color:#0055A4">the posterior density over the parameters for the alternative model evaluated at the hypothesized value</span> by <br><span style="color:#EF4135">the prior for the same model evaluated at the same point</span>.

\begin{equation} \tag{11}
    \textit{BF}_{01}:=\frac{p(\boldsymbol{y}\mid \mathcal{M}_0)}{p(\boldsymbol{y}\mid \mathcal{M}_1)}=\frac{\color{#0055A4}{p(\boldsymbol{\theta}=\boldsymbol{\theta}_0\mid\boldsymbol{y},\mathcal{M}_1)}}{\color{#EF4135}{p(\boldsymbol{\theta}=\boldsymbol{\theta}_0\mid \mathcal{M}_1)}}
\end{equation}

&nbsp;

```{r rstan, eval=F}
stancode1M1 <- "
data {
  int<lower=1> N;           // total number of observations
  vector[N] x;              // predictor variable
  vector[N] y;              // response variable
  vector[2] mu;             // mean vector for theta
  matrix[2,2] Sigma;        // covariance matrix for theta
}

parameters {
  vector[2] theta;          // combined vector for alpha and beta
  real<lower=0> sigma;      // error standard deviation
}

transformed parameters {
  real alpha = theta[1];    // intercept
  real beta = theta[2];     // slope
}

model {
  // Priors
  theta ~ multi_normal(mu, Sigma);    // unit information
  target += -log(sigma);              // Jeffreys prior

  // Likelihood (linear regression model)
  y ~ normal(alpha + beta * x, sigma);
}"

stanmodel1M1 <- stan_model(model_code=stancode1M1)
```

&nbsp;

<a href="#top"> scroll to top </a>

&nbsp;
&nbsp;


## 2. Simulation

&nbsp;

<font size="4"> Effect Size </font>

\begin{equation}
    Y=\beta_0+\beta_1X+\epsilon
\end{equation}

If $X\sim\mathcal{N}(0,\ \sigma_X^2)\ \perp \!\!\! \perp\ \epsilon\sim\mathcal{N}(0,\ \sigma_\epsilon^2)$,&nbsp;&nbsp; then $Y\sim\mathcal{N}(\beta_0,\ \color{lightgray}{\sigma_Y^2=}\beta_1^2\sigma_X^2+\sigma_\epsilon^2)$.

&nbsp;

In the linear regression analysis, the Pearson's correlation coefficient is $r:=\frac{\text{Cov}(Y,X)}{\sigma_X\sigma_Y}$. Recall that $\frac{\text{Cov}(Y,X)}{\sigma_X^2}=\frac{\text{Cov}(\beta_0+\beta_1X+\epsilon,\ X)}{\sigma_X^2}=\beta_1$.

Thus, the **standardized slope** (also known as the beta coefficient in standardized units) is $r=\beta_1\frac{\sigma_X}{\sigma_Y}=\beta_1\frac{\sigma_X}{\sqrt{\beta_1^2\sigma_X^2+\sigma_\epsilon^2}}$.

And, the (unstandardized) regression slope is $\beta_1=\frac{\sigma_\epsilon}{\sigma_X}\cdot\frac{r}{\sqrt{1-r^2}}$.

&nbsp;

* Cohen ([1988](https://doi.org/10.4324/9780203771587), p. 76-81) suggests that the effect sizes are small, medium, and large for $r\in\{0.1,\ 0.3,\ 0.5\}$, respectively.

&nbsp;
&nbsp;

<font size="4"> Data </font>

```{r data}
simData2 <- function(n, intercept=NULL, slope=NULL, sd.X=2, sd.e=10) {
  #' Input - 
  #' n :             number of observations
  #' intercept:      true baseline at x = 0
  #' slope:          true regression coefficient; Cov(Y, X) / sd.X²
  #' sd.X:           true standard deviation of X
  #' sd.e:           true error standard deviation
  #' 
  #' Details - 
  #' sd.Y²:          variance of Y; slope² · sd.X² + sd.e²
  #' r:              standardized slope; Cov(Y, X) / (sd.X · sd.Y)
  #'                 r ≡ slope · sd.X / sd.Y  ==>
  #'             slope ≡ r · sd.e / (sd.X · sqrt(1 - r²))
  #'             
  #'                 r = 0.1 small
  #'                 r = 0.3 medium
  #'                 r = 0.5 large effect size (Cohen, 1988, p. 76-81)
  #' 
  #' Output - dependent and independent variables, y and x
  
  # no random seed
  x <- stats::rnorm(n, 0, sd.X) # independent variable
  epsilon <- stats::rnorm(n, 0, sd.e) # error term
  data.frame("DV"=intercept + slope * x + epsilon, "IV"=x)
}
```

&nbsp;
&nbsp;

<font size="4"> The Bayes Factors </font>

**Note**: All the Bayes factors are $\textit{BF}_{01}$.

```{r BFs}
computeBFs2 <- function(data, precise=F) {
  #' Input -
  #' data:     long format data whose columns must be `DV` and `IV`
  #' precise:  logical; if FALSE (default), use the simpler form for Output 6;
  #'                             otherwise, use the more precise form
  #' 
  #' Global -  stanmodel1M1
  #' Dependency - BayesFactor (v0.9.12-4.7), rstan (v2.32.6), polspline (v1.1.25)
  #' Output -  a vector of the Bayes factors in favor of the null
  #'           1. Savage–Dickey density ratio (Eq. 11) using the normal unit-information prior
  #'           2. lmBF function or regressionBF function
  #'           3. BIC approximation to the Bayes factor in Eq. 8; BIC(fit)
  #'           4. Jeffreys approximate Bayes factor in Eq. 9 using the normal unit-information prior
  #'           5. Jeffreys approximate Bayes factor in Eq. 9 using the Jeffreys prior
  #'           6. Wagenmakers approximate Bayes factor based on the p-value and sample size in Eq. 10
  
  data <- stats::na.omit(data)
  N <- nrow(data) # number of observations
  
  fit0 <- stats::lm(DV ~ 1, data) # null model fit,  ŷ = ^β₀
  fit1 <- stats::lm(DV ~ IV, data) # full model fit, ŷ = ^β₀ + ^β₁ · x
  pVal <- summary(fit1)$coefficients["IV", "Pr(>|t|)"] # p-value of the regression coefficient
  tVal <- stats::qt(pVal/2, N-2, lower.tail=F) # t-statistic (^β₁ >= 0), ["IV", "t value"]
  
  datalist <- list(N=N, x=data$IV, y=data$DV,
                   mu=stats::coef(fit1), Sigma=stats::vcov(fit1)*N)
  stanfitM1 <- rstan::sampling(stanmodel1M1, data=datalist,
                               iter=15000, warmup=5000, chains=2, seed=277, refresh=0)
  betaS <- rstan::extract(stanfitM1, pars="beta")$beta
  dPost <- polspline::dlogspline(0, polspline::logspline(betaS)) # estimated posterior density
  dPrior <- stats::dnorm(0, datalist$mu[2], sqrt(datalist$Sigma[2,2]))
  SDdr01 <- dPost / dPrior # Savage–Dickey density ratio
  
  bf <- BayesFactor::lmBF(DV ~ IV, data, progress=F) # lmBF10
  BF01 <- ifelse(BayesFactor::extractBF(bf)$error < 0.01, # Monte Carlo error
                 1 / BayesFactor::extractBF(bf)$bf, NA) # lmBF01
  
  BICB01 <- exp((stats::BIC(fit1) - stats::BIC(fit0)) / 2) # Eq. 8
  
  # sqrt(N) * exp(-0.5 * stats::qchisq(1-pVal, 1)) #  <== CAUTION, tiny p-value --> infinity Wald statistic
  JAB01_UI <- sqrt(N) * exp(-0.5 * tVal * tVal) # Eq. 9, unit-information prior

  WAB01 <- ifelse(pVal <= .5,
                  ifelse(pVal > .1,
                         ifelse(precise,
                                4 * pVal^(2/3) * sqrt(N) / 3, # 0.1 < p <= 0.5 (more precise)
                                sqrt(pVal * N)), # 0.1 < p <= 0.5 (simpler)
                         3 * pVal * sqrt(N)), # p <= 0.1
                  pVal^(1/4) * sqrt(N)) # p > 0.5, Eq. 10
  
  c("SD"=SDdr01, "lmBF"=BF01, "BIC_approx"=BICB01, 
                 "JAB"=JAB01_UI, "JAB_J"=sqrt(pi/2) * JAB01_UI, "WAB"=WAB01)
}

# set.seed(277)
# test2 <- simData2(30, 10, 2)
# computeBFs2(test2)
```

&nbsp;
&nbsp;

<font size="4"> Simulation Study </font>

We aim to conduct several simulation studies to evaluate the accuracy of approximate objective Bayes factors in their ability to convert $p$-values and $n$ reported in scientific articles to Bayesian measures of evidence.

&nbsp;

```{r sim-not-run, eval=F}
nSim <- 1000 # number of simulation runs for each setting
nObs <- c(10, 20, 30, 100, 500) # number of observations
BETA1 <- c("Null"=0, "Small"=0.5, "Medium"=1.6, "Large"=2.9) # regression coefficient (slope)

reportBF <- reportBF_avg <- reportBF_sd <- reportERR <- reportERR_avg <- reportERR_sd <- 
  reportRULE <- reportRULE_H1 <- reportRULE_H0 <- reportRULE_inc <-
  setNames(vector(mode="list", length=length(nObs) * length(BETA1)), # pre-allocation
           apply(expand.grid(BETA1, nObs), 1, function(r) paste0("n = ", r[2], ", slope = ", r[1])))

index <- 1 # initial list index

for (n in nObs) {
  
  for (slope in BETA1) {
    set.seed(n+slope+277)
    temp <- t(replicate(nSim, computeBFs2(simData2(n, intercept=10, slope=slope)))) # replicate the Bayes factors
    reportBF[[index]] <- temp
    reportBF_avg[[index]] <- colMeans(temp, na.rm=T) # mean Bayes factors in favor of the null
    reportBF_sd[[index]] <- apply(na.omit(temp), 2, sd) # sample standard deviations of them
    
    tempERR <- 100* (temp - temp[,1]) / temp[,1] # percent errors; the first column should be all 0
    reportERR[[index]] <- tempERR # (can take in NA)
    reportERR_avg[[index]] <- colMeans(tempERR, na.rm=T) # mean percent errors
    reportERR_sd[[index]] <- apply(na.omit(tempERR), 2, sd) # sample standard deviations of the percent errors
    
    tempRULE_H1 <- temp < 1/3 # support H1
    tempRULE_H0 <- temp > 3 # support H0
    tempRULE_inc <- temp >= 1/3 & temp <= 3 # inconclusive
    
    # counts of matching decisions for H1
    reportRULE_H1[[index]] <- colSums(tempRULE_H1 * tempRULE_H1[,1], na.rm=T)
    
    # counts of matching decisions for H0
    reportRULE_H0[[index]] <- colSums(tempRULE_H0 * tempRULE_H0[,1], na.rm=T)
    
    # counts of matching inconclusiveness
    reportRULE_inc[[index]] <- colSums(tempRULE_inc * tempRULE_inc[,1], na.rm=T)
    
    tempRULE <- 1*(temp < 1/3) - 1*(temp > 3) # 1: support H1;  0: inconclusive;  -1: support H0
    reportRULE[[index]] <- colSums(tempRULE[,1] == tempRULE, na.rm=T) # counts of matching decisions
    # colMeans(); a value near 1 suggests matching results; the first column should be all 1
    
    index <- index + 1
  }
} # roughly 4.6 hours of run time

# Warning messages:
# 1-8: In polspline::logspline(betaS) :  Not all models could be fitted
```

&nbsp;

```{r prep}
reshapeW2L2 <- function(list, prob=T, lab=c("SD", "BF", "BIC", "JAB", "JAB_J", "WAB")) {
  #' Input -
  #' list:    a list of size length(nObs) * length(BETA1);
  #'          reportERR (unaggregated),
  #'          reportRULE, reportRULE_H1, reportRULE_H0, or reportRULE_inc (aggregated)
  #' prob:    logical; if TRUE (default), return the proportions;
  #'                           otherwise, return the original values
  #' lab:     a vector of characters representing the Bayes factor methods
  #' 
  #' Global -
  #' nSim:    number of simulation runs for each setting
  #' nObs:    number of observations
  #' BETA1:   regression coefficient (slope)
  #' 
  #' Output - unlist the report and reshape the wide into the long
  
  len <- length(lab) # six methods
  nRep <- ifelse(!prob, nSim, 1) # nSim if unaggregated; 1 if aggregated
  long <- data.frame("value"=unname(unlist(list)),
                     "method"=factor(rep(rep(1:len, each=nRep), length(list)), labels=lab),
                     "n"=rep(nObs, each=nRep*len*length(BETA1)),
                     "slope"=rep(rep(unname(BETA1), each=nRep*len), length(nObs)))
  
  if (prob) {
    list2 <- lapply(list, function(vec) {
      if (vec[1] == 0) {
        vec * NA
      } else {
        vec / vec[1] # convert counts to proportions
      }
    })
    list3 <- lapply(list, function(vec) rep(vec[1], length(vec))) # denominator
    long$total <- unname(unlist(list3))
    long$prop <- unname(unlist(list2))
  }
  long
}


formatting <- function(val) {
  #' format the value using scientific notation
  if (abs(val) >= 10) {
    paste("~italic(BF)[\"01\"] %~~%", 
          gsub("e\\+0*(\\d+)", "%*%10^\\1", sprintf("%.2e", val)))
  } else if (abs(val) < 1) {
    paste("~italic(BF)[\"01\"] %~~%", 
          gsub("e-0*(\\d+)", "%*%10^{-\\1}", sprintf("%.2e", val)))
  } else {
    paste("~italic(BF)[\"01\"] %~~%", sprintf("%.2f", val))
  }
}
```

&nbsp;

<a href="#top"> scroll to top </a>

&nbsp;
&nbsp;


## 3. Graphic Results {.tabset .tabset-fade}

&nbsp;

<font size="4"> Visualization </font>

Note: The Jeffreys approximate Bayes factor (`JAB_J`) in Eq. 9, using the Jeffreys prior, will not be plotted.

### Boxplot (Percent Errors)

```{r load, eval=runthis, echo=F}
load("~/Downloads/lmSim.RData")
```
```{r viz-boxplot, eval=runthis, fig.width=10, fig.height=5, results="asis"}
# percent errors
df2 <- reshapeW2L2(reportERR, prob=F)

index <- 0

# boxplot of the percent errors
for (beta in BETA1) {
  baseBF01 <- sapply(reportBF_avg[seq(1, by=length(BETA1), length.out=length(nObs)) + index], 
                     function(x) x[1]) # baseline is the Savage–Dickey density ratio
  baseBF01_lab <- sapply(baseBF01, formatting) # scientific notation
  index <- index + 1
  
  sub <- subset(df2, !(method %in% c("SD", "BF", "JAB_J")) & slope == beta) # subset the long
  anno <- data.frame("label"=baseBF01_lab, "n"=unique(sub$n),
                     "x"=2, "y"=.9 * aggregate(value~n, sub, FUN=min)$value) # for annotation use
  
  print(ggplot(sub, aes(x=method, y=value)) +
          geom_boxplot(alpha=0.3, na.rm=T) +
          geom_text(data=anno, aes(label=label, x=x, y=y), parse=T, col="red") +
          facet_wrap(.~n, scales="free_y", nrow=1,
                     labeller=label_bquote(cols=italic(n)==.(n))) +
          labs(x="\nBayes Factor Methods", y="Percent Error (%)\n") +
          ggtitle(paste0(names(BETA1[BETA1==beta]), " Effect Size of the Slope")) +
          geom_hline(yintercept=0, linetype="dashed", color="gray") + # reference line 0%
          theme_classic() +
          theme(axis.text.x=element_text(angle=90)))
  cat("<br><br><br><br><br>")
}
```

&nbsp;

<a href="#top"> scroll to top </a>

&nbsp;
&nbsp;

### Boxplot ($\textit{BF}_{01}$)

```{r viz-boxplot2, eval=runthis, fig.width=10, fig.height=5, results="asis"}
df_BF <- reshapeW2L2(reportBF, prob=F)

# boxplot of the BF₀₁
for (beta in BETA1) {
  sub <- subset(df_BF, slope == beta & method != "JAB_J") # subset the long
  print(ggplot(sub, aes(x=method, y=value)) +
          geom_boxplot(alpha=0.3, na.rm=T) +
          facet_wrap(.~n, scales="free_y", nrow=1,
                     labeller=label_bquote(cols=italic(n)==.(n))) +
          labs(x="\nBayes Factor Methods", y=expression(italic("BF")["01"])) +
          ggtitle(paste0(names(BETA1[BETA1==beta]), " Effect Size of the Slope")) +
          theme_classic() +
          theme(axis.text.x=element_text(angle=90)))
  cat("<br><br><br><br><br>")
}
```

&nbsp;

<a href="#top"> scroll to top </a>

&nbsp;
&nbsp;

### Line Chart (Proportions of Agreement)

```{r viz-line, eval=runthis, fig.width=9, fig.height=7, warning=F}
# decisions
decis2 <- rbind(reshapeW2L2(reportRULE),
                reshapeW2L2(reportRULE_H1),
                reshapeW2L2(reportRULE_H0),
                reshapeW2L2(reportRULE_inc))
decis2$result <- factor(rep(c("Overall", "BF01 < 1/3", "BF01 > 3", "[1/3, 3]"),
                            each=length(levels(df2$method)) * length(nObs) * length(BETA1)),
                        levels=c("Overall", "BF01 < 1/3", "BF01 > 3", "[1/3, 3]"))
decis2$delta <- paste0("Slope = ", decis2$slope)
decis2 <- decis2[!is.na(decis2[,"prop"]),]
if (F) {
  # compute the pointwise confidence intervals using the normal approximation
  decis2$half.len <- qnorm(1-.05/2) * 
    sqrt(decis2$value * (decis2$total - decis2$value) / decis2$total^3) # 95% CI width
  decis2$lwr <- max(0, decis2$value / decis2$total - decis2$half.len) # 95% CI lower bound
  decis2$upr <- min(1, decis2$value / decis2$total + decis2$half.len) # 95% CI upper bound
} else { # Jeffreys intervals
  decis2$lwr <- qbeta(.05/2, decis2$value, decis2$total - decis2$value + 1) # 95% CI lower bound
  decis2$upr <- qbeta(1-.05/2, decis2$value + 1, decis2$total - decis2$value) # 95% CI upper bound
  decis2$half.len <- (decis2$upr - decis2$lwr) / 2 # 95% CI width
}

# line chart of the proportions of agreement
ggplotly(ggplot(decis2, aes(n, prop, color=method, label=value)) +
           geom_line(linewidth=1.2, na.rm=T) + geom_point(size=2, na.rm=T) +
           geom_ribbon(aes(ymin=lwr, ymax=upr), alpha=.2) +
           scale_color_manual(values=c("black", "gray", "#E69F00", "#CC79A7", "#D55E00", "#0072B2"),
                              labels=c("SD", "BF", "BIC", "JAB", "JAB_J", "WAB")) +
           facet_grid(result~delta) +
           labs(x="Number of Observations", y="Proportion of Agreement", color="Method") +
           scale_x_continuous(breaks=c(10, 100, 500)) +
           scale_y_continuous(breaks=c(0, 0.5, 1)) + # min(decis2$prop, na.rm=T)
           theme_minimal()) # interactive plots
```
&nbsp;
&nbsp;
&nbsp;
```{r viz-line2, eval=runthis, echo=F, fig.width=10, fig.height=10, message=F}
decis2 <- rbind(reshapeW2L2(reportRULE),
                reshapeW2L2(reportRULE_H1),
                reshapeW2L2(reportRULE_H0),
                reshapeW2L2(reportRULE_inc))
decis2$result <- factor(rep(c("Overall", "H1", "H0", "Inconclusive"),
                            each=length(levels(df2$method)) * length(nObs) * length(BETA1)),
                        levels=c("Overall", "H1", "H0", "Inconclusive"),
                        labels=c("Overall", "italic(BF)['01'] < 1/3", 
                                 "italic(BF)['01'] > 3", "italic(BF)['01'] %in% group('[',list(1/3, 3),']')"))
decis2$slope <- factor(decis2$slope, levels=unname(BETA1),
                       labels=paste0("italic(β)[1] == ", unname(BETA1), "~(", names(BETA1), ")"))
set.seed(277)
decis2_plot <- subset(decis2[sample(1:nrow(decis2), nrow(decis2)),], 
                      !(method %in% c("SD", "JAB_J"))) # shuffle to prevent overlapping

ggplot(decis2_plot, aes(n, prop, color=method, shape=method)) +
  geom_line(linewidth=1.2, na.rm=T) + geom_point(size=3.5, na.rm=T) +
  scale_color_manual(values=c("gray", "#E69F00", "#CC79A7", "#0072B2"),
                     labels=c("BF", "BIC", "JAB", "WAB")) +
  scale_shape_manual(values=c(10, 8, 18, 15),
                     labels=c("BF", "BIC", "JAB", "WAB")) +
  facet_grid(result~slope, labeller=label_parsed) +
  labs(x="\nNumber of Observations", y="Proportion of Agreement\n", color="Method", shape="Method",
       title="Simple Linear Regression") +
  scale_x_continuous(breaks=c(10, 100, 500)) +
  scale_y_continuous(breaks=c(0, 0.5, 1)) +
  theme_minimal() +
  theme(legend.position="bottom", text=element_text(size=17),
        legend.title=element_text(margin=margin(0,1,0,0,"cm")),
        legend.text=element_text(margin=margin(0,0.5,0,0.5,"cm"))) +
  guides(color=guide_legend(override.aes=list(linewidth=0)),
         shape=guide_legend(override.aes=list(size=6), nrow=1))
```

&nbsp;

<a href="#top"> scroll to top </a>

&nbsp;
&nbsp;


## 4. Decision Rules

&nbsp;

<span style="color:red"><font size="3"> Debugging </font></span>

```{r debug, error=T}
(mat <- matrix(rep(c(0, -1), 5), nrow=2))
(res <- mat[mat[,1] == -1, 1] == mat[mat[,1] == -1,]) # only one row ==> column vector

colSums(res) #  ERROR!
colSums(rbind(rep(F, 5), res)) # counts of matching decisions

colSums((mat == -1) * (mat == -1)[,1]) # counts of matching decisions
```

&nbsp;

The counts of matching Bayes factors between the gold standard and approximate methods.

```{r rules, eval=runthis}
reportRULE
reportRULE_inc
reportRULE_H1
reportRULE_H0
```

&nbsp;

<a href="#top"> scroll to top </a>

&nbsp;
&nbsp;


## 5. Finite Sample Correction

&nbsp;

Centering the normal unit-information prior on the null hypothesized value $\theta_0$, the customary choice in objective Bayesian testing yields the following JAB* (its Eq. 4; Wagenmakers, [2022](https://doi.org/10.31234/osf.io/egydq)).

\begin{equation}
  \textit{JAB}_{01}^*\approx\sqrt{N}\exp{\left\{-\frac{N-1}{2}\left(\frac{\hat{\theta}-\theta_0}{\sigma_g}\right)^2\right\}}
\end{equation}

Alternatively, the JAB in Eq. 9 centers the prior on the maximum likelihood estimate.

\begin{equation}
  \textit{JAB}_{01}\approx\sqrt{N}\exp{\left\{-\frac{N}{2}\left(\frac{\hat{\theta}-\theta_0}{\sigma_g}\right)^2\right\}}
\end{equation}

Using a power transformation, it can be shown that $\textit{JAB}_{01}^*=\textit{JAB}_{01}^{\frac{N-1}{N}}\cdot N^{\frac{1}{2N}}$.

&nbsp;

$N=10$

```{r (N-1)/N, eval=runthis, fig.width=10, fig.height=5}
par(mfrow=c(1, length(BETA1)))
for (i in 1:length(BETA1)) {
  reportBF[[i]][,5] <- reportBF[[i]][,4]^(9/10) * 10^(1/20) # number of observations is 10
  boxplot(reportBF[[i]][,c(1,4:6)], names=c("SD", "JAB", "JAB*", "WAB"),
          xlab="", ylab=expression(italic("BF")["01"]), main=paste0(names(BETA1)[i], " Slope"), las=2)
}
```

&nbsp;

<a href="#top"> scroll to top </a>

&nbsp;
&nbsp;


## 6. eJAB

&nbsp;

To test the hypotheses $\mathcal{H}_0:\boldsymbol{\theta}=\boldsymbol{\theta}_0$ versus $\mathcal{H}_1:\boldsymbol{\theta}\neq\boldsymbol{\theta}_0$, we define the extended Jeffreys approximate objective Bayes factor (eJAB) as
\begin{equation}
  \mathit{eJAB}_{01}=\sqrt{N}\exp\left\{-\frac{1}{2}\frac{N^{1/q}-1}{N^{1/q}}\cdot Q_{\chi^{2}_{q}}(1-p)\right\},
\end{equation}

where $q$ is the size of the parameter vector $\boldsymbol{\theta}$, $N$ is the sample size, $Q_{\chi^{2}_{q}}(\cdot)$ is the quantile function of the chi-squared distribution with $q$ degrees of freedom, and $p$ is the $p$-value from a null-hypothesis significance test. We further define $\mathit{eJAB}_{10}=1\ \!/\ \!\mathit{eJAB}_{01}$.

&nbsp;

The sample size $N$ is

* the total number of observations for for $t$-tests, linear regression, logistic regression, one-way ANOVA, and chi-squared tests.

* the number of events for Cox models.

* the number of independent observations for one-way repeated-measures ANOVA (Nathoo & Masson, [2016](https://dx.doi.org/10.1016/j.jmp.2015.03.003)).

&nbsp;

The degrees of freedom are

* $q=1$ for $t$-tests, linear regression, logistic regression, and Cox models, where $\mathcal{H}_0$ specifies a point-null hypothesis (Wagenmakers, [2022](https://doi.org/10.31234/osf.io/egydq)).

* $q=I-1$ for one-way ANOVA and one-way repeated-measures ANOVA, where $I$ is the number of conditions.

* $q=(R-1)(C-1)$ for chi-squared tests, where $R$ and $C$ are the numbers of rows and columns, respectively.

Since the Bayes factor is the ratio of the marginal likelihoods of two competing models, $q=p_1-p_0$ also represents the difference in the number of free parameters between the alternative and null models.

&nbsp;

For further reading, please visit <https://osf.io/qeprj/>.

&nbsp;

<a href="#top"> scroll to top </a>

&nbsp;
&nbsp;

