---
title: "Generalized Jeffreys's Approximate Objective Bayes Factor"
subtitle: "Part One: Two-Sample *t*-Test Simulation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
runthis <- T # whether to run chunks 13-17 & 19-24

if(!require(pwr)) {install.packages("pwr")}
if(!require(BayesFactor)) {install.packages("BayesFactor")}
if(!require(rstan)) {install.packages("rstan")}
if(!require(polspline)) {install.packages("polspline")}
if(!require(ggplot2)) {install.packages("ggplot2"); library(ggplot2)}
if(!require(plotly)) {install.packages("plotly"); library(plotly)}
if(!require(remotes)) {install.packages("remotes")}
if(!require(klippy)) {remotes::install_github("rlesur/klippy")}
```
```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```
============================================================================================================ 

Open Science Framework: <https://osf.io/qeprj/>

&nbsp;

# {.tabset .tabset-fade .tabset-pills}
## 1. Methods {.tabset .tabset-fade}

### "*BayesFactor*" R Package

&nbsp;

\begin{align}
&Y_{1j}\overset{\text{i.i.d.}}{\sim}\mathcal{N}(\mu+\sigma_\epsilon \delta/2,\ \sigma_\epsilon^2)\quad\text{for }j=1,\dotsb,n_1, \\
&Y_{2j}\overset{\text{i.i.d.}}{\sim}\mathcal{N}(\mu-\sigma_\epsilon \delta/2,\ \sigma_\epsilon^2)\quad\text{for }j=1,\dotsb,n_2,\text{ where }\delta=(\mu_1-\mu_2)/\sigma_\epsilon \tag{$*$}
\end{align}

\begin{equation} \tag{2}
  \pi(\mu,\sigma_\epsilon^2)\propto 1/\ \sigma_\epsilon^{2}
\end{equation}

\begin{equation} \tag{$\dagger$}
\delta\sim\text{Cauchy}(0,\ \gamma)=\frac{\gamma}{(\delta^2+\gamma^2)\pi}
\end{equation}

$\mathcal{H}_0:\delta=0$ versus $\mathcal{H}_1:\delta\neq0$

&nbsp;

A Bayesian two-sample $t$-test is equivalent to a Bayesian one-way analysis of variance (ANOVA) with two conditions ($a=2$).

\begin{align} \tag{1}
    \mathcal{M}_1:\ Y_{ij}&=\mu+\sigma_\epsilon d_i+\epsilon_{ij}\\ \text{versus}\quad\mathcal{M}_0:\ Y_{ij}&=\mu+\epsilon_{ij},\\ \qquad\qquad&\epsilon_{ij}\overset{\text{i.i.d.}}{\sim}\mathcal{N}(0,\sigma_\epsilon^2)\quad\text{ for condition } i=1,\dotsb,a;\text{ (unbalanced) subject } j=1,\dotsb,n_i
\end{align}

\begin{equation} \tag{2}
  \pi(\mu,\sigma_\epsilon^2)\propto 1/\ \sigma_\epsilon^{2}
\end{equation}

\begin{equation} \tag{3}
  d_i^\star\mid g\overset{\text{i.i.d.}}{\sim}\mathcal{N}(0,g)
\end{equation}

\begin{equation} \tag{4}
  g\sim\text{Scale-inv-}\chi^2(1,h^2)
\end{equation}

\begin{equation} \tag{5}
  (d_1^\star,\dotsb,d_{a-1}^\star)=(d_1,\dotsb,d_a)\cdot\mathbf{Q}
\end{equation}

\begin{equation} \tag{6}
  \mathbf{I}_a-\frac{1}{a}\mathbf{J}_a=\mathbf{Q}\cdot\mathbf{Q}^\top
\end{equation}

By default, $h=0.5$ for the fixed effects (in this case) and $h=1$ for the random effects in Eq. 4.

$\mathbf{Q}$ is an $a\times(a-1)$ matrix of the $a-1$ eigenvectors of unit length corresponding to the nonzero eigenvalues of the left side term in Eq. 6. For example, the projected standardized effect $d^\star=(d_1-d_2)/\sqrt{2}$ when $a=2$. In the other direction (given $d_1+d_2=0$), $d_1=d^\star/\sqrt{2}$ and $d_2=-d^\star/\sqrt{2}$. Read more: <https://rpubs.com/sherloconan/1070217>.

&nbsp;

The Jeffreys–Zellner–Siow (JZS) Bayes factor for the Bayesian $t$-test is computed by integrating across one dimension in Eq. 7 (Ly et al., [2016](https://doi.org/10.1016/j.jmp.2015.06.004), p. 24; Rouder et al., [2012](https://doi.org/10.1016/j.jmp.2012.08.001), p. 360; Rouder et al., [2009](https://doi.org/10.3758/PBR.16.2.225), p. 237; Wei, Nathoo, & Masson, [2023](https://doi.org/10.3758/s13423-023-02295-1), p. 1762; see also the balanced one-way between-subjects ANOVA in Morey et al., [2011](https://doi.org/10.1016/j.jmp.2011.06.004), p. 374), using Gaussian quadrature.

\begin{equation} \tag{7}
\textit{JZS-BF}_{10}=(2\pi)^{-\frac{1}{2}}\gamma\cdot\left(1+\frac{t^2}{\nu}\right)^{\frac{\nu+1}{2}}\int_0^\infty(1+Ng)^{-\frac{1}{2}}\left(1+\frac{t^2}{(1+Ng)\nu}\right)^{-\frac{\nu+1}{2}}g^{-\frac{3}{2}}e^{-\frac{\gamma^2}{2g}}\text{d}g,
\end{equation}

where the default scale $\gamma=\sqrt{2}/2$ (*cf.* $h=0.5$; version 0.9.2 and later), the degrees of freedom $\nu=n_1+n_2-2$, the $t$-statistic $t=\frac{M_1-M_2}{s_p\cdot\sqrt{\frac{1}{n_1}+\frac{1}{n_2}}}$, $M_i$ are the sample condition means, and $s_p^2=\frac{(n_1-1)s_1^2+(n_2-1)s_2^2}{n_1+n_2-2}$ is the pooled variance in a two-sample $t$-test assuming equal variances between the two groups. The sample sizes are $n_1$ and $n_2$, and the sample variances are $s_1^2$ and $s_2^2$, respectively.

The **effective sample size** is $N=\frac{n_1\cdot n_2}{n_1+n_2}$. &nbsp;&nbsp;<span style="color:lightgray">See also the <a href="https://en.wikipedia.org/wiki/Harmonic_mean" target="_blank">harmonic mean</a>.</span>

```{r verify-N}
JZSBF10 <- function(gamma=1/sqrt(2), t, N, nu) {
  #' Input - 
  #' gamma:   scale parameter of the Cauchy prior, i.e.,
  #'          `rscale="medium"` or `rscale=1/sqrt(2)` in BayesFactor::ttestBF, or
  #'          `rscaleFixed="medium"` or `rscaleFixed=0.5` in BayesFactor::anovaBF
  #' t:       observed t test statistic
  #' N:       "effective" sample size
  #' nu:      degrees of freedom
  #' Output - the JZS Bayes factor for a t-test in Eq. 7

  coef <- (2*pi)^(-0.5) * gamma * (1+t^2/nu)^((nu+1)/2)
  integrand <- function(g) {
    (1+N*g)^(-0.5) * (1+t^2/((1+N*g)*nu))^(-(nu+1)/2) * g^(-1.5) * exp(-gamma^2/(2*g))
  }
  unname(coef * integrate(integrand, lower=0, upper=Inf)$value)
}

test <- t.test(extra ~ group, data=sleep, var.equal=T) # two-sample t-test
t_obs <- test$statistic # t-statistic
n1 <- table(sleep$group)[1]; n2 <- table(sleep$group)[2] # balanced sample sizes, n1 = n2 = 10

JZSBF10(t=t_obs, N=n1*n2/(n1+n2), nu=n1+n2-2) # verifying the formula of the effective sample size
BayesFactor::ttestBF(formula=extra ~ group, data=sleep)
```

&nbsp;

Unlike the *anovaBF* function, the *ttestBF* function of the "*BayesFactor*" R package yields negligible Monte Carlo errors and does not require specifying the number of Markov chain Monte Carlo iterations.

&nbsp;

Default functions return the same Bayes factor estimates for the independent-samples (two-sample) $t$-test and one-way ANOVA with two conditions. However, those functions return systematically different estimates for the paired $t$-test and one-way repeated-measures ANOVA with two conditions (see <a href="https://forum.cogsci.nl/discussion/2330/bayesian-t-tests-vs-anovas-in-within-subject-designs-2-conditions" target="_blank">discussion</a> and <a href="https://cran.r-project.org/web/packages/BayesFactor/vignettes/priors.html" target="_blank">vignettes</a>).

&nbsp;
&nbsp;

<span style="color:red"><font size="3"> Caution Extreme Bayes Factor Estimates </font></span>

```{r ttestBF}
set.seed(277)
(bf <- BayesFactor::ttestBF(rnorm(1000), rnorm(1000, 2.5)))
(val <- BayesFactor::extractBF(bf)$bf) # the extracted and stored value is infinity
```

&nbsp;

<a href="#top"> scroll to top </a>

&nbsp;
&nbsp;


### BIC Approximation

&nbsp;

Approximating Bayes Factors from Bayesian Information Criterion (BIC)

&nbsp;
&nbsp;

The BIC for model $i$ with the number of free parameters $k_i$ and the sample size $n$ is $\text{BIC}(\mathcal{M}_i)=-2\ln{\mathcal{L}_i}+k_i\ln{n}\quad$ for $n\gg k_i$,

where $\mathcal{L}_i=p(\boldsymbol{y}\mid\hat{\boldsymbol{\theta}}_i,\mathcal{M}_i)$ is the maximum likelihood.

&nbsp;

Use the second-order Taylor approximation of log-marginal likelihood of each model.

\begin{equation} \tag{8}
\mathit{BF}_{10}\approx\exp\left\{\frac{\text{BIC}(\mathcal{M}_0)-\text{BIC}(\mathcal{M}_1)}{2}\right\}
\end{equation}

Derivation in Wagenmakers ([2007](https://dx.doi.org/10.3758/BF03194105), Appendix B).

Discussion in Nathoo and Masson ([2016](https://dx.doi.org/10.1016/j.jmp.2015.03.003)).

&nbsp;

```{r SBC, eval=F}
?BIC
```

&nbsp;

<a href="#top"> scroll to top </a>

&nbsp;
&nbsp;


### Approximate Bayes Factor

&nbsp;

The test statistic in a Wald test is $W=\left(\frac{\hat{\theta}-\theta_0}{\text{se}(\hat{\theta})}\right)^2\sim\chi_1^2$ under $\mathcal{H}_0$, where $\hat{\theta}$ is the maximum likelihood estimate, and $\text{se}(\hat{\theta})$ is the standard error.

Wagenmakers ([2022](https://doi.org/10.31234/osf.io/egydq); its Eq. 6) reviewed the Jeffreys-style approximate Bayes factor (JAB) and proposed a piecewise approximation (WAB) from $p$-value and effective sample size $N$ in Eq. 10.

The complication with JAB is that the approximation is valid **only** when $\text{se}(\hat{\theta})\ll\sigma_g$, where $\sigma_g$ is the scale of the prior distribution $g(\theta)$. Note that with very small sample sizes, this assumption is not true and Eq. 9 is biased against $\mathcal{H}_0$.

\begin{equation} \tag{9}
\textit{JAB}_{01}=A\cdot\sqrt{N}\exp\left\{-\frac{1}{2}W\right\}
\end{equation}

A normal unit-information prior for $\theta$ yields $A=1$. Based on the Jeffreys prior, the general approximation specifies $A=\sqrt{\pi/2}\approx1.253$.

<span style="color:gray">The normal unit-information prior is a normal distribution with mean zero and variance equal to the information in a single observation.</span>

<span style="color:gray">Equation 2 specifies the Jeffreys priors for the grand mean and error variance.</span>

&nbsp;

Wagenmakers's piecewise approximation is

\[
\textit{WAB}_{01}\approx
\begin{cases} \tag{10}
  \text{$\ p^{\frac{1}{4}}\sqrt{N}$,} & \text{$p>0.5$} \\
  \\
  \text{$\ \sqrt{pN}$,} & \text{$0.1<p\leqslant0.5$ (simpler)} \\
  \\
  \text{$\ \frac{4}{3}p^{\frac{2}{3}}\sqrt{N}$,} & \text{$0.1<p\leqslant0.5$ (more precise)} \\
  \\
  \text{$\ 3p\sqrt{N}$,} & \text{$p\leqslant0.1$}
\end{cases} 
\]

&nbsp;

```{r Wald, eval=F}
qchisq(p, 1, lower.tail=F) # inverse CDF of chisq(1): 1-p → W = t² → JAB
qchisq(1-p, 1) # same
```

&nbsp;
&nbsp;
&nbsp;

**Note**: The two-sided $p$-value of the $t$-statistic differs from the $p$-value of the Wald statistic, particularly when the sample size is insufficient.

```{r pVal}
# suppose t(𝜈 = 10 + 10 - 2) = 0.77, compute the p-values
c(pt(0.77, 18, lower.tail=F)*2, pchisq(0.77^2, 1, lower.tail=F))

# suppose a two-sided p = 0.45, compute the test statistics t² and W
c(qt(0.45/2, 18, lower.tail=F)^2, qchisq(0.45, 1, lower.tail=F)) # try increasing 𝜈
```

&nbsp;

<a href="#top"> scroll to top </a>

&nbsp;

Sampling Distributions

* If $X_i\overset{\text{i.i.d.}}{\sim}\mathcal{N}(0,1)$ for $i=1,\dotsb,n$, then $Z=\sum_{i=1}^n X_i\sim\chi_n^2$ with $\mathbb{E}[Z]=n$ and $\text{Var}(Z)=2n$.

* If $Z_1\sim\chi_m^2\ \perp \!\!\! \perp\ Z_2\sim\chi_n^2$, then $F=\frac{Z_1/m}{Z_2/n}\sim F(m,n)$ with $\mathbb{E}[F]=\frac{n}{n-2}$ for $n>2$ and $\text{Var}(Z)=\frac{2n^2(m+n-2)}{m(n-2)^2(n-4)}$ for $n>4$.

* If $X\sim\mathcal{N}(0,1)\ \perp \!\!\! \perp\ Z\sim\chi_n^2$, then $T=\frac{X}{\sqrt{Z/n}}\sim t_n$ with $\mathbb{E}[T]=0$ for $n>1$ and $\text{Var}(T)=\frac{n}{n-2}$ for $n>2$.

&nbsp;
&nbsp;


### Savage–Dickey Density Ratio

&nbsp;

**The Savage–Dickey density ratio** is a special form of the Bayes factor for **nested** models by dividing the value of <br><span style="color:#0055A4">the posterior density over the parameters for the alternative model evaluated at the hypothesized value</span> by <br><span style="color:#EF4135">the prior for the same model evaluated at the same point</span>.

\begin{equation} \tag{11}
    \textit{BF}_{01}:=\frac{p(\boldsymbol{y}\mid \mathcal{M}_0)}{p(\boldsymbol{y}\mid \mathcal{M}_1)}=\frac{\color{#0055A4}{p(\boldsymbol{\theta}=\boldsymbol{\theta}_0\mid\boldsymbol{y},\mathcal{M}_1)}}{\color{#EF4135}{p(\boldsymbol{\theta}=\boldsymbol{\theta}_0\mid \mathcal{M}_1)}}
\end{equation}

&nbsp;

```{r rstan, eval=F}
stancode1M1 <- "
data {
  int<lower=1> N;           // total number of observations
  vector[N] x;              // predictor variable
  vector[N] y;              // response variable
  vector[2] mu;             // mean vector for theta
  matrix[2,2] Sigma;        // covariance matrix for theta
}

parameters {
  vector[2] theta;          // combined vector for alpha and beta
  real<lower=0> sigma;      // error standard deviation
}

transformed parameters {
  real alpha = theta[1];    // intercept
  real beta = theta[2];     // slope
}

model {
  // Priors
  theta ~ multi_normal(mu, Sigma);    // unit information
  target += -log(sigma);              // Jeffreys prior

  // Likelihood (linear regression model)
  y ~ normal(alpha + beta * x, sigma);
}"

stanmodel1M1 <- stan_model(model_code=stancode1M1)
```

&nbsp;

<a href="#top"> scroll to top </a>

&nbsp;
&nbsp;


## 2. Simulation

&nbsp;

<font size="4"> Data </font>

```{r data, fig.width=5.5, fig.height=4.5, warning=F}
simData <- function(n1, n2, m1=0, delta=NULL, sd=1, power=NULL) {
  #' Input - 
  #' n1:      sample size of the first group
  #' n2:      sample size of the second group
  #' m1:      population mean of the first group
  #' delta:   population mean difference, m2 - m1
  #' sd:      population standard deviation
  #' power:   statistical power;
  #'          need to input either `delta` or `power`
  #' 
  #' Dependency - pwr (v1.3-0)
  #' Output - two-sample (unbalanced) data in the long format
  
  if (is.null(delta) & !is.null(power)) {
    delta <- sd * pwr::pwr.t2n.test(n1=n1, n2=n2, power=power)$d
  } else if (is.null(delta) & is.null(power)) {
    delta <- 0 # simulation under the null
  } else if (!is.null(delta) & !is.null(power)) {stop("Bad input!")}
  
  # no random seed
  data.frame("resp"=c(stats::rnorm(n1, m1,         sd),
                      stats::rnorm(n2, m1 + delta, sd)),
             "grp"=rep(c("grp1", "grp2"), c(n1, n2)),
             stringsAsFactors=T)
}


num <- 1000
p_par <- p_nonpar <- numeric(num)
for (i in 1:num) {
  set.seed(i)
  data <- simData(n1=30, n2=30, power=.8)
  p_par[i] <- t.test(resp ~ grp, data=data, var.equal=T)$p.value
  p_nonpar[i] <- wilcox.test(formula=resp ~ grp, data=data)$p.value # Mann–Whitney 𝑈-test
}
plot(p_par, p_nonpar, xlim=range(c(p_par, p_nonpar)), ylim=range(c(p_par, p_nonpar)),
     xlab=expression(italic(p) * "-Values from " * italic(t) * "-Tests"),
     ylab=expression(italic(p) * "-Values from " * italic(U) * "-Tests"),
     main=paste0("Two-Sample Designs (", num, " Runs)"))
abline(0, 1, col="gray", lty=2, lwd=1.5)
```

&nbsp;
&nbsp;

<font size="4"> The Bayes Factors </font>

**Note**: All the Bayes factors are $\textit{BF}_{01}$.

```{r BFs}
computeBFs <- function(data, total=F, precise=F) {
  #' Input -
  #' data:     two-sample data in the long format whose columns must be `resp` and `grp`
  #' total:    logical; if FALSE (default), use the effective sample size for Output 5-7
  #'                             otherwise, use the total number of observations
  #' precise:  logical; if FALSE (default), use the simpler form for Output 7;
  #'                             otherwise, use the more precise form
  #' 
  #' Global -  stanmodel1M1
  #' Dependency - BayesFactor (v0.9.12-4.7), rstan (v2.32.6), polspline (v1.1.25)
  #' Output -  a vector of the Bayes factors in favor of the null
  #'           1. Savage–Dickey density ratio (Eq. 11) using the normal unit-information prior
  #'           2. ttestBF function (Eq. 7)
  #'           3. BIC approximation to the Bayes factor in Eq. 8; BIC(fit)
  #'           4. SBC approximation to the Bayes factor in Eq. 8; AIC(fit, k=log(N)) <== CAUTION
  #'           5. Jeffreys approximate Bayes factor in Eq. 9 using the normal unit-information prior
  #'           6. Jeffreys approximate Bayes factor in Eq. 9 using the Jeffreys prior
  #'           7. Wagenmakers approximate Bayes factor based on the p-value and sample size in Eq. 10
  
  data <- stats::na.omit(data)
  grp_sizes <- unname(table(data$grp)) # sample sizes, n1 and n2
  Ntot <- sum(grp_sizes) # total number of observations
  N <- prod(grp_sizes) / Ntot # effective sample size in the two-sample t-test, n1 * n2 / (n1 + n2)
  
  fit0 <- stats::lm(resp ~ 1, data) # null model fit
  fit1 <- stats::lm(resp ~ grp, data) # full model fit
  pVal <- summary(fit1)$coefficients["grpgrp2", "Pr(>|t|)"] # p-value of the two-sample t-test
  tVal <- stats::qt(pVal/2, Ntot-2) # t-statistic (m2 >= m1), ["grpgrp2", "t value"]
  
  datalist <- list(N=Ntot, x=as.numeric(data$grp)-1, y=data$resp,
                   mu=stats::coef(fit1), Sigma=stats::vcov(fit1)*Ntot)
  stanfitM1 <- rstan::sampling(stanmodel1M1, data=datalist,
                               iter=15000, warmup=5000, chains=2, seed=277, refresh=0)
  betaS <- rstan::extract(stanfitM1, pars="beta")$beta
  dPost <- polspline::dlogspline(0, polspline::logspline(betaS)) # estimated posterior density
  dPrior <- stats::dnorm(0, datalist$mu[2], sqrt(datalist$Sigma[2,2]))
  SDdr01 <- dPost / dPrior # Savage–Dickey density ratio
  
  bf <- BayesFactor::ttestBF(formula=resp ~ grp, data=data) # Eq. 7
  BF01 <- ifelse(BayesFactor::extractBF(bf)$error < 0.01, # negligible Monte Carlo errors
                 1 / BayesFactor::extractBF(bf)$bf, NA)
  
  BICB01 <- exp((stats::BIC(fit1) - stats::BIC(fit0)) / 2) # equiv. AIC(fit, k=log(n1 + n2)); Eq. 8
  SBCB01 <- exp((stats::AIC(fit1, k=log(N)) - stats::AIC(fit0, k=log(N))) / 2) # <== CAUTION 
  
  N <- ifelse(total, Ntot, N)
  # sqrt(N) * exp(-0.5 * stats::qchisq(1-pVal, 1)) #  <== CAUTION, tiny p-value --> infinity Wald statistic
  JAB01_UI <- sqrt(N) * exp(-0.5 * tVal * tVal) # Eq. 9, unit-information prior

  WAB01 <- ifelse(pVal <= .5,
                  ifelse(pVal > .1,
                         ifelse(precise,
                                4 * pVal^(2/3) * sqrt(N) / 3, # 0.1 < p <= 0.5 (more precise)
                                sqrt(pVal * N)), # 0.1 < p <= 0.5 (simpler)
                         3 * pVal * sqrt(N)), # p <= 0.1
                  pVal^(1/4) * sqrt(N)) # p > 0.5, Eq. 10
  
  c("SD"=SDdr01, "ttestBF"=BF01, "BIC_approx"=BICB01, "SBC_approx"=SBCB01, 
                 "JAB"=JAB01_UI, "JAB_J"=sqrt(pi/2) * JAB01_UI, "WAB"=WAB01)
}

# set.seed(277)
# (test <- simData(10, 15, power=.8))
# computeBFs(test)
```

&nbsp;
&nbsp;

<font size="4"> Simulation Study </font>

We aim to conduct several simulation studies to evaluate the accuracy of approximate objective Bayes factors in their ability to convert $p$-values and $N$ reported in scientific articles to Bayesian measures of evidence.

&nbsp;

Read Cohen's $d$: <https://en.wikipedia.org/wiki/Effect_size#Cohen's_d>

&nbsp;

```{r sim-not-run, eval=F}
useTotalN <- T

nSim <- 1000 # number of simulation runs for each setting
nSubj <- c(10, 20, 30, 100, 500) # sample sizes of the first group
SMD <- c("Null"=0, "Small"=0.2, "Medium"=0.5, "Large"=0.8) # (standardized) mean differences, m2-m1, given sd=1

reportBF <- reportBF_avg <- reportBF_sd <- reportERR <- reportERR_avg <- reportERR_sd <- 
  reportRULE <- reportRULE_H1 <- reportRULE_H0 <- reportRULE_inc <-
  setNames(vector(mode="list", length=length(nSubj) * length(SMD)), # pre-allocation
           apply(expand.grid(names(SMD), nSubj), 1, function(r) paste0("n1 = ", r[2], ", delta = ", r[1])))

index <- 1 # initial list index

for (n1 in nSubj) {
  n2 <- round(n1 * 1.2) # set the sample size of the second group <==
  
  for (delta in SMD) {
    set.seed(n1+delta)
    temp <- t(replicate(nSim, computeBFs(simData(n1, n2, delta=delta), useTotalN))) # replicate the Bayes factors
    reportBF[[index]] <- temp
    reportBF_avg[[index]] <- colMeans(temp, na.rm=T) # mean Bayes factors in favor of the null
    reportBF_sd[[index]] <- apply(na.omit(temp), 2, sd) # sample standard deviations of them
    
    tempERR <- 100* (temp - temp[,1]) / temp[,1] # percent errors; the first column should be all 0
    reportERR[[index]] <- tempERR # (can take in NA)
    reportERR_avg[[index]] <- colMeans(tempERR, na.rm=T) # mean percent errors
    reportERR_sd[[index]] <- apply(na.omit(tempERR), 2, sd) # sample standard deviations of the percent errors
    
    tempRULE_H1 <- temp < 1/3 # support H1
    tempRULE_H0 <- temp > 3 # support H0
    tempRULE_inc <- temp >= 1/3 & temp <= 3 # inconclusive

    # counts of matching decisions for H1
    reportRULE_H1[[index]] <- colSums(tempRULE_H1 * tempRULE_H1[,1], na.rm=T)

    # counts of matching decisions for H0
    reportRULE_H0[[index]] <- colSums(tempRULE_H0 * tempRULE_H0[,1], na.rm=T)

    # counts of matching inconclusiveness
    reportRULE_inc[[index]] <- colSums(tempRULE_inc * tempRULE_inc[,1], na.rm=T)

    tempRULE <- 1*(temp < 1/3) - 1*(temp > 3) # 1: support H1;  0: inconclusive;  -1: support H0
    reportRULE[[index]] <- colSums(tempRULE[,1] == tempRULE, na.rm=T) # counts of matching decisions
    # colMeans(); a value near 1 suggests matching results; the first column should be all 1
    
    index <- index + 1
  }
} # roughly 8.3 hours of run time

# Warning messages:
# 1-10: In polspline::logspline(betaS) :  Not all models could be fitted
```

&nbsp;

```{r prep}
reshapeW2L <- function(list, prob=T, lab=c("SD", "BF", "BIC", "SBC", "JAB", "JAB_J", "WAB")) {
  #' Input -
  #' list:    a list of size length(nSubj) * length(SMD);
  #'          reportERR (unaggregated),
  #'          reportRULE, reportRULE_H1, reportRULE_H0, or reportRULE_inc (aggregated)
  #' prob:    logical; if TRUE (default), return the proportions;
  #'                           otherwise, return the original values
  #' lab:     a vector of characters representing the Bayes factor methods
  #' 
  #' Global -
  #' nSim:    number of simulation runs for each setting
  #' nSubj:   sample sizes of the first group
  #' SMD:     (standardized) mean differences, m2 - m1, given sd = 1
  #' 
  #' Output - unlist the report and reshape the wide into the long
  
  len <- length(lab) # seven methods
  nRep <- ifelse(!prob, nSim, 1) # nSim if unaggregated; 1 if aggregated
  long <- data.frame("value"=unname(unlist(list)),
                     "method"=factor(rep(rep(1:len, each=nRep), length(list)), labels=lab),
                     "n1"=rep(nSubj, each=nRep*len*length(SMD)),
                     "delta"=rep(rep(unname(SMD), each=nRep*len), length(nSubj)))
  
  if (prob) {
    list2 <- lapply(list, function(vec) {
      if (vec[1] == 0) {
        vec * NA
      } else {
        vec / vec[1] # convert counts to proportions
      }
    })
    list3 <- lapply(list, function(vec) rep(vec[1], length(vec))) # denominator
    long$total <- unname(unlist(list3))
    long$prop <- unname(unlist(list2))
  }
  long
}


formatting <- function(val) {
  #' format the value using scientific notation
  if (abs(val) >= 10) {
    paste("~italic(BF)[\"01\"] %~~%", 
          gsub("e\\+0*(\\d+)", "%*%10^\\1", sprintf("%.2e", val)))
  } else if (abs(val) < 1) {
    paste("~italic(BF)[\"01\"] %~~%", 
          gsub("e-0*(\\d+)", "%*%10^{-\\1}", sprintf("%.2e", val)))
  } else {
    paste("~italic(BF)[\"01\"] %~~%", sprintf("%.2f", val))
  }
}
```

&nbsp;

<a href="#top"> scroll to top </a>

&nbsp;
&nbsp;


## 3A. Graphic Results (Effective $N=n_1\cdot n_2/(n_1+n_2)$) {.tabset .tabset-fade}

&nbsp;

<font size="4"> Visualization </font>

Note: The effective number of observations is used in the BF, SBC, JAB, and WAB methods. The Jeffreys approximate Bayes factor (`JAB_J`) in Eq. 9, using the Jeffreys prior, will not be plotted.

### Boxplot (Percent Errors)

```{r load, eval=runthis, echo=F}
load("~/Downloads/ttestSim_totFALSE.RData")
useTotalN <- F
```
```{r viz-boxplot, eval=runthis, fig.width=10, fig.height=5, results="asis"}
# percent errors
df <- reshapeW2L(reportERR, prob=F)

index <- 0

# boxplot of the percent errors
for (smd in SMD) {
  baseBF01 <- sapply(reportBF_avg[seq(1, by=length(SMD), length.out=length(nSubj)) + index], 
                     function(x) x[1]) # baseline is the Savage–Dickey density ratio
  baseBF01_lab <- sapply(baseBF01, formatting) # scientific notation
  index <- index + 1
  
  sub <- subset(df, !(method %in% c("SD", "JAB_J")) & delta == smd) # subset the long
  anno <- data.frame("label"=baseBF01_lab, "n1"=unique(sub$n1),
                     "x"=3, "y"=.8 * aggregate(value~n1, sub, FUN=max)$value) # for annotation use
  
  print(ggplot(sub, aes(x=method, y=value)) +
        geom_boxplot(alpha=0.3, na.rm=T) +
        geom_text(data=anno, aes(label=label, x=x, y=y), parse=T, col="red") +
        facet_wrap(.~n1, scales="free_y", nrow=1,
                     labeller=label_bquote(cols=italic(n)[1]==.(n1))) +
        labs(x="\nBayes Factor Methods", y="Percent Error (%)\n") +
        ggtitle(paste0(names(SMD[SMD==smd]), " Standardized Mean Difference")) +
        geom_hline(yintercept=0, linetype="dashed", color="gray") + # reference line 0%
        theme_classic() +
        theme(axis.text.x=element_text(angle=90)))
  cat("<br><br><br><br><br><br><br>")
}
```

&nbsp;

<a href="#top"> scroll to top </a>

&nbsp;
&nbsp;

### Boxplot ($\textit{BF}_{01}$)

```{r viz-boxplot2, eval=runthis, fig.width=10, fig.height=5, results="asis"}
df_BF <- reshapeW2L(reportBF, prob=F)

# boxplot of the BF₀₁
for (smd in SMD) {
  sub <- subset(df_BF, delta == smd & method != "JAB_J") # subset the long
  print(ggplot(sub, aes(x=method, y=value)) +
          geom_boxplot(alpha=0.3, na.rm=T) +
          facet_wrap(.~n1, scales="free_y", nrow=1,
                     labeller=label_bquote(cols=italic(n)[1]==.(n1))) +
          labs(x="\nBayes Factor Methods", y=expression(italic("BF")["01"])) +
          ggtitle(paste0(names(SMD[SMD==smd]), " Standardized Mean Difference")) +
          theme_classic() +
          theme(axis.text.x=element_text(angle=90)))
  cat("<br><br><br><br><br>")
}
```

&nbsp;

<a href="#top"> scroll to top </a>

&nbsp;
&nbsp;

### Line Chart (Proportions of Agreement)

```{r viz-line, eval=runthis, fig.width=9, fig.height=7, warning=F}
# decisions
decis <- rbind(reshapeW2L(reportRULE),
               reshapeW2L(reportRULE_H1),
               reshapeW2L(reportRULE_H0),
               reshapeW2L(reportRULE_inc))
decis$result <- factor(rep(c("Overall", "BF01 < 1/3", "BF01 > 3", "[1/3, 3]"),
                           each=length(levels(df$method)) * length(nSubj) * length(SMD)),
                       levels=c("Overall", "BF01 < 1/3", "BF01 > 3", "[1/3, 3]"))
decis$delta <- paste0("ES = ", decis$delta)
decis <- decis[!is.na(decis[,"prop"]),]
if (F) {
  # compute the pointwise confidence intervals using the normal approximation
  decis$half.len <- qnorm(1-.05/2) * 
    sqrt(decis$value * (decis$total - decis$value) / decis$total^3) # 95% CI width
  decis$lwr <- max(0, decis$value / decis$total - decis$half.len) # 95% CI lower bound
  decis$upr <- min(1, decis$value / decis$total + decis$half.len) # 95% CI upper bound
} else { # Jeffreys intervals
  decis$lwr <- qbeta(.05/2, decis$value, decis$total - decis$value + 1) # 95% CI lower bound
  decis$upr <- qbeta(1-.05/2, decis$value + 1, decis$total - decis$value) # 95% CI upper bound
  decis$half.len <- (decis$upr - decis$lwr) / 2 # 95% CI width
}

# line chart of the proportions of agreement
ggplotly(ggplot(decis, aes(n1, prop, color=method, label=value)) +
           geom_line(linewidth=1.2, na.rm=T) + geom_point(size=2, na.rm=T) +
           geom_ribbon(aes(ymin=lwr, ymax=upr), alpha=.2) +
           scale_color_manual(values=c("black", "gray", "#E69F00", "#009E73", "#CC79A7", "#D55E00", "#0072B2"),
                              labels=c("SD", "BF", "BIC", "SBC", "JAB", "JAB_J", "WAB")) +
           facet_grid(result~delta) +
           labs(x="Sample Size of the First Group", y="Proportion of Agreement", color="Method") +
           scale_x_continuous(breaks=c(10, 100, 500)) +
           scale_y_continuous(breaks=c(0, 0.5, 1)) + # min(decis$prop, na.rm=T)
           theme_minimal()) # interactive plots
```
&nbsp;
&nbsp;
&nbsp;
```{r viz-line2, eval=runthis, echo=F, fig.width=10, fig.height=10, message=F}
decis <- rbind(reshapeW2L(reportRULE),
               reshapeW2L(reportRULE_H1),
               reshapeW2L(reportRULE_H0),
               reshapeW2L(reportRULE_inc))
decis$result <- factor(rep(c("Overall", "H1", "H0", "Inconclusive"),
                           each=length(levels(df$method)) * length(nSubj) * length(SMD)),
                       levels=c("Overall", "H1", "H0", "Inconclusive"),
                       labels=c("Overall", "italic(BF)['01'] < 1/3", 
                                "italic(BF)['01'] > 3", "italic(BF)['01'] %in% group('[',list(1/3, 3),']')"))
decis$delta <- factor(decis$delta, levels=unname(SMD),
                      labels=paste0("italic(δ) == ", unname(SMD), "~(", names(SMD), ")"))
set.seed(277)
decis_plot <- subset(decis[sample(1:nrow(decis), nrow(decis)),], 
                     !(method %in% c("SD", "JAB_J"))) # shuffle to prevent overlapping

ggplot(decis_plot, aes(n1, prop, color=method, shape=method)) +
  geom_line(linewidth=1.2, na.rm=T) + geom_point(size=3.5, na.rm=T) +
  scale_color_manual(values=c("gray", "#E69F00", "#009E73", "#CC79A7", "#0072B2"),
                     labels=c("BF", "BIC", "SBC", "JAB", "WAB")) +
  scale_shape_manual(values=c(10, 8, 16, 18, 15),
                     labels=c("BF", "BIC", "SBC", "JAB", "WAB")) +
  facet_grid(result~delta, labeller=label_parsed) +
  labs(x="\nSample Size of the First Group", y="Proportion of Agreement\n", color="Method", shape="Method",
       title=expression("Two-Sample " * italic(t) * "-Test"),
       caption=ifelse(useTotalN,
                      "Note: The total number of observations is used in the SD, BIC, JAB, and WAB methods.",
                      "Note: The effective number of observations is used in the BF, SBC, JAB, and WAB methods.")) +
  scale_x_continuous(breaks=c(10, 100, 500)) +
  scale_y_continuous(breaks=c(0, 0.5, 1)) + # the smallest is 0.25 for BIC method under the null
  theme_minimal() +
  theme(legend.position="bottom", text=element_text(size=17),
        legend.title=element_text(margin=margin(0,1,0,0,"cm")),
        legend.text=element_text(margin=margin(0,0.5,0,0.5,"cm"))) +
  guides(color=guide_legend(override.aes=list(linewidth=0)),
         shape=guide_legend(override.aes=list(size=6), nrow=1))
```

&nbsp;

<a href="#top"> scroll to top </a>

&nbsp;
&nbsp;


### **Eureka!**

Since all the methods have solutions in $t$-tests and balanced one-way ANOVA, simulation studies may not be necessary.

```{r theoretical, fig.width=9, fig.height=8}
BF10 <- function(t) {
  gamma <- 1/sqrt(2) # scale parameter of the Cauchy prior
  Neff <- n1*n2/(n1+n2) # effective sample size
  nu <- n1+n2-2 # degrees of freedom
  coef <- (2*pi)^(-0.5) * gamma * (1+t^2/nu)^((nu+1)/2)
  integrand <- function(g) {
    (1+Neff*g)^(-0.5) * (1+t^2/((1+Neff*g)*nu))^(-(nu+1)/2) * g^(-1.5) * exp(-gamma^2/(2*g))
  }
  unname(coef * integrate(integrand, lower=0, upper=Inf)$value)
}

BIC01 <- function(t) {
  N <- n1+n2 # total number of observations
  nu <- N-2 # degrees of freedom
  sqrt(N * (1 + t^2 / nu)^(-N))
}

SBC01 <- function(t) {
  Neff <- n1*n2/(n1+n2) # effective sample size
  N <- n1+n2 # total number of observations
  nu <- N-2 # degrees of freedom
  sqrt(Neff * (1 + t^2 / nu)^(-N))
}

JAB01 <- function(t) {
  Neff <- n1*n2/(n1+n2) # effective sample size
  sqrt(Neff) * exp(-0.5 * t^2)
}

WAB01 <- function(t) {
  Neff <- n1*n2/(n1+n2) # effective sample size
  pVal <- pt(t, n1+n2-2, lower.tail=F) * 2 # two-sided p-value
  ifelse(pVal <= .5,
         ifelse(pVal > .1,
                ifelse(F,
                       4 * pVal^(2/3) * sqrt(Neff) / 3, # 0.1 < p <= 0.5 (more precise)
                       sqrt(pVal * Neff)), # 0.1 < p <= 0.5 (simpler)
                3 * pVal * sqrt(Neff)), # p <= 0.1
         pVal^(1/4) * sqrt(Neff)) # p > 0.5
}

PBF01 <- function(t) {
  N <- n1+n2 # total number of observations
  a <- 2 # number of groups
  F_ratio <- t^2 # F = t²
  zeta <- -0.5 # adjustment of the shape parameters
  ifelse(N >= 345,
         (N/2)^((a-1)/2) * gamma(1+zeta) * 
           ((N-a) / (N-a+(a-1)*F_ratio))^((N-a)/2-1-zeta) / gamma((a-1)/2+1+zeta), # in case of Inf/Inf
         gamma((N-1)/2) * gamma(1+zeta) * 
           ((N-a) / (N-a+(a-1)*F_ratio))^((N-a)/2-1-zeta) / (gamma((a-1)/2+1+zeta) * gamma((N-a)/2)))
}

TSBF01 <- function(t) {
  # Johnson (2005), https://doi.org/10.1111/j.1467-9868.2005.00521.x
  t <- abs(t)
  N <- n1+n2 # total number of observations
  m <- N-2
  ifelse(t > 1,
         ((m+1) / (m+t^2))^((m+1)/2) * t,
         1)
}


n1 <- 30 # sample size of the first group
n2 <- 35 # sample size of the second group
ts <- qt(c(.05, .01, .005)/2, n1+n2-2, lower.tail=F) # critical values of t
x <- seq(0, ts[3], by=0.1) # input sequence
num <- length(x)
bf <- bic <- sbc <- jab <- wab <- pbf <- tsbf <- numeric(num)
for (i in 1:num) {
  bf[i] <- BF10(x[i])     # JZS Bayes factor
  bic[i] <- BIC01(x[i])   # BIC approximation
  sbc[i] <- SBC01(x[i])   # BIC approximation using the effective sample size
  jab[i] <- JAB01(x[i])   # Jeffreys approximate Bayes factor
  wab[i] <- WAB01(x[i])   # Wagenmakers approximate Bayes factor
  pbf[i] <- PBF01(x[i])   # Pearson Bayes factor
  tsbf[i] <- TSBF01(x[i]) # Test-based Bayes factor
}

plot(x, bf, ylab=expression(italic(BF)[10]), xlab=expression("|" * italic("t") * "|-Statistic"),
     main=expression("Two-Sample " * italic(t) * "-Test"),
     sub=bquote(italic(n)[1]==.(n1)~~"&"~~italic(n)[2]==.(n2))) # scatter plot
points(x, 1/bic, col="#E69F00", pch=8)
points(x, 1/sbc, col="#009E73", pch=16)
points(x, 1/jab, col="#CC79A7", pch=18)
points(x, 1/wab, col="#0072B2", pch=15)
points(x, 1/pbf, col="#56B4E9", pch=17)
points(x, 1/tsbf, col="#EF4135", pch=3)
abline(h=c(1/3, 3), col="gray")
abline(v=c(ts[1], ts[2]), col="red")
legend("topleft", c("ttestBF", "BIC approx", "SBC approx", "JAB", "WAB", "PBF", "TSBF"),
       col=c("black", "#E69F00", "#009E73", "#CC79A7", "#0072B2", "#56B4E9", "#EF4135"),
       pch=c(1, 8, 16, 18, 15, 17, 3), lty=0, lwd=2, bty="n")

plot(x, 1/bf, ylab=expression(italic(BF)["01"]), xlab=expression("|" * italic("t") * "|-Statistic"),
     main=expression("Two-Sample " * italic(t) * "-Test"),
     sub=bquote(italic(n)[1]==.(n1)~~"&"~~italic(n)[2]==.(n2))) # scatter plot
points(x, bic, col="#E69F00", pch=8)
points(x, sbc, col="#009E73", pch=16)
points(x, jab, col="#CC79A7", pch=18)
points(x, wab, col="#0072B2", pch=15)
points(x, pbf, col="#56B4E9", pch=17)
points(x, tsbf, col="#EF4135", pch=3)
abline(h=c(1/3, 3), col="gray")
abline(v=c(ts[1], ts[2]), col="red")
legend("bottomleft", c("ttestBF", "BIC approx", "SBC approx", "JAB", "WAB", "PBF", "TSBF"),
       col=c("black", "#E69F00", "#009E73", "#CC79A7", "#0072B2", "#56B4E9", "#EF4135"),
       pch=c(1, 8, 16, 18, 15, 17, 3), lty=0, lwd=2, bty="n")
```

&nbsp;

<a href="#top"> scroll to top </a>

&nbsp;
&nbsp;


## 3B. Graphic Results (Total $N=n_1+n_2$) {.tabset .tabset-fade}

&nbsp;

<font size="4"> Visualization </font>

Note: The total number of observations is used in the SD, BIC, JAB, and WAB methods. The Jeffreys approximate Bayes factor (`JAB_J`) in Eq. 9, using the Jeffreys prior, will not be plotted.

### Boxplot (Percent Errors)

```{r load2, eval=runthis, echo=F}
load("~/Downloads/ttestSim_totTRUE.RData")
useTotalN <- T
```
```{r viz-boxplot3, eval=runthis, fig.width=10, fig.height=5, results="asis", echo=F}
# percent errors
df <- reshapeW2L(reportERR, prob=F)

index <- 0

# boxplot of the percent errors
for (smd in SMD) {
  baseBF01 <- sapply(reportBF_avg[seq(1, by=length(SMD), length.out=length(nSubj)) + index], 
                     function(x) x[1]) # baseline is the Savage–Dickey density ratio
  baseBF01_lab <- sapply(baseBF01, formatting) # scientific notation
  index <- index + 1
  
  sub <- subset(df, !(method %in% c("SD", "JAB_J")) & delta == smd) # subset the long
  anno <- data.frame("label"=baseBF01_lab, "n1"=unique(sub$n1),
                     "x"=3, "y"=.8 * aggregate(value~n1, sub, FUN=max)$value) # for annotation use
  
  print(ggplot(sub, aes(x=method, y=value)) +
        geom_boxplot(alpha=0.3, na.rm=T) +
        geom_text(data=anno, aes(label=label, x=x, y=y), parse=T, col="red") +
        facet_wrap(.~n1, scales="free_y", nrow=1,
                     labeller=label_bquote(cols=italic(n)[1]==.(n1))) +
        labs(x="\nBayes Factor Methods", y="Percent Error (%)\n") +
        ggtitle(paste0(names(SMD[SMD==smd]), " Standardized Mean Difference")) +
        geom_hline(yintercept=0, linetype="dashed", color="gray") + # reference line 0%
        theme_classic() +
        theme(axis.text.x=element_text(angle=90)))
  cat("<br><br><br><br><br><br><br>")
}
```

&nbsp;

<a href="#top"> scroll to top </a>

&nbsp;
&nbsp;

### Boxplot ($\textit{BF}_{01}$)

```{r viz-boxplot4, eval=runthis, fig.width=10, fig.height=5, results="asis", echo=F}
df_BF <- reshapeW2L(reportBF, prob=F)

# boxplot of the BF₀₁
for (smd in SMD) {
  sub <- subset(df_BF, delta == smd & method != "JAB_J") # subset the long
  print(ggplot(sub, aes(x=method, y=value)) +
          geom_boxplot(alpha=0.3, na.rm=T) +
          facet_wrap(.~n1, scales="free_y", nrow=1,
                     labeller=label_bquote(cols=italic(n)[1]==.(n1))) +
          labs(x="\nBayes Factor Methods", y=expression(italic("BF")["01"])) +
          ggtitle(paste0(names(SMD[SMD==smd]), " Standardized Mean Difference")) +
          theme_classic() +
          theme(axis.text.x=element_text(angle=90)))
  cat("<br><br><br><br><br>")
}
```

&nbsp;

<a href="#top"> scroll to top </a>

&nbsp;
&nbsp;

### Line Chart (Proportions of Agreement)

```{r viz-line3, eval=runthis, fig.width=9, fig.height=7, warning=F, echo=F}
# decisions
decis <- rbind(reshapeW2L(reportRULE),
               reshapeW2L(reportRULE_H1),
               reshapeW2L(reportRULE_H0),
               reshapeW2L(reportRULE_inc))
decis$result <- factor(rep(c("Overall", "BF01 < 1/3", "BF01 > 3", "[1/3, 3]"),
                           each=length(levels(df$method)) * length(nSubj) * length(SMD)),
                       levels=c("Overall", "BF01 < 1/3", "BF01 > 3", "[1/3, 3]"))
decis$delta <- paste0("ES = ", decis$delta)
decis <- decis[!is.na(decis[,"prop"]),]
if (F) {
  # compute the pointwise confidence intervals using the normal approximation
  decis$half.len <- qnorm(1-.05/2) * 
    sqrt(decis$value * (decis$total - decis$value) / decis$total^3) # 95% CI width
  decis$lwr <- max(0, decis$value / decis$total - decis$half.len) # 95% CI lower bound
  decis$upr <- min(1, decis$value / decis$total + decis$half.len) # 95% CI upper bound
} else { # Jeffreys intervals
  decis$lwr <- qbeta(.05/2, decis$value, decis$total - decis$value + 1) # 95% CI lower bound
  decis$upr <- qbeta(1-.05/2, decis$value + 1, decis$total - decis$value) # 95% CI upper bound
  decis$half.len <- (decis$upr - decis$lwr) / 2 # 95% CI width
}

# line chart of the proportions of agreement
ggplotly(ggplot(decis, aes(n1, prop, color=method, label=value)) +
           geom_line(linewidth=1.2, na.rm=T) + geom_point(size=2, na.rm=T) +
           geom_ribbon(aes(ymin=lwr, ymax=upr), alpha=.2) +
           scale_color_manual(values=c("black", "gray", "#E69F00", "#009E73", "#CC79A7", "#D55E00", "#0072B2"),
                              labels=c("SD", "BF", "BIC", "SBC", "JAB", "JAB_J", "WAB")) +
           facet_grid(result~delta) +
           labs(x="Sample Size of the First Group", y="Proportion of Agreement", color="Method") +
           scale_x_continuous(breaks=c(10, 100, 500)) +
           scale_y_continuous(breaks=c(0, 0.5, 1)) + # min(decis$prop, na.rm=T)
           theme_minimal()) # interactive plots
```
&nbsp;
&nbsp;
&nbsp;
```{r viz-line4, eval=runthis, echo=F, fig.width=10, fig.height=10, message=F}
decis <- rbind(reshapeW2L(reportRULE),
               reshapeW2L(reportRULE_H1),
               reshapeW2L(reportRULE_H0),
               reshapeW2L(reportRULE_inc))
decis$result <- factor(rep(c("Overall", "H1", "H0", "Inconclusive"),
                           each=length(levels(df$method)) * length(nSubj) * length(SMD)),
                       levels=c("Overall", "H1", "H0", "Inconclusive"),
                       labels=c("Overall", "italic(BF)['01'] < 1/3", 
                                "italic(BF)['01'] > 3", "italic(BF)['01'] %in% group('[',list(1/3, 3),']')"))
decis$delta <- factor(decis$delta, levels=unname(SMD),
                      labels=paste0("italic(δ) == ", unname(SMD), "~(", names(SMD), ")"))
set.seed(277)
decis_plot <- subset(decis[sample(1:nrow(decis), nrow(decis)),], 
                     !(method %in% c("SD", "JAB_J"))) # shuffle to prevent overlapping

ggplot(decis_plot, aes(n1, prop, color=method, shape=method)) +
  geom_line(linewidth=1.2, na.rm=T) + geom_point(size=3.5, na.rm=T) +
  scale_color_manual(values=c("gray", "#E69F00", "#009E73", "#CC79A7", "#0072B2"),
                     labels=c("BF", "BIC", "SBC", "JAB", "WAB")) +
  scale_shape_manual(values=c(10, 8, 16, 18, 15),
                     labels=c("BF", "BIC", "SBC", "JAB", "WAB")) +
  facet_grid(result~delta, labeller=label_parsed) +
  labs(x="\nSample Size of the First Group", y="Proportion of Agreement\n", color="Method", shape="Method",
       title=expression("Two-Sample " * italic(t) * "-Test"),
       caption=ifelse(useTotalN,
                      "Note: The total number of observations is used in the SD, BIC, JAB, and WAB methods.",
                      "Note: The effective number of observations is used in the BF, SBC, JAB, and WAB methods.")) +
  scale_x_continuous(breaks=c(10, 100, 500)) +
  scale_y_continuous(breaks=c(0, 0.5, 1)) + # the smallest is 0.25 for BIC method under the null
  theme_minimal() +
  theme(legend.position="bottom", text=element_text(size=17),
        legend.title=element_text(margin=margin(0,1,0,0,"cm")),
        legend.text=element_text(margin=margin(0,0.5,0,0.5,"cm"))) +
  guides(color=guide_legend(override.aes=list(linewidth=0)),
         shape=guide_legend(override.aes=list(size=6), nrow=1))
```

&nbsp;

<a href="#top"> scroll to top </a>

&nbsp;
&nbsp;


## 4. Finite Sample Correction

&nbsp;

Centering the normal unit-information prior on the null hypothesized value $\theta_0$, the customary choice in objective Bayesian testing yields the following JAB* (its Eq. 4; Wagenmakers, [2022](https://doi.org/10.31234/osf.io/egydq)).

\begin{equation}
  \textit{JAB}_{01}^*\approx\sqrt{N}\exp{\left\{-\frac{N-1}{2}\left(\frac{\hat{\theta}-\theta_0}{\sigma_g}\right)^2\right\}}
\end{equation}

Alternatively, the JAB in Eq. 9 centers the prior on the maximum likelihood estimate.

\begin{equation}
  \textit{JAB}_{01}\approx\sqrt{N}\exp{\left\{-\frac{N}{2}\left(\frac{\hat{\theta}-\theta_0}{\sigma_g}\right)^2\right\}}
\end{equation}

Using a power transformation, it can be shown that $\textit{JAB}_{01}^*=\textit{JAB}_{01}^{\frac{N-1}{N}}\cdot N^{\frac{1}{2N}}$.

&nbsp;

$n_1=10,\quad n_2=12$

```{r (N-1)/N, eval=runthis, fig.width=10, fig.height=5}
par(mfrow=c(1, length(SMD)))
for (i in 1:length(SMD)) {
  reportBF[[i]][,6] <- reportBF[[i]][,5]^(21/22) * 22^(1/44) # total number of observations is 10 + 12
  boxplot(reportBF[[i]][,c(1,5:7)], names=c("SD", "JAB", "JAB*", "WAB"),
          xlab="", ylab=expression(italic("BF")["01"]), main=paste0(names(SMD)[i], " SMD"), las=2)
}
```

&nbsp;

<a href="#top"> scroll to top </a>

&nbsp;
&nbsp;


## 5. eJAB

&nbsp;

To test the hypotheses $\mathcal{H}_0:\boldsymbol{\theta}=\boldsymbol{\theta}_0$ versus $\mathcal{H}_1:\boldsymbol{\theta}\neq\boldsymbol{\theta}_0$, we define the extended Jeffreys approximate objective Bayes factor (eJAB) as
\begin{equation}
  \mathit{eJAB}_{01}=\sqrt{N}\exp\left\{-\frac{1}{2}\frac{N^{1/q}-1}{N^{1/q}}\cdot Q_{\chi^{2}_{q}}(1-p)\right\},
\end{equation}

where $q$ is the size of the parameter vector $\boldsymbol{\theta}$, $N$ is the sample size, $Q_{\chi^{2}_{q}}(\cdot)$ is the quantile function of the chi-squared distribution with $q$ degrees of freedom, and $p$ is the $p$-value from a null-hypothesis significance test. We further define $\mathit{eJAB}_{10}=1\ \!/\ \!\mathit{eJAB}_{01}$.

&nbsp;

The sample size $N$ is

* the total number of observations for for $t$-tests, linear regression, logistic regression, one-way ANOVA, and chi-squared tests.

* the number of events for Cox models.

* the number of independent observations for one-way repeated-measures ANOVA (Nathoo & Masson, [2016](https://dx.doi.org/10.1016/j.jmp.2015.03.003)).

&nbsp;

The degrees of freedom are

* $q=1$ for $t$-tests, linear regression, logistic regression, and Cox models, where $\mathcal{H}_0$ specifies a point-null hypothesis (Wagenmakers, [2022](https://doi.org/10.31234/osf.io/egydq)).

* $q=I-1$ for one-way ANOVA and one-way repeated-measures ANOVA, where $I$ is the number of conditions.

* $q=(R-1)(C-1)$ for chi-squared tests, where $R$ and $C$ are the numbers of rows and columns, respectively.

Since the Bayes factor is the ratio of the marginal likelihoods of two competing models, $q=p_1-p_0$ also represents the difference in the number of free parameters between the alternative and null models.

&nbsp;

For further reading, please visit <https://osf.io/qeprj/>.

&nbsp;

<a href="#top"> scroll to top </a>

&nbsp;
&nbsp;

